---
title: "pile_hi_form_2021"
author: "Xavier Francis"
date: "2025-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
## R Markdown
#load packages
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(data.table) 
library(ggplot2)
library(purrr)
library(here)

```



#Data cleaning chunk 1. Here and in the next several data cleaning chunks we are getting rid of duplicate observations, unknown behaviors, and dealing with undirected behaviors
```{r}
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}


`2021_interactions_raw` <- read_csv(here("ch2 data","2021_interactions_raw.csv"))
raw.allbehavedata <- `2021_interactions_raw`



birdIDs_capture <- read_csv(here("ch2 data","2021_birdIDs.csv")) 
birdIDs <- birdIDs_capture %>%
  dplyr::select(mark_id3) %>%
  filter(mark_id3!="") # all 20 birds





bird.list <- sort(unique(birdIDs$mark_id3)) # change each time
length(bird.list)
bird.list
length(bird.list)
# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")
# list of days with observations
obsday.list <- as.character(unique(raw.allbehavedata$date)) #only days with observations
obsday.list.wday <- as.data.frame(obsday.list)
colnames(obsday.list.wday) <- "date"
obsday.list.wday$nday <- seq(1:length(obsday.list.wday$date))

```



#Data cleaning chunk 2; "other" and "unknown" behaviors
```{r}
behave<-raw.allbehavedata 
unique(behave$behavior)
# other
behave.other<-filter(behave, behavior=="other") %>% dplyr::select(-behavior) %>%
  rename(behavior=other) %>%
  dplyr::select(orderkey, sessionKEY, session_start_timeStamp, date, full_time, time, actor, subject, behavior, everything())
behave.exl.other<-filter(behave, behavior!="other") %>% 
  dplyr::select(-other)
behave.goodother<-bind_rows(behave.exl.other, behave.other)
# unknown
behave.unknown <- filter(behave.goodother, behavior=="unknown") 
behave.unknown$behavior[behave.unknown$behavior=="unknown"] <-"unk"
behave.unknown <- behave.unknown %>%
  unite(behavior, behavior, unknown, sep = "_")
behave$behavior[behave$behavior=="unk agonistic"] <-"unk_agonistic"
behave.exl.unk<-behave.goodother %>% filter(behavior!="unknown") %>%
  dplyr::select(-unknown)
behave.goodunk<-bind_rows(behave.unknown, behave.exl.unk)
behave<-behave.goodunk
head(behave)

```



#Data cleaning chunk 3; prepare cleaned df
```{r}
# keep only observations where the actor is in the list of bird IDs
behave.goodID <- subset(behave, actor %in% bird.list)
# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
behave.goodID <- subset(behave.goodID, subject %in% bird.list)
#finds n behaviors by sessionKEY, date, time, actor, subject, behavior
behave.key <- behave.goodID %>% 
  dplyr::group_by(sessionKEY, date, full_time,actor, subject, behavior) %>% 
  tally() # counts behaviors that were observed within the same minute
#finds max n behaviors per actor by date, time, and behavior type. Return just that summary
behave.maxkey<- behave.key %>% 
  group_by(actor, subject, date, full_time, behavior) %>%
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
head(behave.maxkey)
#finds n disps per actor by date for summarized/trimmed data
behaveXday <- behave.maxkey %>% 
  group_by(actor, subject, date, full_time, behavior) %>% 
  summarise(n=sum(n))
head(behaveXday)
min(behaveXday$n)





unique(behaveXday$behavior)
# behaviors in long hi_format (1 event at a time)
behaveXday.long <- behaveXday %>%
  uncount(n) %>%
  mutate(n=1)
sum(behaveXday.long$n)
sum(behaveXday$n)
head(behaveXday.long)
## DUPLICATE UNDIRECTED BEHAVIORS
# filter the undirected behaviors 
behaveXday.undirected <- behaveXday.long %>% 
  filter(grepl(pattern="each other",tolower(behavior)))
  
behaveXday.undirected1 <-behaveXday.long %>%
  filter( behavior=="s2s") # spar as well?
behaveXday.undirected <- bind_rows(behaveXday.undirected,behaveXday.undirected1 )
# duplicate undirected dataframe and change subject and actor
behaveXday.undirected.dupl <- behaveXday.undirected %>%
  rename(subject=actor, actor=subject) %>%
  dplyr::select(actor, subject, everything())
# bind both dataframes
behaveXday.directed_2021.comb <- bind_rows(behaveXday.undirected, behaveXday.undirected.dupl) 
behaveXday.undir2dir <- behaveXday.directed_2021.comb %>%
  separate( behavior,into = c("behavior", "direction"), sep = "_") %>%
  dplyr::select(-direction)
unique(behaveXday.undir2dir$behavior)
## COMBINE DATAFRAMES
# bind to directed dataframe
behaveXday.directed_2021 <- behaveXday.long %>% 
  filter(!grepl(pattern="each other",tolower(behavior)) & behavior!="s2s")
unique(behaveXday.directed_2021$behavior)
behaveXday.directed_2021 <- bind_rows(behaveXday.directed_2021, behaveXday.undir2dir)
behaveXday.directed_2021$behavior[behaveXday.directed_2021$behavior=="NN"] <- "nn"
behaveXday.directed_2021$behavior[behaveXday.directed_2021$behavior=="alloP"] <- "allop"
behaveXday.directed_2021$behavior[behaveXday.directed_2021$behavior=="allopreen"] <- "allop"
length(behaveXday.directed_2021$actor)
unique(behaveXday.directed_2021$behavior)
12454 + 2587 # 15041
head(behaveXday.directed_2021)

```

#Add obs session to df
```{r}
#add obs session
head(behaveXday.directed_2021)

behaveXday.directed_2021 <- ungroup(behaveXday.directed_2021)
behaveXday.directed_2021 <- arrange(behaveXday.directed_2021, date, full_time)
#procedure to add obs session

#store the first row as a separate df. It will be binded back on later 
row_1_obs <- behaveXday.directed_2021[1,]

#put events side by side. This will allow us to subtract consecutive times from one another to determine whether an event took place on a separate observation session. 
obs_period_pt_1 <- mutate(behaveXday.directed_2021, actor_2 = lead(actor),
                          subject_2 = lead(subject),                                   
                          full_time_2 = lead(full_time), 
                          date2 = lead(date),
                          behavior_2 = lead (behavior)
                          )




#add empty columns for observation period and date period
obs_period_pt_1$obs_period <- NA
obs_period_pt_1$date_period <- NA

q <- 1

#loop to calculate observation sessions. If consecutive events were on the same day but there was a long time between events, it is determined that the 2nd event took place in a separate observation session. The ticker goes up, and is reset if events took place on different days.changed time from 30 min to 35 minutes cuz one instance where 30 min just added an obs session by a bit
for (i in 1:nrow(obs_period_pt_1)) {ifelse(obs_period_pt_1$date[i] == obs_period_pt_1$date2[i]
                                           & obs_period_pt_1$full_time_2[i] - obs_period_pt_1$full_time[i] >= 2100, q <- q+1, q <- q)
  ifelse(obs_period_pt_1$date[i] != obs_period_pt_1$date2[i], q <- 1, q <- q)
  obs_period_pt_1$obs_period [i] <- q
}

#remove original columns and keep mutated columns. These will replace original columns
obs_period_pt_2 <- subset(obs_period_pt_1, select = c(date2,  full_time_2, actor_2,  subject_2, behavior_2, obs_period, date_period  ))

#get rid of the bottom, which will be full of NAs, as the last event in the original data does not have anything that comes after
obs_period_pt_3 <- slice(obs_period_pt_2, 1:(n() - 1))


#rename mutated columns with original names 
setnames(obs_period_pt_3, old = c('date2',  'full_time_2', 'actor_2',  'subject_2','behavior_2'), 
         new = c('date',  'full_time', 'actor',  'subject', 'behavior'))

#add observation period "1" onto first row that we subseted off, as the first event in the data has to take place in obs session 1
row_1_obs$obs_period <- 1 
row_1_obs$date_period <- NA

#ensure that row 1 matches the df
row_1_obs_ready <-  subset(row_1_obs, select = c(date,  full_time, actor,  subject, behavior, obs_period, date_period  ))

#bind first row and df together
obs_period_pt_4 <- rbind(row_1_obs_ready, obs_period_pt_3)
obs_period_pt_4$date_period <- paste(obs_period_pt_4$date,obs_period_pt_4$obs_period)


unique(obs_period_pt_4$obs_period)



head(obs_period_pt_4)
```

#bind bin number and rank period onto raw data
```{r}
# MOPA_power_score_2021 <- read_csv(here("ch2 data","MOPA_bin_rank_hi_form_2021.csv"))
# three_day_bins_2021 <- read_csv(here("ch2 data","2021_3daybins.csv"))
# 
# 
# three_day_bins_2021_just_bin.n.date <- subset(three_day_bins_2021, select = c("bin', "date"))
# 
# #subset for columns of interest
# MOPA_power_score_2021_sub <- subset(MOPA_power_score_2021, select = c("bin", "id", "rank"))
# 
# 
# #filter main df so that df only contains crowds and displacements
# aggresion_sub<- filter(obs_period_pt_4, behavior %in% c("crowd","displace", "chase"))
# 
# 
# 
# #filter for study period
# aggresion_sub <- aggresion_sub[(aggresion_sub$date >= "2021-04-05" & aggresion_sub$date <= "2021-05-18" ),]
# 
# 
# head(MOPA_power_score_2021)
# MOPA_power_score_2021_sub <- subset(MOPA_power_score_2021, select = c("bin", "id", "rank"))
# 
# 
# #bind phase df and rank df onto main aggression df
# aggresion_sub_pt2 <- aggresion_sub %>%
#   left_join(three_day_bins_2021_just_bin.n.date, by = c("date")) %>%
#   arrange(date, full_time) %>%
#   left_join(MOPA_power_score_2021_sub, by = c("bin", "actor"= "id")) %>%
#   rename(actor_rank = rank, actor_power = power) %>%
#   left_join(MOPA_power_score_2021_sub, by = c("bin", "subject"= "id")) %>%
#   rename(subject_rank = rank, subject_power = power)


```








#day version bind bin number and rank on data 
```{r}
MOPA_power_score_2021 <- read_csv(here("ch3 data","MOPA_date_rank_2021.csv"))
three_day_bins_2021 <- read_csv(here("ch2 data","2021_3daybins.csv"))
#calculate power
MOPA_power_score_2021$power <- 1 - MOPA_power_score_2021$dom_ec

three_day_bins_2021_just_bin.n.date <- subset(three_day_bins_2021, select = c("bin", "date"))


#subset for columns of interest
MOPA_power_score_2021_sub <- subset(MOPA_power_score_2021, select = c("date", "id", "power", "rank"))


#filter main df so that df only contains crowds and displacements
aggresion_sub<- filter(obs_period_pt_4, behavior %in% c("crowd","displace", "chase"))



#filter for study period
aggresion_sub <- aggresion_sub[(aggresion_sub$date >= "2021-04-05	" & aggresion_sub$date <= "2021-05-18" ),]




#bindrank df onto main aggression df 
aggression_sub_pt2 <- aggresion_sub %>% 
  left_join(three_day_bins_2021_just_bin.n.date, by = c("date")) %>%
  arrange(date, full_time) %>% 
  left_join(MOPA_power_score_2021_sub, by = c("date", "actor"= "id")) %>%
  rename(actor_rank = rank, actor_power = power) %>%
  left_join(MOPA_power_score_2021_sub, by = c("date", "subject"= "id")) %>%
  rename(subject_rank = rank, subject_power = power)

```



#prepare df for rule detecting
```{r}

aggression_sub_pt2$actor_rank <- as.numeric(aggression_sub_pt2$actor_rank)
aggression_sub_pt2$subject_rank <- as.numeric(aggression_sub_pt2$subject_rank)

#here

#Prepare/order & give each row an id + numeric time
df <- aggression_sub_pt2 %>%
  arrange(date_period, date, full_time) %>%
  mutate(.row = row_number(),
         tsec = as.numeric(full_time))  # seconds since midnight

#Previous-by-actor (used for  base A to B within 60s check) 
by_actor <- df %>%
  group_by(date_period, actor) %>%
  mutate(prev_subj_by_actor = lag(subject),
         prev_tsec_by_actor = lag(tsec)) %>%
  ungroup()

#Last involvement time for each participant (actor OR subject) 
# Build a long view with one entry per participant per event
long <- bind_rows(
  by_actor %>% transmute(date_period, .row, participant = actor, tsec),
  by_actor %>% transmute(date_period, .row, participant = subject, tsec)
) %>%
  arrange(date_period, tsec, .row) %>%
  group_by(date_period, participant) %>%
  mutate(prev_involve_tsec = lag(tsec)) %>%
  ungroup()

#For each original row, get the most recent involvement of either A or B before this row
last_involve <- long %>%
  group_by(date_period, .row) %>%
  summarise(last_involve_tsec = max(prev_involve_tsec, na.rm = TRUE), .groups = "drop") %>%
  mutate(last_involve_tsec = ifelse(is.finite(last_involve_tsec), last_involve_tsec, NA_real_))

#Apply  rule of getting rid of consecutive 60 sec possible chases
aggression_no_repeat_custom <- by_actor %>%
  left_join(last_involve, by = c("date_period", ".row")) %>%
  mutate(
    #Base repeat condition. previous by same actor is the same subject and within 60s
    within_actor_repeat = !is.na(prev_tsec_by_actor) &
                          subject == prev_subj_by_actor &
                          (tsec - prev_tsec_by_actor) <= 60,

    #Detect if there was an intervening event involving A or B after that previous A to B
    interrupted_by_A_or_B = within_actor_repeat &
                            !is.na(last_involve_tsec) &
                            (last_involve_tsec > prev_tsec_by_actor),

    keep = !within_actor_repeat | interrupted_by_A_or_B
  ) %>%
  filter(keep) %>%
dplyr::select(-.row, -tsec,
         -prev_subj_by_actor, -prev_tsec_by_actor,
         -last_involve_tsec, -within_actor_repeat, -interrupted_by_A_or_B, -keep)


obs <- aggression_no_repeat_custom$date_period


dater <- unique(obs)
datex <- c(1: length(unique(obs)))
actorer <- unique(aggression_no_repeat_custom$actor)
actex <- c(1:length(actorer))

#sort by time and date
pile_filt_binded_2021 <- aggression_no_repeat_custom %>%
  arrange(aggression_no_repeat_custom$date, aggression_no_repeat_custom$full_time)


colSums(is.na(pile_filt_binded_2021))

#bottom rows of df are full of NA rows from where bird was removed or didnt aggressively interact. Checked for NAs throughout rest of df. clean through rest so removing final NA rows 
pile_filt_binded_2021 <- na.omit(pile_filt_binded_2021)
#to here
 

pile_aggression_list_2021_again <- list()


#separate data into list of each obs session each day for each obs
for(i in datex) { 
  output <- filter(pile_filt_binded_2021, pile_filt_binded_2021$date_period == dater[i])
      if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
}
  pile_aggression_list_2021_again[[i]] <- output}


#ensure df arranged by date and time
pile_aggression_list_2021_again <- lapply(pile_aggression_list_2021_again, function(x) arrange(x, date, full_time))





#modified data frame to put consecutive events side by side

pile_side_by_side <- lapply(pile_aggression_list_2021_again, function(x) mutate(x, actor_2 = lead(actor), 
                                                                                        subject_2 = lead(subject), 
                                                                                        full_time_2 = lead(full_time), 
                                                                                        date2 = lead(date),
                                                                                  actor_2_rank = lead(actor_rank),
                                                                                  subject_2_rank = lead(subject_rank))) 


```

#Calculate time threshold
```{r}
dateact1 <-  c(1:length(pile_side_by_side))

#rename df for calculation of time to pile in seconds
pile_side_by_side_time_2021 <- pile_side_by_side


#pick out pile along rule while keeping all dataframes in a list (yay!!!)
for(i in dateact1 ) {
  pile_side_by_side_time_2021[[i]] <- mutate(pile_side_by_side_time_2021[[i]], 
                                           initiated =  "yes", 
                                           rule_followed = ifelse(
                                             subject_2 == subject
                                             & actor != actor_2
                                             & full_time_2 - full_time > 0
                                             & date2 == date,
                                             "yes","no"), time_to_pile = full_time_2 - full_time)
  
}




 #combine dataframes into one huge dataframe
 pile_time_huge_2021 <- bind_rows(pile_side_by_side_time_2021, .id = "column_label")
 
 
 

#get df where rule was followed
pile_time_length_hi_form_2021 <- subset(pile_time_huge_2021, rule_followed != "no")
 
 #subset for columns of interest
 pile_time_length_hi_form_2021 <- subset(pile_time_length_hi_form_2021, select = c(date, full_time, actor, subject, 
                                                                         actor_2, subject_2, full_time_2, date2, 
                                                                         initiated, rule_followed, time_to_pile, date_period, bin, behavior,actor_rank, subject_rank))
 

pile_time_length_hi_form_2021$time_to_pile <- as.numeric(pile_time_length_hi_form_2021$time_to_pile)
 


ggplot(pile_time_length_hi_form_2021, aes(x=time_to_pile)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=10,
                 colour="black", fill="white") + 
  geom_density() +   scale_x_continuous(name = "length in s of pile events", limits = c(0,3000)) 


pile_time_length_hi_form_2021


quantile(pile_time_length_hi_form_2021$time_to_pile, probs = c(0.25, 0.5, 0.75, .95))
#25%   4 
#50%   13
#75%   46
#95%  220

```


#calculate observed rule follows
```{r}



#loop to set conditionals to pick out pile

#counter corresponds with which birds dataframe I am looking at and is used to control for false positives 
#for example, if pile is detected for a bird A whos specific dataframe I am not looking at (im viewing bird Bs df) , this may be a false positive, as that bird A's interactions are not all listed in the dataframe of another bird B. In other words, bird A may have done something else to another bird (C, D, F etc.) that is not listed, as it would be only listed in that specific birds dataframe, that being Bird A's own dataframe

dateact1 <-  c(1:length(pile_side_by_side))

pile_side_by_side_thresh <- pile_side_by_side

for(i in dateact1) {
  pile_side_by_side[[i]] <- mutate(pile_side_by_side_thresh[[i]], 
                                           initiated =  "yes", 
                                           rule_followed = ifelse(
                                             subject_2 == subject
                                             & actor != actor_2
                                             & full_time_2 - full_time <= 220  
                                             & full_time_2 - full_time > 0
                                             & date2 == date,
                                             "yes","no"), time_to_pile = full_time_2 - full_time)
} 



#combine dataframes into one huge dataframe
pilehuge_2021 <- bind_rows(pile_side_by_side, .id = "column_label")

#sort by time and date
pilehugesorted_2021 <- pilehuge_2021 %>%
  arrange(pilehuge_2021$date, pilehuge_2021$full_time)




#Remove duplicate events, events that have been represented twice, and events that could not confidently be counted as the start of pile
pilehugesorted_no_dupr_2021 <- subset(pilehugesorted_2021, initiated != "no") 

#subset for desired columns
pile_no_dupr_2021 <- subset(pilehugesorted_no_dupr_2021, select = c(date, full_time, actor, subject, 
                                                                         actor_2, subject_2, full_time_2, date2, 
                                                                         initiated, rule_followed, time_to_pile, date_period, bin, behavior,actor_rank, subject_rank ))

head(pile_no_dupr_2021)


#determine how many time rule was follows
pile_real_2021 <- length(which(pile_no_dupr_2021$rule_followed == "yes"))

pile_real.frame_2021 <- data.frame(pile_real_2021)


pile_real.frame_2021

#create df of only rule follows
pile_doner_all_yes_2021 <- filter(pile_no_dupr_2021, rule_followed == "yes")
```





# day level randomize order of events
```{r}
#randomize order of events


#REMEMBER TO BIND THRESHOLD TO DF
start_time <- format(Sys.time(), "%H:%M:%S")



pile_obs_day_2021 <- unique(pile_filt_binded_2021$date)
pile_count_length_list_masterframe_day_2021 <- list()

pbouter = txtProgressBar(min=0, max = length(pile_obs_day_2021), style = 3)

set.seed(29)

for(k in pile_obs_day_2021) { 
  runouter <- k
  setTxtProgressBar(pbouter, k) #update progress bar
  loop.run_outer <- paste("outer run", runouter) 
  loop.run_outer
  
  pile_filt_binded_2021_rand <- subset(pile_filt_binded_2021, date == k )  
  #remove extra columns that create spam in loop
  pile_filt_binded_2021_rand <- subset(pile_filt_binded_2021_rand, select = c(date, full_time, actor, subject, bin, behavior, obs_period, date_period,  actor_rank, subject_rank ))
  
  
  obs <- pile_filt_binded_2021_rand$date_period
  #separate data into a list with each df=a date so we can randomize within days
  #sort_by_date <- function(x, i, y) {y <- filter(x, date == i )
  #mutate(y, date_counter = i)}
  
  datelist <- list()
  dater <- unique(obs)
  datex <- c(1: length(unique(obs)))
  #boop <- filter(both_numbered_sorted, both_numbered_sorted$date == datex [2])
  
  #loop to create a separate dataframe for each bird that only contains interactions #pertaining to that bird.
  #Each dataframe is stored within a list to make the dataframes easy to work with
  #separate data into list of each obs session each day 
  for(i in datex) { 
    output <- filter(pile_filt_binded_2021_rand, pile_filt_binded_2021_rand$date_period == dater[i] )
    if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
    datelist[[paste0("list", dater[i])]] <- output}
  
  
  
  
  #run ids should be 1 through however many times you want to run through the permutation
  runIDs <- seq(1,1000)
  
  #create an empty masterframe to be filled with each run of the permutation
  masterframe_day_pile_2021.hi_form.df <- data.frame(runIDs = character(),
                                                 actor = character(),
                                                 actor_rank = integer(),
                                                 subject = character(),
                                                 subject_rank = integer(),
                                                 date = character(),
                                                 full_time = integer(),
                                                 bin = integer(),
                                                 behavior = character(),
                                                 actor_2 = character(),
                                                 actor_2_rank = integer(),
                                                 subject_2 = character(),
                                                 subject_2_rank = integer(),
                                                 full_time_2 = integer(),
                                                 date2 = character(),
                                                 counter = character(),
                                                 initiated = character(),
                                                 rule_followed = character(),
                                                 time_to_pile = integer(),
                                                 stringsAsFactors=FALSE)
  nruns <- 1000
  #pb = txtProgressBar(min=0, max = nruns, style = 3)
  
  #forloop starts here'
  for (r in 1:length(runIDs)) { #r = 1 is a test you can uncomment to see if it works before #running it 100 times
    #r = 1
    run <- r
    #setTxtProgressBar(pb, r) #update progress bar
    loop.run <- paste("inner run", run) 
    loop.run
    #for example, if r = 1, looprun will say run 1 which will allow you to keep each run's #data straight
    
    
    #loop.seed <- set.seed[r] didnt work so ignore. not needed
    
    
    #renamed data
    loop.datalist <- datelist
    xorder <- c()
    listlength <- as.numeric(length(loop.datalist))
    datex2 <- 1:listlength
    full_time_list <-c()
    loop.data.randomized.time.added <- c()
    #get rid of anything thats not actor and receiver (we will bind back on the date and time #later anyways to preserve
    #their order)
    
    #remove date from each df in the list
    #create a randomized sequence of numbers to bind to the subset data (ex. 2,6,1,3,5,4)
    #sort the data frame to put the randomized numbers in order from 1 to whatever and voila, #the order of events has been randomized (ex. from above 1,2,3,4,5,6)
    
    for(i in datex2) {
      full_time_list[[i]] <- subset(loop.datalist[[i]], select = c(full_time))
      loop.datalist[[i]] <- subset(loop.datalist[[i]], select = -c(full_time))
      
      xorder[[i]] <- data.frame(xorder = sample(1:length(loop.datalist[[i]]$actor),
                                                length(loop.datalist[[i]]$actor)))
      loop.datalist[[i]] <- mutate(loop.datalist[[i]], 
                                   xorderx = xorder[[i]]$xorder)
      loop.datalist[[i]] <-
        loop.datalist[[i]][order(loop.datalist[[i]]$xorderx),]
      #bind time back on
      loop.data.randomized.time.added[[i]] <- bind_cols(loop.datalist[[i]], full_time_list[[i]])
      loop.data.randomized.time.added[[i]] <- subset(loop.data.randomized.time.added[[i]], select = -c(xorderx))
      
    }
    
    practdrlist <- loop.data.randomized.time.added
    
    

    
    
    pile_2021_rand_split_by_bird_ordered <- lapply(practdrlist, function(x) arrange(x, date, full_time))
    
    
    
    
    #modified each data frame to put consecutive events side by side
    
    pile_rand_test_2021 <- lapply(pile_2021_rand_split_by_bird_ordered, function(x) mutate(x, actor_2 = lead(actor), subject_2 = lead(subject), full_time_2 = lead(full_time), date2 = lead(date), actor_2_rank = lead(actor_rank), subject_2_rank = lead(subject_rank))) 
    
    
    
    
    #successfully picked out pile rule while keeping all dataframes in a list (yay!!!)
    #loop to set conditionals to pick out pile
    
    #counter corresponds with which birds dataframe I am looking at and is used to control for false positives 
    #for example, if pile is detected for a bird whos specific dataframe I am not looking #at , this may be a false positive, as that birds interactions are not all listed in the #dataframe of another bird. In other words, the bird may have done something else to another #bird that is not listed, as it would be only listed in that specific birds dataframe
    dateact2 <-  c(1:length(pile_rand_test_2021))
    
    
    
    
    for(i in dateact2) {
      pile_rand_test_2021[[i]] <- mutate(pile_rand_test_2021[[i]], 
                                           initiated =  "yes", 
                                           rule_followed = ifelse(
                                             subject_2 == subject
                                             & actor != actor_2
                                             & full_time_2 - full_time <=  220
                                             & full_time_2 - full_time > 0
                                             & date2 == date,
                                             "yes","no"), time_to_pile = full_time_2 - full_time)
    } 
    
    
    
    
    
    
    
    
    
    #combine dataframes into one huge dataframe
    testhuge <- bind_rows(pile_rand_test_2021, .id = "column_label")
    
    #sort by time and date
    testhugesorted <- testhuge %>%
      arrange(testhuge$date, testhuge$full_time)
    
    
    
    #need to check but i think it works to get rid of all "initiated no" rows to get rid of #duplicates, as pile 
    #is literally possible whenever an aggressive event takes place
    #Update. It does work
    
    testhugesorted_no_dup <- subset(testhugesorted, initiated != "no") 
    
    pile_no_dup_rand_2021 <- subset(testhugesorted_no_dup, select = c(date, full_time, actor, actor_rank, subject, subject_rank, bin, behavior, actor_2, actor_2_rank, subject_2, subject_2_rank, full_time_2, date2, initiated, rule_followed, time_to_pile))
    
    
    
    
    
    
    #end of pile process
    #end of rule code
    
    
    
    #bind vector of run IDs for each run to the dataframe so that you can have a numeric run #id next to each data point
    #that coresponds with that run which will allow you to sort/put them in ascending order later
    runIDs <- rep(loop.run, length(pile_no_dup_rand_2021$actor))
    
    pile_complete <- cbind(runIDs, pile_no_dup_rand_2021)
    
    
    #put each runs data into teh masterframe
    masterframe_day_pile_2021.hi_form.df <- rbind(masterframe_day_pile_2021.hi_form.df, pile_complete)
    
  }
  pile_count_length_list_masterframe_day_2021[[paste0("day", k)]] <- masterframe_day_pile_2021.hi_form.df
}

pile_count_length_list_masterframe_day_2021_binded <-  bind_rows(pile_count_length_list_masterframe_day_2021, .id = "column_label")


end_time <- format(Sys.time(), "%H:%M:%S")
start_time
end_time


#write.csv(pile_count_length_list_masterframe_day_2021_binded,  here("ch3 data","pile_countmaster_day_level_hi_form_2021_custom_thresh.csv"), row.names = FALSE)

```

```{r}


#pile_count_length_list_masterframe_day_2021_binded <- read_csv(here("ch3 data",("pile_countmaster_day_level_hi_form_2021_custom_thresh.csv")))
pile_doner_all_yes_2021

head(pile_doner_all_yes_2021) 



pile_count_length_sum_2021 <- pile_doner_all_yes_2021 %>% group_by( date) %>%
  dplyr::summarize (
    count =n())

pile_count_length_sum_2021


#write.csv(pile_count_length_sum_2021,  here("ch2 data","pile_count_length_sum_2021.csv"), row.names = FALSE)



pile_dates_2021 <- unique(pile_count_length_list_masterframe_day_2021_binded$date)
```



#day1
```{r}
#day1

current_date <- pile_dates_2021[1]
current_date



pile_count_length_list_masterframe_day_2021_binded_day1 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day1_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day_2021_binded_day1$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day_2021_binded_day1$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day_2021_binded_day1$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day1 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day1_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day1

hi_form_pile_tally2021_day1 <-length(which(pile_count_length_list_2021_day1_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day1 <-   1 - hi_form_pile_tally2021_day1/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day1

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day1 <- (pile_count_length_list_2021_day1_real$count - mean_perm) / sd_perm



#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day1 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day_2021_binded_day1$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.061 ,
  z_score = z_score_day1,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day1 <- rbind(hi_form_df_piece_day1, new_row)

```

#day2
```{r}
#day2

current_date <- pile_dates_2021[2]
current_date



pile_count_length_list_masterframe_day2_2021_binded_day2 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day2_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day2_2021_binded_day2$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day2_2021_binded_day2$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day2_2021_binded_day2$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day2 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day2_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day2

hi_form_pile_tally2021_day2 <-length(which(pile_count_length_list_2021_day2_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day2 <-   1 - hi_form_pile_tally2021_day2/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day2

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day2 <- (pile_count_length_list_2021_day2_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day2 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day2_2021_binded_day2$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.115 ,
  z_score = z_score_day2,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day2 <- rbind(hi_form_df_piece_day2, new_row)

```

#day3
```{r}
#day3

current_date <- pile_dates_2021[3]
current_date



pile_count_length_list_masterframe_day3_2021_binded_day3 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day3_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day3_2021_binded_day3$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day3_2021_binded_day3$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day3_2021_binded_day3$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day3 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day3_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day3

hi_form_pile_tally2021_day3 <-length(which(pile_count_length_list_2021_day3_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day3 <-   1 - hi_form_pile_tally2021_day3/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day3

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day3 <- (pile_count_length_list_2021_day3_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day3 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day3_2021_binded_day3$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.016 ,
  z_score = z_score_day3,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day3 <- rbind(hi_form_df_piece_day3, new_row)

```

#day4
```{r}
#day4

current_date <- pile_dates_2021[4]
current_date



pile_count_length_list_masterframe_day4_2021_binded_day4 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day4_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day4_2021_binded_day4$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day4_2021_binded_day4$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day4_2021_binded_day4$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day4 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day4_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day4

hi_form_pile_tally2021_day4 <-length(which(pile_count_length_list_2021_day4_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day4 <-   1 - hi_form_pile_tally2021_day4/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day4

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day4 <- (pile_count_length_list_2021_day4_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day4 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day4_2021_binded_day4$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.001 ,
  z_score = z_score_day4,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day4 <- rbind(hi_form_df_piece_day4, new_row)

```

#day5
```{r}
#day5

current_date <- pile_dates_2021[5]
current_date



pile_count_length_list_masterframe_day5_2021_binded_day5 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day5_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day5_2021_binded_day5$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day5_2021_binded_day5$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day5_2021_binded_day5$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day5 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day5_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day5

hi_form_pile_tally2021_day5 <-length(which(pile_count_length_list_2021_day5_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day5 <-   1 - hi_form_pile_tally2021_day5/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day5

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day5 <- (pile_count_length_list_2021_day5_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day5 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day5_2021_binded_day5$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.564 ,
  z_score = z_score_day5,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day5 <- rbind(hi_form_df_piece_day5, new_row)

```

#day6
```{r}
#day6

current_date <- pile_dates_2021[6]
current_date



pile_count_length_list_masterframe_day6_2021_binded_day6 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day6_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day6_2021_binded_day6$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day6_2021_binded_day6$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day6_2021_binded_day6$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day6 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day6_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day6

hi_form_pile_tally2021_day6 <-length(which(pile_count_length_list_2021_day6_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day6 <-   1 - hi_form_pile_tally2021_day6/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day6

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day6 <- (pile_count_length_list_2021_day6_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day6 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day6_2021_binded_day6$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day6,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day6 <- rbind(hi_form_df_piece_day6, new_row)

```

#day7
```{r}
#day7

current_date <- pile_dates_2021[7]
current_date



pile_count_length_list_masterframe_day7_2021_binded_day7 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day7_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day7_2021_binded_day7$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day7_2021_binded_day7$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day7_2021_binded_day7$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day7 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept =  pile_count_length_list_2021_day7_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day7

hi_form_pile_tally2021_day7 <-length(which(pile_count_length_list_2021_day7_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day7 <-   1 - hi_form_pile_tally2021_day7/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day7

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day7 <- (pile_count_length_list_2021_day7_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day7 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day7_2021_binded_day7$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day7,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day7 <- rbind(hi_form_df_piece_day7, new_row)

```

#day8
```{r}
#day8

current_date <- pile_dates_2021[8]
current_date



pile_count_length_list_masterframe_day8_2021_binded_day8 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day8_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day8_2021_binded_day8$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day8_2021_binded_day8$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day8_2021_binded_day8$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day8 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day8_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day8

hi_form_pile_tally2021_day8 <-length(which(pile_count_length_list_2021_day8_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day8 <-   1 - hi_form_pile_tally2021_day8/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day8

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day8 <- (pile_count_length_list_2021_day8_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day8 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day8_2021_binded_day8$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.001 ,
  z_score = z_score_day8,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day8 <- rbind(hi_form_df_piece_day8, new_row)

```

#day9
```{r}
#day9

current_date <- pile_dates_2021[9]
current_date



pile_count_length_list_masterframe_day9_2021_binded_day9 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day9_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day9_2021_binded_day9$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day9_2021_binded_day9$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day9_2021_binded_day9$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day9 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day9_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day9

hi_form_pile_tally2021_day9 <-length(which(pile_count_length_list_2021_day9_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day9 <-   1 - hi_form_pile_tally2021_day9/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day9

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day9 <- (pile_count_length_list_2021_day9_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day9 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day9_2021_binded_day9$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0,
  z_score = z_score_day9,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day9 <- rbind(hi_form_df_piece_day9, new_row)

```

#day10
```{r}
#day10

current_date <- pile_dates_2021[10]
current_date



pile_count_length_list_masterframe_day10_2021_binded_day10 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day10_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day10_2021_binded_day10$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day10_2021_binded_day10$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day10_2021_binded_day10$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day10 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day10_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day10

hi_form_pile_tally2021_day10 <-length(which(pile_count_length_list_2021_day10_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day10 <-   1 - hi_form_pile_tally2021_day10/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day10

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day10 <- (pile_count_length_list_2021_day10_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day10 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day10_2021_binded_day10$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.181 ,
  z_score = z_score_day10,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day10 <- rbind(hi_form_df_piece_day10, new_row)

```

#day11
```{r}
#day11

current_date <- pile_dates_2021[11]
current_date



pile_count_length_list_masterframe_day11_2021_binded_day11 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day11_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day11_2021_binded_day11$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day11_2021_binded_day11$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day11_2021_binded_day11$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day11 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day11_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day11

hi_form_pile_tally2021_day11 <-length(which(pile_count_length_list_2021_day11_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day11 <-   1 - hi_form_pile_tally2021_day11/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day11

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day11 <- (pile_count_length_list_2021_day11_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day11 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day11_2021_binded_day11$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.001 ,
  z_score = z_score_day11,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day11 <- rbind(hi_form_df_piece_day11, new_row)

```

#day12
```{r}
#day12

current_date <- pile_dates_2021[12]
current_date



pile_count_length_list_masterframe_day12_2021_binded_day12 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day12_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day12_2021_binded_day12$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day12_2021_binded_day12$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day12_2021_binded_day12$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day12 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day12_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day12

hi_form_pile_tally2021_day12 <-length(which(pile_count_length_list_2021_day12_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day12 <-   1 - hi_form_pile_tally2021_day12/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day12

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day12 <- (pile_count_length_list_2021_day12_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day12 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day12_2021_binded_day12$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.015 ,
  z_score = z_score_day12,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day12 <- rbind(hi_form_df_piece_day12, new_row)

```

#day13
```{r}
#day13

current_date <- pile_dates_2021[13]
current_date



pile_count_length_list_masterframe_day13_2021_binded_day13 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day13_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day13_2021_binded_day13$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day13_2021_binded_day13$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day13_2021_binded_day13$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day13 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day13_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day13

hi_form_pile_tally2021_day13 <-length(which(pile_count_length_list_2021_day13_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day13 <-   1 - hi_form_pile_tally2021_day13/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day13

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day13 <- (pile_count_length_list_2021_day13_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day13 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day13_2021_binded_day13$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls =0.001 ,
  z_score = z_score_day13,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day13 <- rbind(hi_form_df_piece_day13, new_row)

```

#day14
```{r}
#day14

current_date <- pile_dates_2021[14]
current_date



pile_count_length_list_masterframe_day14_2021_binded_day14 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day14_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day14_2021_binded_day14$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day14_2021_binded_day14$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day14_2021_binded_day14$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day14 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day14_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day14

hi_form_pile_tally2021_day14 <-length(which(pile_count_length_list_2021_day14_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day14 <-   1 - hi_form_pile_tally2021_day14/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day14

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day14 <- (pile_count_length_list_2021_day14_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day14 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day14_2021_binded_day14$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day14,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day14 <- rbind(hi_form_df_piece_day14, new_row)

```

#day15
```{r}
#day15

current_date <- pile_dates_2021[15]
current_date



pile_count_length_list_masterframe_day15_2021_binded_day15 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day15_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day15_2021_binded_day15$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day15_2021_binded_day15$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day15_2021_binded_day15$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day15 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day15_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day15

hi_form_pile_tally2021_day15 <-length(which(pile_count_length_list_2021_day15_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day15 <-   1 - hi_form_pile_tally2021_day15/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day15

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day15 <- (pile_count_length_list_2021_day15_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day15 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day15_2021_binded_day15$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.058 ,
  z_score = z_score_day15,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day15 <- rbind(hi_form_df_piece_day15, new_row)

```

#day16
```{r}
#day16

current_date <- pile_dates_2021[16]
current_date



pile_count_length_list_masterframe_day16_2021_binded_day16 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day16_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day16_2021_binded_day16$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day16_2021_binded_day16$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day16_2021_binded_day16$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day16 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day16_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day16

hi_form_pile_tally2021_day16 <-length(which(pile_count_length_list_2021_day16_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day16 <-   1 - hi_form_pile_tally2021_day16/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day16

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day16 <- (pile_count_length_list_2021_day16_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day16 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day16_2021_binded_day16$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.013 ,
  z_score = z_score_day16,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day16 <- rbind(hi_form_df_piece_day16, new_row)

```

#day17
```{r}
#day17

current_date <- pile_dates_2021[17]
current_date



pile_count_length_list_masterframe_day17_2021_binded_day17 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day17_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day17_2021_binded_day17$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day17_2021_binded_day17$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day17_2021_binded_day17$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day17 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day17_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day17

hi_form_pile_tally2021_day17 <-length(which(pile_count_length_list_2021_day17_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day17 <-   1 - hi_form_pile_tally2021_day17/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day17

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day17 <- (pile_count_length_list_2021_day17_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day17 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day17_2021_binded_day17$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day17,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day17 <- rbind(hi_form_df_piece_day17, new_row)

```

#day18
```{r}
#day18

current_date <- pile_dates_2021[18]
current_date



pile_count_length_list_masterframe_day18_2021_binded_day18 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day18_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day18_2021_binded_day18$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day18_2021_binded_day18$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day18_2021_binded_day18$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day18 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day18_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day18

hi_form_pile_tally2021_day18 <-length(which(pile_count_length_list_2021_day18_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day18 <-   1 - hi_form_pile_tally2021_day18/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day18

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day18 <- (pile_count_length_list_2021_day18_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day18 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day18_2021_binded_day18$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.596 ,
  z_score = z_score_day18,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day18 <- rbind(hi_form_df_piece_day18, new_row)

```

#day19
```{r}
#day19

current_date <- pile_dates_2021[19]
current_date



pile_count_length_list_masterframe_day19_2021_binded_day19 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day19_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day19_2021_binded_day19$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day19_2021_binded_day19$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day19_2021_binded_day19$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day19 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day19_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day19

hi_form_pile_tally2021_day19 <-length(which(pile_count_length_list_2021_day19_real$count> rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day19 <-   1 - hi_form_pile_tally2021_day19/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day19

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day19 <- (pile_count_length_list_2021_day19_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day19 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day19_2021_binded_day19$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.001 ,
  z_score = z_score_day19,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day19 <- rbind(hi_form_df_piece_day19, new_row)

```

#day20
```{r}
#day20

current_date <- pile_dates_2021[20]
current_date



pile_count_length_list_masterframe_day20_2021_binded_day20 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day20_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day20_2021_binded_day20$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day20_2021_binded_day20$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day20_2021_binded_day20$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day20 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day20_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day20

hi_form_pile_tally2021_day20 <-length(which(pile_count_length_list_2021_day20_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day20 <-   1 - hi_form_pile_tally2021_day20/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day20

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day20 <- (pile_count_length_list_2021_day20_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day20 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day20_2021_binded_day20$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.076 ,
  z_score = z_score_day20,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day20 <- rbind(hi_form_df_piece_day20, new_row)

```

#day21
```{r}
#day21

current_date <- pile_dates_2021[21]
current_date



pile_count_length_list_masterframe_day21_2021_binded_day21 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day21_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day21_2021_binded_day21$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day21_2021_binded_day21$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day21_2021_binded_day21$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day21 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day21_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day21

hi_form_pile_tally2021_day21 <-length(which(pile_count_length_list_2021_day21_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day21 <-   1 - hi_form_pile_tally2021_day21/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day21

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day21 <- (pile_count_length_list_2021_day21_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day21 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day21_2021_binded_day21$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day21,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day21 <- rbind(hi_form_df_piece_day21, new_row)

```

#day22
```{r}
#day22

current_date <- pile_dates_2021[22]
current_date



pile_count_length_list_masterframe_day22_2021_binded_day22 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day22_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day22_2021_binded_day22$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day22_2021_binded_day22$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day22_2021_binded_day22$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day22 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day22_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day22

hi_form_pile_tally2021_day22 <-length(which(pile_count_length_list_2021_day22_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day22 <-   1 - hi_form_pile_tally2021_day22/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day22

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day22 <- (pile_count_length_list_2021_day22_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day22 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day22_2021_binded_day22$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day22,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day22 <- rbind(hi_form_df_piece_day22, new_row)

```

#day23
```{r}
#day23

current_date <- pile_dates_2021[23]
current_date



pile_count_length_list_masterframe_day23_2021_binded_day23 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day23_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day23_2021_binded_day23$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day23_2021_binded_day23$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day23_2021_binded_day23$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day23 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day23_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day23

hi_form_pile_tally2021_day23 <-length(which(pile_count_length_list_2021_day23_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day23 <-   1 - hi_form_pile_tally2021_day23/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day23

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day23 <- (pile_count_length_list_2021_day23_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day23 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day23_2021_binded_day23$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day23,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day23 <- rbind(hi_form_df_piece_day23, new_row)

```

#day24
```{r}
#day24

current_date <- pile_dates_2021[24]
current_date



pile_count_length_list_masterframe_day24_2021_binded_day24 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day24_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day24_2021_binded_day24$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day24_2021_binded_day24$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day24_2021_binded_day24$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day24 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day24_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day24

hi_form_pile_tally2021_day24 <-length(which(pile_count_length_list_2021_day24_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day24 <-   1 - hi_form_pile_tally2021_day24/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day24

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day24 <- (pile_count_length_list_2021_day24_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day24 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day24_2021_binded_day24$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day24,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day24 <- rbind(hi_form_df_piece_day24, new_row)

```

#day25
```{r}
#day25

current_date <- pile_dates_2021[25]
current_date



pile_count_length_list_masterframe_day25_2021_binded_day25 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day25_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day25_2021_binded_day25$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day25_2021_binded_day25$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day25_2021_binded_day25$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day25 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day25_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day25

hi_form_pile_tally2021_day25 <-length(which(pile_count_length_list_2021_day25_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day25 <-   1 - hi_form_pile_tally2021_day25/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day25

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day25 <- (pile_count_length_list_2021_day25_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day25 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day25_2021_binded_day25$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.028 ,
  z_score = z_score_day25,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day25 <- rbind(hi_form_df_piece_day25, new_row)

```

#day26
```{r}
#day26

current_date <- pile_dates_2021[26]
current_date



pile_count_length_list_masterframe_day26_2021_binded_day26 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day26_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day26_2021_binded_day26$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day26_2021_binded_day26$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day26_2021_binded_day26$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day26 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day26_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day26

hi_form_pile_tally2021_day26 <-length(which(pile_count_length_list_2021_day26_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day26 <-   1 - hi_form_pile_tally2021_day26/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day26

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day26 <- (pile_count_length_list_2021_day26_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day26 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day26_2021_binded_day26$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day26,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day26 <- rbind(hi_form_df_piece_day26, new_row)

```

#day27
```{r}
#day27

current_date <- pile_dates_2021[27]
current_date



pile_count_length_list_masterframe_day27_2021_binded_day27 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day27_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day27_2021_binded_day27$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day27_2021_binded_day27$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day27_2021_binded_day27$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day27 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day27_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day27

hi_form_pile_tally2021_day27 <-length(which(pile_count_length_list_2021_day27_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day27 <-   1 - hi_form_pile_tally2021_day27/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day27

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day27 <- (pile_count_length_list_2021_day27_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day27 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day27_2021_binded_day27$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day27,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day27 <- rbind(hi_form_df_piece_day27, new_row)

```

#day28
```{r}
#day28

current_date <- pile_dates_2021[28]
current_date



pile_count_length_list_masterframe_day28_2021_binded_day28 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day28_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day28_2021_binded_day28$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day28_2021_binded_day28$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day28_2021_binded_day28$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day28 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day28_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day28

hi_form_pile_tally2021_day28 <-length(which(pile_count_length_list_2021_day28_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day28 <-   1 - hi_form_pile_tally2021_day28/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day28

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day28 <- (pile_count_length_list_2021_day28_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day28 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day28_2021_binded_day28$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.005 ,
  z_score = z_score_day28,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day28 <- rbind(hi_form_df_piece_day28, new_row)

```

#day29
```{r}
#day29

current_date <- pile_dates_2021[29]
current_date



pile_count_length_list_masterframe_day29_2021_binded_day29 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day29_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day29_2021_binded_day29$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day29_2021_binded_day29$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day29_2021_binded_day29$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day29 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day29_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day29

hi_form_pile_tally2021_day29 <-length(which(pile_count_length_list_2021_day29_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day29 <-   1 - hi_form_pile_tally2021_day29/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day29

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day29 <- (pile_count_length_list_2021_day29_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day29 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day29_2021_binded_day29$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.003 ,
  z_score = z_score_day29,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day29 <- rbind(hi_form_df_piece_day29, new_row)

```

#day30
```{r}
#day30

current_date <- pile_dates_2021[30]
current_date



pile_count_length_list_masterframe_day30_2021_binded_day30 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day30_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day30_2021_binded_day30$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day30_2021_binded_day30$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day30_2021_binded_day30$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day30 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day30_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day30

hi_form_pile_tally2021_day30 <-length(which(pile_count_length_list_2021_day30_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day30 <-   1 - hi_form_pile_tally2021_day30/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day30

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day30 <- (pile_count_length_list_2021_day30_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day30 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day30_2021_binded_day30$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.004 ,
  z_score = z_score_day30,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day30 <- rbind(hi_form_df_piece_day30, new_row)

```

#day31
```{r}
#day31

current_date <- pile_dates_2021[31]
current_date



pile_count_length_list_masterframe_day31_2021_binded_day31 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day31_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day31_2021_binded_day31$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day31_2021_binded_day31$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day31_2021_binded_day31$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day31 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day31_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day31

hi_form_pile_tally2021_day31 <-length(which(pile_count_length_list_2021_day31_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day31 <-   1 - hi_form_pile_tally2021_day31/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day31

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day31 <- (pile_count_length_list_2021_day31_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day31 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day31_2021_binded_day31$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day31,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day31 <- rbind(hi_form_df_piece_day31, new_row)

```

#day32
```{r}
#day32

current_date <- pile_dates_2021[32]
current_date



pile_count_length_list_masterframe_day32_2021_binded_day32 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day32_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day32_2021_binded_day32$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day32_2021_binded_day32$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day32_2021_binded_day32$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day32 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day32_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day32

hi_form_pile_tally2021_day32 <-length(which(pile_count_length_list_2021_day32_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day32 <-   1 - hi_form_pile_tally2021_day32/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day32

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day32 <- (pile_count_length_list_2021_day32_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day32 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day32_2021_binded_day32$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.129 ,
  z_score = z_score_day32,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day32 <- rbind(hi_form_df_piece_day32, new_row)

```

#day33
```{r}
#day33

current_date <- pile_dates_2021[33]
current_date



pile_count_length_list_masterframe_day33_2021_binded_day33 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day33_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day33_2021_binded_day33$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day33_2021_binded_day33$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day33_2021_binded_day33$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day33 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day33_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day33

hi_form_pile_tally2021_day33 <-length(which(pile_count_length_list_2021_day33_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day33 <-   1 - hi_form_pile_tally2021_day33/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day33

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day33 <- (pile_count_length_list_2021_day33_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day33 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day33_2021_binded_day33$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day33,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day33 <- rbind(hi_form_df_piece_day33, new_row)

```

#day34
```{r}
#day34

current_date <- pile_dates_2021[34]
current_date



pile_count_length_list_masterframe_day34_2021_binded_day34 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day34_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day34_2021_binded_day34$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day34_2021_binded_day34$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day34_2021_binded_day34$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day34 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day34_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day34

hi_form_pile_tally2021_day34 <-length(which(pile_count_length_list_2021_day34_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day34 <-   1 - hi_form_pile_tally2021_day34/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day34

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day34 <- (pile_count_length_list_2021_day34_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day34 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day34_2021_binded_day34$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0 ,
  z_score = z_score_day34,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day34 <- rbind(hi_form_df_piece_day34, new_row)

```

#day35
```{r}
#day35

current_date <- pile_dates_2021[35]
current_date



pile_count_length_list_masterframe_day35_2021_binded_day35 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day35_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day35_2021_binded_day35$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day35_2021_binded_day35$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day35_2021_binded_day35$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day35 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day35_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day35

hi_form_pile_tally2021_day35 <-length(which(pile_count_length_list_2021_day35_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day35 <-   1 - hi_form_pile_tally2021_day35/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day35

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day35 <- (pile_count_length_list_2021_day35_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day35 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day35_2021_binded_day35$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "same",
  p_cmls = 0.121 ,
  z_score = z_score_day35,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day35 <- rbind(hi_form_df_piece_day35, new_row)

```

#day36
```{r}
#day36

current_date <- pile_dates_2021[36]
current_date



pile_count_length_list_masterframe_day36_2021_binded_day36 <- subset(pile_count_length_list_masterframe_day_2021_binded, date == current_date)

pile_count_length_list_2021_day36_real <- subset(pile_count_length_sum_2021, date == current_date)




#put run #s in an ascending vector
cum_run_IDS <- unique(pile_count_length_list_masterframe_day36_2021_binded_day36$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_pile_followed_hi_form_2021 <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_piles <- length(which(pile_count_length_list_masterframe_day36_2021_binded_day36$rule_followed == "yes" & 
                                       pile_count_length_list_masterframe_day36_2021_binded_day36$runIDs == j))
  
  random_times_pile_followed_hi_form_2021 <- c(random_times_pile_followed_hi_form_2021, len_of_piles)
}

#turn vector into a dataframe for analysis
rand_pile_.frame_hi_form_2021 <- data.frame(random_times_pile_followed_hi_form_2021)

#plot time (the x intercept is the # of time a rule was used in the real data, found with 
#"pile_real <- length(which(pile_no_dupr$rule_followed == "yes"))
#pile_real.frame <- data.frame(pile_real))"
hi_form_pile_2021_count_plot_day36 <- ggplot(rand_pile_.frame_hi_form_2021, aes(x=random_times_pile_followed_hi_form_2021)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=5,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pile_count_length_list_2021_day36_real$count, colour = "red", size = 2)  + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of pile events", breaks=seq(0, 200, 50)) +  
  coord_cartesian(xlim = c(0, 200), ylim = c(0,.05)) +
  scale_y_continuous( breaks = seq(0, .05, .025)) 

hi_form_pile_2021_count_plot_day36

hi_form_pile_tally2021_day36 <-length(which(pile_count_length_list_2021_day36_real$count > rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021 ))



hi_form_ppile_tally2021_day36 <-   1 - hi_form_pile_tally2021_day36/length(rand_pile_.frame_hi_form_2021$random_times_pile_followed_hi_form_2021)

hi_form_ppile_tally2021_day36

#p = 
#calculate z score
#calculate mean and SD of permuted counts
mean_perm <- mean(random_times_pile_followed_hi_form_2021)
sd_perm <- sd(random_times_pile_followed_hi_form_2021)

# Calculate z-score
z_score_day36 <- (pile_count_length_list_2021_day36_real$count - mean_perm) / sd_perm


#create df to store important data for later stats and figures. to be binded later
hi_form_df_piece_day36 <- data.frame(group = integer(),
                                    bin = integer(),
                                    date =  character(),
                                    rule = character(),
                                    cmls = character(),
                                    p_cmls = integer(),
                                    z_score= integer())

 bin_for_store <- unique(pile_count_length_list_masterframe_day36_2021_binded_day36$bin)



#input data to be combined into df
new_row <- data.frame(
  group = 1,
  bin = bin_for_store,
  date = current_date,
  rule = "pile",
  cmls = "more",
  p_cmls = 0.008 ,
  z_score = z_score_day36,
  stringsAsFactors = FALSE
)
#bind row of data to df
hi_form_df_piece_day36 <- rbind(hi_form_df_piece_day36, new_row)

```


#combine days
```{r}
#bind all summary dfs to make group masterframe
hi_form_df_piece_master_piece_pile_2021_day_by_day <- rbind(hi_form_df_piece_day1,  hi_form_df_piece_day2,  hi_form_df_piece_day3,  hi_form_df_piece_day4,  hi_form_df_piece_day5,  hi_form_df_piece_day6,  hi_form_df_piece_day7,  hi_form_df_piece_day8,  hi_form_df_piece_day9,  hi_form_df_piece_day10,  hi_form_df_piece_day11,  hi_form_df_piece_day12,  hi_form_df_piece_day13,  hi_form_df_piece_day14,  hi_form_df_piece_day15,  hi_form_df_piece_day16,  hi_form_df_piece_day17,  hi_form_df_piece_day18,  hi_form_df_piece_day19,  hi_form_df_piece_day20,  hi_form_df_piece_day21,  hi_form_df_piece_day22,  hi_form_df_piece_day23,  hi_form_df_piece_day24,  hi_form_df_piece_day25,  hi_form_df_piece_day26,  hi_form_df_piece_day27,  hi_form_df_piece_day28,  hi_form_df_piece_day29,  hi_form_df_piece_day30,  hi_form_df_piece_day31,  hi_form_df_piece_day32,  hi_form_df_piece_day33,  hi_form_df_piece_day34,  hi_form_df_piece_day35,  hi_form_df_piece_day36)


#write.csv(hi_form_df_piece_master_piece_pile_2021_day_by_day,  here("ch2 data","hi_form_df_piece_master_piece_pile_2021_day_by_day.csv"), row.names = FALSE)
```






#plot rule following comp to random over days
```{r}
#load dfs
#load pile_count_length_sum_2021 by running code above
pile_count_length_list_masterframe_day_2021_binded <- read_csv(here("ch2 data",("pile_countmaster_day_level_hi_form_2021_custom_thresh.csv")))

hi_form_df_piece_master_piece_pile_2021_day_by_day <- read_csv(here("ch2 data",("hi_form_df_piece_master_piece_pile_2021_day_by_day.csv")))

head(pile_count_length_sum_2021)
head(pile_count_length_list_masterframe_day_2021_binded)
head(hi_form_df_piece_master_piece_pile_2021_day_by_day)

str(pile_count_length_list_masterframe_day_2021_binded)


#Canonical set of runs ( 1..1000 labels)
all_runs <- pile_count_length_list_masterframe_day_2021_binded %>%
  distinct(runIDs) %>% pull()

#Count "rule_followed == yes" per (date, runIDs); then complete to 1000 runs/day
rand_master_for_fig_pile <- pile_count_length_list_masterframe_day_2021_binded %>%
  filter(rule_followed == "yes") %>%
  select(date, runIDs) %>%
  group_by(date, runIDs) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(date) %>%
  tidyr::complete(runIDs = all_runs, fill = list(count = 0L)) %>%
  ungroup() %>%
  mutate(rule = "pile")





str(pile_count_length_sum_2021)
str(rand_master_for_fig_pile)
str(hi_form_df_piece_master_piece_pile_2021_day_by_day)

#write.csv(rand_master_for_fig_pile,  here("ch2 data","rand_master_for_fig_pile.csv"), row.names = FALSE)



#Observed rule follows
obs_daily <- pile_count_length_sum_2021 %>%
  select(date, count) %>%
  rename(obs_count = count)

#Randomized stats per day (FULL RANGE)
rand_stats <- rand_master_for_fig_pile %>%
  filter(rule == "pile") %>%
  group_by(date) %>%
  summarise(
    n_draws  = n(),
    min_rand = min(count, na.rm = TRUE),
    max_rand = max(count, na.rm = TRUE),
    mean_rand = mean(count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(date)

#Bottom z-scores
z_df <- hi_form_df_piece_master_piece_pile_2021_day_by_day %>%
  filter(rule == "pile") %>%
  select(date, cmls, z_score) %>%
  mutate(
    cmls = factor(
      cmls,
      levels = c("less", "same", "more"),
      labels = c("Less than random", "Same as random", "More than random")
    )
  )



#Global x-axis so both panels align
global_min <- min(c(obs_daily$date, rand_stats$date, z_df$date), na.rm = TRUE)
global_max <- max(c(obs_daily$date, rand_stats$date, z_df$date), na.rm = TRUE)
global_all_dates <- seq(global_min, global_max, by = "day")

#TOP tube + mean spine + observed, with global x
p_top <- ggplot() +
  geom_vline(xintercept = global_all_dates, linewidth = 0.25, alpha = 0.15) +
  #full-range tube
  geom_ribbon(
    data = rand_stats,
    aes(x = date, ymin = min_rand, ymax = max_rand),
    fill = "grey40", alpha = 0.18
  ) +
  #random mean rule use spine
  geom_line(
    data = rand_stats,
    aes(x = date, y = mean_rand, linetype = "Random mean"),
    linewidth = 0.9, color = "grey20"
  ) +
  # points for random mean
  geom_point(
    data = rand_stats,
    aes(x = date, y = mean_rand),
    color = "grey20", size = 1.8
  ) +
  #observed rule use line
  geom_line(
    data = obs_daily,
    aes(x = date, y = obs_count, color = "Observed"),
    linewidth = 1.1
  ) +
  #Points for observed rule use
  geom_point(
    data = obs_daily,
    aes(x = date, y = obs_count, color = "Observed"),
    size = 1.8
  ) +
  scale_linetype_manual(values = c("Random mean" = "dashed"), name = NULL) +
  scale_color_manual(values = c("Observed" = "#d62728"), name = NULL) +
  scale_x_date(
    breaks = global_all_dates,
    date_labels = "%b %d",
    expand = expansion(mult = c(0.001, 0.001))
  ) +
  coord_cartesian(xlim = c(global_min, global_max)) +
  labs(x = "Date", y = "Motif count",
       title = "Observed vs. randomized motif occurrences (daily)") +
  theme_minimal(base_size = 20) +
  theme(
    axis.text.x = element_text(angle = 90),
    axis.ticks.x = element_line(linewidth = 0.6),
    axis.ticks.length.x = grid::unit(3, "pt"),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(linewidth = 0.25),
    legend.position = "top"
  )
#BOTTOM z-scores, with the same global x
p_bottom <- ggplot(z_df, aes(x = date, y = z_score, fill = cmls)) +
  geom_vline(xintercept = global_all_dates, linewidth = 0.25, alpha = 0.15) +
  geom_col() +
  geom_hline(yintercept = 0, linewidth = 0.4) +
  scale_fill_manual(
    values = c("Less than random" = "#d62728", "Same as random" = "grey70", "More than random" = "#2ca02c"),
    name = NULL
  ) +
  scale_x_date(
    breaks = global_all_dates,
    date_labels = "%b %d",
    expand = expansion(mult = c(0.001, 0.001))
  ) +
  coord_cartesian(xlim = c(global_min, global_max)) +
  labs(x = "Date", y = "Z-score", title = "Daily difference (observed vs. random)") +
  theme_minimal(base_size = 20) +
  theme(
    axis.text.x = element_text(angle = 90),
    axis.ticks.x = element_line(linewidth = 0.6),
    axis.ticks.length.x = grid::unit(3, "pt"),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(linewidth = 0.25),
    legend.position = "top"
  )

#Stack 
(p_top / p_bottom) + patchwork::plot_layout(heights = c(2, 1))




#16 by 22?

```