---
title: "Additional Group 1-3 stats and plots"
author: "Xavier Francis"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#global thresh
```{r}
combined_data <- rbind(time_between_events1, time_between_events2, time_between_events3)

# Compute quantiles (95th, 75th, and 50th percentiles) of time_to_retal
quantiles <- quantile(combined_data$time_to_retal, probs = c(0.95, 0.75, 0.50), na.rm = TRUE)

# Display result
print(quantiles)
#95%     75%     50% 
#1363.15  413.75  138.00 
```

#Calculate Threshold Retaliation
```{r}
head(retal_length_ready_2021)
head(retal_length_ready2022G2)

#bind retal dfs and calculate mean retaliation speed
retal_length_comb <- rbind(retal_length_ready_2021, retal_length_ready2022G2)
retal_length_comb$time_to_retal <- as.numeric(retal_length_comb$time_to_retal)
retal_mean <- mean(retal_length_comb$time_to_retal) %>% 
round()

retal_mean
#269

 quantile(retal_length_comb$time_to_retal, probs = c(0.25, 0.5, 0.75, .95))
#25% 2
#50% 33
#75% 194 
#95% 1273    
```
#Calculate pileon threshold
```{r}
head(pile_length_ready_2021)
head(pile_length_ready_2022G1)
head(pile_length_ready_2022G2)


#bind retal dfs and calculate mean retaliation speed
pile_length_comb <- rbind(pile_length_ready_2021, pile_length_ready_2022G1,pile_length_ready_2022G2)
pile_length_comb$time_to_pile <- as.numeric(pile_length_comb$time_to_pile)
pile_mean <- mean(pile_length_comb$time_to_pile) %>% 
round()

pile_mean
#41
 quantile(pile_length_comb$time_to_pile, probs = c(0.25, 0.5, 0.75,.95))
#4 #12 #46 145 




surv <- c(seq(0,1000, by = 1))


surv.df <- data.frame(seconds = surv)

percentage_action <-c()




for (y in surv){ perc <-length(which(pile_length_comb$time_to_pile <= y))
perc_comb <- perc/length(pile_length_comb$time_to_pile)
percentage_action <- c(percentage_action, perc_comb)
 
}

percentage_action_ready <- data.frame( seconds = surv,
                                          percentage = percentage_action)

ggplot(data=percentage_action_ready, aes(x=seconds, y=percentage, group=1)) +
  geom_line()+
  geom_point() +
  scale_y_continuous(name = "percentage",
                     breaks = seq(0, 1, .1), 
                     limits = c(0, 1)) +
  scale_x_continuous(name = "length of all events (s)",
                     breaks = seq(0, 1000, 50), 
                     limits = c(0, 1000))



na.omit(pile_length_comb)

ggplot(data = pile_length_comb, aes(x=factor(0), y= time_to_pile)) + geom_boxplot(outlier.shape = NA) + 
stat_summary(fun=mean, color="darkred",geom="crossbar", show.legend=FALSE) + scale_x_discrete() +
    coord_cartesian(ylim = c(0, 400)) + scale_y_continuous(breaks = seq(0, 1000, 25)) 


#pile_75_thresh <- 
```
<!-- #relaxed pile (triple lead thresh) -->
<!-- ```{r} -->
<!-- head(pile_time_no_dupr_2021) -->
<!-- head(pile_time_no_dupr_2022G1) -->
<!-- head(pile_time_no_dupr_2022G2) -->

<!-- #bind retal dfs and calculate mean retaliation speed -->
<!-- pile_length_relax_comb <- rbind(pile_time_no_dupr_2021, pile_time_no_dupr_2022G1,pile_time_no_dupr_2022G2) -->
<!-- pile_length_relax_comb$time_to_pile <- as.numeric(pile_length_relax_comb$time_to_pile) -->

<!-- mean(pile_length_relax_comb$time_to_pile) -->
<!-- #79 -->


<!-- quantile(pile_length_relax_comb$time_to_pile, probs = c(0.25, 0.5, 0.75)) -->
<!-- # %25 8 -->
<!-- # median 35 -->
<!-- # %75 98 -->
<!-- ``` -->



#Calculate passalong threshold
```{r}
head(pass_length_ready)
head(pass_length_ready_2022G1)
head(pass_length_ready_2022G2)


pass_length_comb <- rbind(pass_length_ready,pass_length_ready_2022G1,pass_length_ready_2022G2)
pass_length_comb$time_to_pass <- as.numeric(pass_length_comb$time_to_pass)

pass_mean <- mean(pass_length_comb$time_to_pass) %>% 
round()

pass_mean
#395
 quantile(pass_length_comb$time_to_pass, probs = c(0.25, 0.5, 0.75, 0.95))
#25% 50% 75% 95%
 #43 173 494 1476
```
#Calculate opportunism threshold
```{r}
head(opp_length_ready_2021)
head(opp_length_ready_2022G1)
head(opp_length_ready_2022G2)

opp_length_comb <- rbind(opp_length_ready_2021,opp_length_ready_2022G1,opp_length_ready_2022G2)
opp_length_comb$time_to_opp <- as.numeric(opp_length_comb$time_to_opp)

opp_mean <- mean(opp_length_comb$time_to_opp) %>% 
round()

opp_mean
#676 mean
 quantile(opp_length_comb$time_to_opp, probs = c(0.25, 0.5, 0.75, 0.95))
#76   25%
#294   50%
#767  75%
#2444  95%
```


#Calculate punishment threshold
```{r}
head(pun_length_ready_2021)
head(pun_length_ready_2022G1)
head(pun_length_ready_2022G2)

pun_length_comb <- rbind(pun_length_ready_2021,pun_length_ready_2022G1,pun_length_ready_2022G2)
pun_length_comb$time_to_pun <- as.numeric(pun_length_comb$time_to_pun)

pun_mean <- mean(pun_length_comb$time_to_pun) %>% 
round()

pun_mean
#367
 quantile(pun_length_comb$time_to_pun, probs = c(0.25, 0.5, 0.75, 0.95))
#25% 41
#50% 149   
#75% 402
#95% 1404
```





#observed count of rule follows compared to random faceted plot
```{r}

#make all column names consistent
colnames(rand_retal_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G2)[1] <- 'count_of_rule_follows'







#bind all dfs containing the counts of rule follows from the randomized reference models. bind possible grouping factors
tdr_count_rand_comb <- rbind(
  cbind(rand_retal_.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(rand_pile_.frame, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(rand_pile_.frame_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(rand_pile_.frame_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(rand_pass_.frame, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G1, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G2, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(rand_opp_.frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G1, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G2, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(rand_pun_.frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G1, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G2, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)







#write.csv(tdr_count_rand_comb,  here("ch1 data","tdr_count_rand_comb.csv"), row.names = FALSE)







tdr_count_rand_comb <- read_csv(here("ch1 data","tdr_count_rand_comb.csv"))



tdr_count_rand_comb_mean <- tdr_count_rand_comb  %>% group_by(groupid, rule) %>% summarise(mean_value = mean(count_of_rule_follows, na.rm = TRUE))



#make all column names consistent
colnames(retal_real.frame)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2021)[1] <- 'count_of_rule_follows'
colnames(pile_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(pile_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_frame_2021)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G2_frame)[1] <- 'count_of_rule_follows'

# #bind all dfs containing the counts of rule follows from observed data. bind possible grouping factors
tdr_count_real_comb<- rbind(
  cbind(retal_real.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(retal_real.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(retal_real.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(pile_real.frame_2021, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(pile_real_2022G1_frame, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(pile_real_2022G2_frame, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(pass_along_real_frame_2021, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(pass_along_real_2022G1_frame, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(pass_along_real_2022G2_frame, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(opp_real_2021_frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(opp_real_2022G1_frame, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(opp_real_2022G2_frame, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(punishment_real_2021_frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(punishment_real_2022G1_frame, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(punishment_real_2022G2_frame, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

#write.csv(tdr_count_real_comb,  here("ch1 data","tdr_count_real_comb.csv"), row.names = FALSE)




tdr_count_real_comb <- read_csv(here("ch1 data","tdr_count_real_comb.csv"))


#tdr_count_rand_comb$rule <- factor(tdr_count_rand_comb$rule, levels = c("Retaliation", "Pile-on", "Pass-along", "Opportunism", "Punishment"))



#tdr_count_rand_comb$Group <- factor(tdr_count_rand_comb$Group, levels = c("Retalialtion Group 1", "Retaliation Group 2", "Retaliation Group 3", "Pile-on Group 1", "Pile-on Group 2", "Pile-on Group 3", "Pass-along Group 1", "Pass-along Group 2", "Pass-along Group 3", "Opportunism Group 1", "Opportunism Group 2", "Opportunism Group 3", "Punishment Group 1", "Punishment Group 2", "Punishment Group 3"))



#store each rule and group as a factor in a particular order I would like them displayed in my plot
#custom_order <- c("Retalialtion Group 1", "Retaliation Group 2", "Retaliation Group 3", "Pile-on Group 1", "Pile-on Group 2", "Pile-on Group 3", "Pass-along Group 1", "Pass-along Group 2", "Pass-along Group 3", "Opportunism Group 1", "Opportunism Group 2", "Opportunism Group 3", "Punishment Group 1", "Punishment Group 2", "Punishment Group 3")

# Convert group column to a factor with the desired order

#tdr_count_real_comb <- read.csv("tdr_count_real_comb", header=TRUE, stringsAsFactors=FALSE)



custom_order <- c("Retaliation","Pile-on",  "Pass-along", "Opportunism", "Punishment")
tdr_count_rand_comb$rule <- factor(tdr_count_rand_comb$rule, levels = custom_order)

tdr_count_real_comb$rule <- factor(tdr_count_real_comb$rule, levels = custom_order)




#store p values in particular order 
pval_df <- data.frame(
  Group = custom_order,    # Match facets with p values
  label = c("p = 0.174", "p = 0.16", "p = 0.036", "p = 0",
            "p = 0.122", "p = 0.042", "p = 0", "p = 0",
            "p = 0", "p = 0", "p = 0", "p = 0.492",
            "p = 0", "p = 0.028", "p = 0.003"),   
  # 90% of max count in each group to keep text inside
    x_pos = max(tdr_count_rand_comb$count_of_rule_follows) * 0.9,  
   # 90% of max density per group
    y_pos = max(density(tdr_count_rand_comb$count_of_rule_follows)$y) * 0.9 
)


bold_values_count <- c("p = 0.036", "p = 0","p = 0.042", "p = 0.028", "p = 0.003")

pval_df$font_weight <- ifelse(pval_df$label %in% bold_values_count, "bold", "plain")


# Ensure pval_df respect the custom order
pval_df$Group <- factor(pval_df$Group, levels = custom_order)






# Plot faceted set of violin plots that show how often each rule was followed compared to random chance. Observed rule follows vs randomized reference mdoels
rule_count_faceted_plot <- ggplot(tdr_count_rand_comb, aes(x = groupid, y = count_of_rule_follows)) +  # x is empty for density distribution
  geom_violin(fill = "grey", color = "black", alpha = 1) +  # Violin plot
  stat_summary(fun = mean, geom = "crossbar", width = 0.8, color = "black", size = 0.3)+
    geom_point(data = tdr_count_real_comb, aes(x = groupid, y = count_of_rule_follows), color = "red", size = 2, shape = 8, stroke = 2) +  # Add red star
  #geom_text(data = pval_df, aes(x = Inf, y = Inf, label = label, fontface = font_weight), 
           # hjust = 1.5, vjust = 1.5, color = "black", size = 7) +  # Custom text in top-right
  facet_wrap( ~ rule, scales = "free_x", ncol = 2, nrow = 5 ) +
  theme_minimal() + 
  theme(
    strip.text = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 25),  # Increase axis text size
    axis.title = element_text(size = 25, face = "bold"),
    axis.text.x = element_text(size = 25),
  axis.ticks.x = element_line(),
    panel.spacing = unit(1, "lines"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    )#,
    #axis.text.x = element_blank(),  # Remove x-axis labels (since x is empty)
    #axis.ticks.x = element_blank()   # Remove x-axis ticks
  ) +
  labs(
       x = "Density",
       y = "Count")

rule_count_faceted_plot

#ggsave("figure.pdf", plot = rule_count_faceted_plot, width = 10, height = 26, units = "in")






#18 by 20 landscape or
#18 by 22 portraite

#boxplot version
rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")

# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_rand_comb$rule))
#obtain each group and sort in order
groups <- sort(unique(tdr_count_rand_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,            
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "-", groupid)) |>
  pull(xcat)
tdr_rand2 <- tdr_count_rand_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )
#same thing for real rule follows
tdr_real2 <- tdr_count_real_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5
#plot
rule_boxplot <- ggplot(tdr_rand2, aes(x = xcat, y = count_of_rule_follows)) +
#separators behind boxes
geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.6) +
#randomized reference as gray boxplots
geom_boxplot(fill = "grey70", color = "black", width = 0.8, outlier.shape = 16) +
  #observed value as a red horizontal line across each category
geom_segment(
    data = tdr_real2,
    aes(x = as.numeric(xcat) - 0.35,
        xend = as.numeric(xcat) + 0.35,
        y = count_of_rule_follows,
        yend = count_of_rule_follows),
    inherit.aes = FALSE,
    linewidth = 1.2, color = "red"
  ) +
  labs(
    x = NULL,
    y = "Count of rule follows"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line()
  )


rule_boxplot





#compressed violin version

rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_rand_comb$rule))
groups <- sort(unique(tdr_count_rand_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_rand2 <- tdr_count_rand_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )
#same thing for observed rule follows
tdr_real2 <- tdr_count_real_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5

rule_violin <- ggplot(tdr_rand2, aes(x = xcat, y = count_of_rule_follows)) +
#separators behind violins
geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
#randomized reference as gray violins
geom_violin(fill = "grey70", color = "black", alpha = 1,
              width = 0.9, trim = FALSE, scale = "width") +
#mean bar across each violin
   stat_summary(fun = mean, geom = "crossbar", width = 0.6,
              color = "black", linewidth = 0.3)+
#observed value as a red horizontal line across each category
  geom_segment(
    data = tdr_real2,
    aes(x = as.numeric(xcat) - 0.35,
        xend = as.numeric(xcat) + 0.35,
        y = count_of_rule_follows,
        yend = count_of_rule_follows),
    inherit.aes = FALSE,
    linewidth = 1.2, color = "red"
  ) +
  labs(
    x = NULL,
    y = "Count of rule follows", 
    title = "Observed vs. randomized reference by rule and group (violin)"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_violin
#15 by 12 or 12 by 15
#did 15 by 12 but looks same honestly








#short bar plot to attach
head(tdr_count_rand_comb_mean)
head(tdr_count_real_comb)


tdr_count_both_comb <- tdr_count_rand_comb_mean %>%
  left_join(tdr_count_real_comb, by = c("groupid", "rule"))
#obtain the difference between observed count and the mean of the reference distribution
tdr_count_both_comb$dif <- tdr_count_both_comb$count_of_rule_follows - tdr_count_both_comb$mean_value



rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_both_comb$rule))
groups <- sort(unique(tdr_count_both_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_count_both_comb2 <- tdr_count_both_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5



rule_bar <- ggplot(tdr_count_both_comb2, aes(x = xcat, y = dif)) +
  #separators behind bars
  geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
  #bars from supplied y
  geom_col(fill = "grey60", color = "black") +
  labs(
    x = NULL,
    y = "Count of rule follows",
    title = "Observed vs. randomized reference by rule and group (bar)"
  ) +
  theme_minimal(base_size = 20) +
  scale_y_continuous(limits = c(-100, 100)) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 90, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_bar






library(patchwork)

# keep the top plot as-is visually, just hide its x labels/ticks so they don't repeat
rule_violin_top <- rule_violin +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin  = margin(5.5, 5.5, 0, 5.5)  # tighten gap to the bar plot
  )

#keep  bar plot exactly (fixed y-lims already set), just tweak bottom margin a tad
rule_bar_bottom <- rule_bar +
  theme(
    plot.margin = margin(0, 5.5, 5.5, 5.5)
  )

#stack them: violins on top (bigger), bars below (smaller)
combined <- rule_violin_top / rule_bar_bottom +
  plot_layout(heights = c(2, 1), guides = "keep")

combined
#12 by 10

```

#just one rule
```{r}
target_rule <- "Punishment"

# Filter and drop unused levels so xcat reindexes to 1..n
tdr_rand_one <- tdr_rand2 |>
  filter(rule == target_rule) |>
  droplevels()

tdr_real_one <- tdr_real2 |>
  filter(rule == target_rule) |>
  mutate(
    xcat = factor(xcat, levels = levels(tdr_rand_one$xcat)),  # align levels
    xpos = as.numeric(xcat)                                   # numeric for segment
  )

# No separators for one rule
sep_positions <- numeric(0)

rule_violin_one <- ggplot(tdr_rand_one, aes(x = xcat, y = count_of_rule_follows)) +
  geom_violin(fill = "grey70", color = "black", alpha = 1,
              width = 0.9, trim = FALSE, scale = "width") +
  stat_summary(fun = mean, geom = "crossbar", width = 0.6,
               color = "black", linewidth = 0.3) +
  geom_segment(
    data = tdr_real_one,
    aes(x = xpos - 0.35, xend = xpos + 0.35,
        y = count_of_rule_follows, yend = count_of_rule_follows),
    inherit.aes = FALSE,
    linewidth = 1.2, color = "red"
  ) +
  labs(
    x = NULL,
    y = "Count of rule follows",
    title = "Observed vs. randomized reference – Punishment only"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_violin_one


```





#summary heatmap
```{r}
summary_heatmap_file <- read_csv(here("ch1 data","Data_for_summary_heatmap.csv")) 

summary_heatmap_file

summary_heatmap_file$EMD <- as.numeric(summary_heatmap_file$EMD)


summary_heatmap_file <- summary_heatmap_file %>%
  mutate(Group_ID = case_when(
    Group_ID == "Group_1" ~ "Group 1",
    Group_ID == "Group_2" ~ "Group 2",
    Group_ID == "Group_3" ~ "Group 3",
    TRUE ~ Group_ID  # Keep other values unchanged
  ))


#duplicate mirror of rules to make sur eboth dirrections are there
summary_heatmap_file_rule_flipped <- summary_heatmap_file  %>%
  bind_rows(
    summary_heatmap_file %>%
      rename(Rule_1 = Rule_2, Rule_2 = Rule_1)
  )




#Ensure all rule combinations are present
rules <- unique(c(summary_heatmap_file$Rule_1, summary_heatmap_file$Rule_2))
all_combos <- expand.grid(Rule_1 = rules, Rule_2 = rules)



#join all combos to df    
summary_heatmap_file_final <- all_combos %>%
  left_join(summary_heatmap_file_rule_flipped, by = c("Rule_1", "Rule_2"))



#make 3 copies of each rule vs rule combo
just_same_combo <- subset(summary_heatmap_file_final, Rule_1 == Rule_2) %>%
  slice(rep(1:n(), each = 3))  

num_rules <- length(unique(just_same_combo$Rule_1))

#create group labels for each one of just the same vs same rules
  just_same_combo <- mutate(just_same_combo, Group_ID = rep(c("Group 1", "Group 2", "Group 3"), times = num_rules))      

  summary_heatmap_file_cut <- subset(summary_heatmap_file_final, Rule_1 != Rule_2)

summary_heatmap_file_final <- rbind(just_same_combo, summary_heatmap_file_cut)

#replace na p values with "-"
summary_heatmap_file_final <- summary_heatmap_file_final %>%
  mutate(p = ifelse(is.na(p), "-", p))

#replace empty font weight nas
summary_heatmap_file_final <- summary_heatmap_file_final %>%
  mutate(font_weight = ifelse(is.na(font_weight), "plain", font_weight))



# Define values to be bolded
bold_values <- c("<0.001", "0.006", "0.002", "0.012", "0.004")

summary_heatmap_file_final$font_weight <- ifelse(summary_heatmap_file_final$p %in% bold_values, "bold", "plain")



summary_heatmap_file_final$Rule_1 <- factor(summary_heatmap_file_final$Rule_1, levels = c("Retaliation", "Pile-on", "Pass-along", "Opportunism", "Punishment"))
summary_heatmap_file_final$Rule_2 <- factor(summary_heatmap_file_final$Rule_2, levels = c("Punishment", "Opportunism", "Pass-along","Pile-on", "Retaliation"))

library(dplyr)
library(ggplot2)

# Desired rule order
desired_levels <- c("Retaliation", "Pile-on", "Pass-along", "Opportunism", "Punishment")

# Bold these p-values
bold_values <- c("<0.001", "0.006", "0.002", "0.012", "0.004")

df <- summary_heatmap_file_final %>%
  mutate(
    # positional indices for triangles
    xi = match(as.character(Rule_2), desired_levels),  # Rule_2 (x-axis)
    yj = match(as.character(Rule_1), desired_levels),  # Rule_1 (y-axis)

    # factor ordering for display
    Rule_2 = factor(Rule_2, levels = desired_levels),
    Rule_1 = factor(Rule_1, levels = rev(desired_levels)),

    # labels: upper triangle = EMD (1 decimal), lower triangle = p, diagonal blank
    label_value = case_when(
      yj < xi ~ ifelse(is.na(EMD), "", format(round(EMD, 1), nsmall = 1)),
      yj > xi ~ p,
      TRUE ~ ""
    ),

    # bold specific p-values
    fontface = case_when(
      yj > xi & p %in% bold_values ~ "bold",
      TRUE ~ "plain"
    )
  )

summary_heatmap <- ggplot(df, aes(x = Rule_2, y = Rule_1, fill = EMD)) +
  geom_tile(color = "grey30", linewidth = 0.25) +
  scale_fill_distiller(
    palette = "Spectral", direction = -1, na.value = "grey90", name = "EMD",
    guide = guide_colorbar(
      direction = "horizontal",
      barwidth  = unit(3, "cm"),
      barheight = unit(0.25, "cm"),
      title.position = "top",
      title.hjust = 0.5
    )
  ) +
  coord_fixed() +
  geom_text(aes(label = label_value, fontface = fontface),
            color = "black", size = 5.3, lineheight = 0.9) +  # size 5 text
  facet_wrap(~ Group_ID, ncol = 3, nrow = 1) +
  scale_x_discrete(limits = desired_levels) +
  scale_y_discrete(limits = rev(desired_levels)) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position   = "bottom",
    legend.title      = element_text(size = 9),
    legend.text       = element_text(size = 8),
    legend.margin     = margin(1, 0, 1, 0),
    legend.box.margin = margin(-5, 0, 0, 0),
    strip.text        = element_text(size = 13, face = "bold"),
    axis.text.x       = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 9),
    axis.text.y       = element_text(size = 9),
    axis.title.x      = element_text(size = 11, face = "bold", margin = margin(t = 8)),
    axis.title.y      = element_text(size = 11, face = "bold", margin = margin(r = 8)),
    panel.spacing     = unit(0.28, "cm"),
    panel.border      = element_rect(color = "black", fill = NA, linewidth = 0.8),
    plot.margin       = margin(8, 8, 8, 8)
  ) +
  labs(x = "Rule 2", y = "Rule 1")

summary_heatmap



#12 by 10











#install.packages("ggtext")
#library(ggtext)  # For rendering markdown-styled text

# Plot
summary_heatmap <- ggplot(summary_heatmap_file_final, aes(x = Rule_1, y = Rule_2, fill = EMD )) + 
  geom_tile(colour="grey20",size=0.25, linewidth = 0.3) +
      scale_fill_distiller(palette = "Spectral", direction = -1) +  # Spectral palette
  coord_fixed() +  # Ensures square cells
  geom_text(aes(label = p,  fontface = font_weight), color = "black", size = 12) +  # Add text labels
  facet_wrap(~ Group_ID,  ncol = 3, nrow = 1) +# Free x-axis but consistent bar widths
  theme_minimal()+
  theme(
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    legend.key.size = unit(0.1, "cm"),
    strip.text  = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 14, vjust=0.6),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 14, face = "bold",  margin = margin(t = 15)),  # X-axis title
    axis.title.y = element_text(size = 14, face = "bold",  margin = margin(r = 15)),
    panel.spacing = unit(.3, "cm"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    )
  )  +
  labs(
    x = "Rule 1",  # Add your custom x-axis title here
    y = "Rule 2"   # Add your custom y-axis title here
  )

summary_heatmap
#20 by 38 default


ggsave("figure.pdf", plot = summary_heatmap, width = 38, height = 20, units = "in")


getwd()

str(summary_heatmap_file_final)



#gridextra version with different scales

#G1
summary_heatmap_df_G1 <- subset(summary_heatmap_file_final, Group_ID == "Group 1")
# Plot G1
summary_heatmap_G1 <- ggplot(summary_heatmap_df_G1, aes(x = Rule_1, y = Rule_2, fill = EMD )) + 
  geom_tile(colour="white",size=0.25) +
  coord_fixed() +  # Ensures square cells
  geom_text(aes(label = p), color = "black", size = 10) +  # Add text labels
  scale_fill_gradient(low="white", high="red", na.value = "white") +
  theme_minimal()  +
  ggtitle("Group 1") +  # Add title
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  
    strip.text = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 14, vjust=0.6),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 14,  margin = margin(t = 15)),  # X-axis title
    axis.title.y = element_text(size = 14,   margin = margin(r = 15)),
    plot.background = element_rect(color = "black", linewidth = 3),
    plot.margin = margin(17, 17, 17, 17),  # Increase the space around the plot (top, right, bottom, left)
    panel.spacing = unit(.3, "cm"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    ),
  )  +
  labs(
    x = "Rule 1",  # Add your custom x-axis title here
    y = "Rule 2"   # Add your custom y-axis title here
  )

summary_heatmap_G1



#G2
summary_heatmap_df_G2 <- subset(summary_heatmap_file_final, Group_ID == "Group 2")
# Plot G2
summary_heatmap_G2 <- ggplot(summary_heatmap_df_G2, aes(x = Rule_1, y = Rule_2, fill = EMD )) + 
  geom_tile(colour="white",size=0.25) +
  coord_fixed() +  # Ensures square cells
  geom_text(aes(label = p), color = "black", size = 10) +  # Add text labels
  scale_fill_gradient(low="white", high="red", na.value = "white") +
  theme_minimal()  +
  ggtitle("Group 2") +  # Add title
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  
    strip.text = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 14, vjust=0.6),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 14,  margin = margin(t = 15)),  # X-axis title
    axis.title.y = element_text(size = 14,   margin = margin(r = 15)),
    plot.background = element_rect(color = "black", linewidth = 3),
    plot.margin = margin(17, 17, 17, 17),  # Increase the space around the plot (top, right, bottom, left)
    panel.spacing = unit(.3, "cm"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    ),
  )  +
  labs(
    x = "Rule 1",  # Add your custom x-axis title here
    y = "Rule 2"   # Add your custom y-axis title here
  )

summary_heatmap_G2


#G3
summary_heatmap_df_G3 <- subset(summary_heatmap_file_final, Group_ID == "Group 3")
# Plot G3
summary_heatmap_G3 <- ggplot(summary_heatmap_df_G3, aes(x = Rule_1, y = Rule_2, fill = EMD )) + 
  geom_tile(colour="white",size=0.25) +
  coord_fixed() +  # Ensures square cells
  geom_text(aes(label = p), color = "black", size = 10) +  # Add text labels
  scale_fill_gradient(low="white", high="red", na.value = "white") +
  theme_minimal()  +
  ggtitle("Group 3") +  # Add title
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  
    strip.text = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 14, vjust=0.6),
    axis.text.y = element_text(size = 14), 
    axis.title.x = element_text(size = 14,  margin = margin(t = 15)),  # X-axis title
    axis.title.y = element_text(size = 14,   margin = margin(r = 15)),
    plot.background = element_rect(color = "black", linewidth = 3),
    plot.margin = margin(17, 17, 17, 17),  # Increase the space around the plot (top, right, bottom, left)
    panel.spacing = unit(.3, "cm"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    ),
  )  +
  labs(
    x = "Rule 1",  # Add your custom x-axis title here
    y = "Rule 2"   # Add your custom y-axis title here
  )

summary_heatmap_G3

library(gridExtra)
library(grid)

spacer <- nullGrob()  # Empty space for spacing

grid.arrange(summary_heatmap_G1, spacer, summary_heatmap_G2, spacer, summary_heatmap_G3, ncol = 1, heights = c(1, 0.1, 1, 0.1, 1))

#20 by 38 default
#ignore for now


```










#individ faceted heatmap of observed rule follows
```{r}





#bind individ dataframes into one summary_heatmap_file
indiv_combined <- rbind(
  cbind(summary.retal.individ_2021, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(summary.retal.individ_2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(summary.retal.individ_2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(summary.pile.individ_2021, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(summary.pile.individ_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(summary.pile.individ_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(summary.pass.individ_2021, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(summary.pass.individ_2022G1, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(summary.pass.individ_2022G2, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(summary.opp.individ_2021, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(summary.opp.individ_2022G1, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(summary.opp.individ_2022G2, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(summary.pun.individ_2021, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(summary.pun.individ_2022G1, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(summary.pun.individ_2022G2, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

unique(indiv_combined$actor_2)


#write.csv(indiv_combined,  here("ch1 data","indiv_combined.csv"), row.names = FALSE)


#indiv_combined <- read_csv(here("ch1 data","indiv_combined.csv"))






# Create a consistent category scale across facets
indiv_combined$actor_2 <- factor(
  indiv_combined$actor_2,
  levels = c("GGG", "OBB", "OGO", "OOP", "OPP", "PGG", "POP", "PPO", "BBB", "BOB", "GGO", "GOO", "GOP", "GPG", "OOO", "PBB", "PBO", "POO", "PPP", "GGP", "GPO", "OPO", "BBO", "PBP") # Include all categories across all groups
)

#indiv_combined$Group <- factor(indiv_combined$Group, levels = c("Retalialtion Group 1", "Retaliation Group 2", "Retaliation Group 3", "Pile-on Group 1", "Pile-on Group 2", "Pile-on Group 3", "Pass-along Group 1", "Pass-along Group 2", "Pass-along Group 3", "Opportunism Group 1", "Opportunism Group 2", "Opportunism Group 3", "Punishment Group 1", "Punishment Group 2", "Punishment Group 3"))


#store each group as a factor in a particular order I would like them displayed in my plot
indiv_combined$groupid <- factor(indiv_combined$groupid, levels = c("Group 1", "Group 2", "Group 3"))

#store each rule as a factor in a particular order I would like them displayed in my plot
indiv_combined$rule <- factor(indiv_combined$rule, levels = c("Retaliation", "Pile-on", "Pass-along", "Opportunism", "Punishment"))


# 
# 
# #store each rule and group as a factor in a particular order I would like them displayed in my plot
# custom_order <- c("Retalialtion Group 1", "Retaliation Group 2", "Retaliation Group 3", "Pile-on Group 1", "Pile-on Group 2", "Pile-on Group 3", "Pass-along Group 1", "Pass-along Group 2", "Pass-along Group 3", "Opportunism Group 1", "Opportunism Group 2", "Opportunism Group 3", "Punishment Group 1", "Punishment Group 2", "Punishment Group 3")
# 
# 
# 
# 
# # Convert group column to a factor with the desired order
# indiv_combined$Group <- factor(indiv_combined$Group, levels = custom_order)
# 
# #store p values in particular order 
# pval_individ_df <- data.frame(
#   Group = custom_order,    # Match facet group names
#   label = c("p = 0.174", "p = 0.16", "p = 0.036", "p = 0",
#             "p = 0.122", "p = 0.042", "p = 0", "p = 0",
#             "p = 0", "p = 0", "p = 0", "p = 0.492",
#             "p = 0", "p = 0.028", "p = 0.003"),  
#   # 90% of max count in each group to keep text inside
#     x_pos = max(indiv_combined$count) * 0.9,  # 90% of max count in each group to keep text inside
#   # 90% of max density per group
#     y_pos = max(density(indiv_combined$count)$y) * 0.9  # 90% of max density per group
# )
# 


bold_values_individ_count <- c()

#pval_df$font_weight <- ifelse(pval_df$label %in% bold_values_individ_count, "bold", "plain")

# Ensure pval_time_df respect the custom order
#pval_individ_df$Group <- factor(pval_individ_df$Group, levels = custom_order)





# Plot faceted set of heatmaps that show how often each individual used each rule
TDR_individ_heatmap <- ggplot(indiv_combined, aes(x = rule, y = actor_2, fill = count )) + 
  geom_tile(colour="white",size=0.25) +
  scale_fill_gradient(low="white", high="red", na.value = "white") +
  #geom_text(data = pval_individ_df, aes(x = Inf, y = Inf, label = label), 
            #inherit.aes = FALSE,
           # hjust = 1.5, vjust = 1.5, color = "black", size = 10, fontface = "bold") +
  geom_text(aes(label = count), color = "black", size = 12) +  # Add text labels
  facet_wrap(~ groupid, scales = "free", ncol = 1, nrow = 3) +# Free x-axis but consistent bar widths
  theme_minimal()  +
  theme(legend.text = element_text(size = 30),
    legend.title = element_text(size = 32),
    legend.key.size = unit(1.5, "cm"),
    strip.text  = element_text(size = 45, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 35, vjust=0.5),
    axis.text.y = element_text(size = 30), 
    axis.title.x = element_text(size = 50,   margin = margin(t = 15)),  # X-axis title
    axis.title.y = element_text(size = 50,   margin = margin(r = 15)),
     panel.spacing = unit(.3, "cm"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    )
  )  +
  labs(
    x = "Time Dependent Rule",  # Add custom x-axis title here
    y = "Parakeet ID"   # Add custom y-axis title here
  )

TDR_individ_heatmap

#ggsave("figure.pdf", plot = TDR_individ_heatmap, width = 18, height = 37, units = "in")


library(viridis)

#gridextra version with different scales

#G1
indiv_combined_G1 <- subset(indiv_combined, groupid == "Group 1")

#rescale within a group; if all values are the same, map to 0.5 so tiles aren't white
rescale_within <- function(x) {
  if (all(is.na(x))) return(x)
  rng <- range(x, na.rm = TRUE)
  if (rng[1] == rng[2]) rep(0.5, length(x)) else scales::rescale(x, to = c(0, 1), from = rng)
}

#Create a per rule scaled value
indiv_combined_G1 <- indiv_combined_G1 %>%
  group_by(rule) %>%
  mutate(scaled_count = rescale_within(count)) %>%
  ungroup()

# Plot G1
library(scales)

TDR_individ_heatmap_G1 <- ggplot(indiv_combined_G1, aes(x = rule, y = actor_2, fill = scaled_count)) + 
  geom_tile(colour = "grey20", size = 0.2, linewidth = 0.3) +
  geom_text(aes(label = count), color = "black", size = 12) +
  scale_fill_distiller(
    palette = "Spectral",
    direction = -1,
    limits = c(0, 1),
    name = "Within-motif intensity",
    labels = label_percent(accuracy = 1)
    # na.value = "white"  # uncomment if you want explicit NA tiles to be white
  ) +
  theme_minimal() +
  ggtitle("Group 1") +
  theme(
    plot.title = element_text(size = 45, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 45, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 30, vjust = 0.5),
    axis.text.y = element_text(size = 30),
    axis.title.x = element_text(size = 40, margin = margin(t = 15)),
    axis.title.y = element_text(size = 40, margin = margin(r = 15)),
    plot.background = element_rect(color = "black", linewidth = 3),
    plot.margin = margin(17, 17, 17, 17),
    panel.spacing = unit(.3, "cm"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  ) +
  labs(
    x = "Time Dependent Rule",
    y = "Parakeet ID"
  )

TDR_individ_heatmap_G1

#ggsave("ch1_indiv_combined_2021.pdf", plot = TDR_individ_heatmap_G1, width = 18, height = 27, units = "in")
















#G2
indiv_combined_G2 <- subset(indiv_combined, groupid == "Group 2")

#rescale within a group; if all values are the same, map to 0.5 so tiles aren't white
rescale_within <- function(x) {
  if (all(is.na(x))) return(x)
  rng <- range(x, na.rm = TRUE)
  if (rng[1] == rng[2]) rep(0.5, length(x)) else scales::rescale(x, to = c(0, 1), from = rng)
}

#Create a per rule scaled value
indiv_combined_G2 <- indiv_combined_G2 %>%
  group_by(rule) %>%
  mutate(scaled_count = rescale_within(count)) %>%
  ungroup()

# Plot G2

library(scales)

TDR_individ_heatmap_G2 <- ggplot(indiv_combined_G2, aes(x = rule, y = actor_2, fill = scaled_count)) + 
  geom_tile(colour = "grey20", size = 0.2, linewidth = 0.3) +
  geom_text(aes(label = count), color = "black", size = 12) +
  scale_fill_distiller(
    palette = "Spectral",
    direction = -1,
    limits = c(0, 1),
    name = "Within-motif intensity",
    labels = label_percent(accuracy = 1)
    # na.value = "white"  # uncomment if you want explicit NA tiles to be white
  ) +
  theme_minimal() +
  ggtitle("Group 2") +
  theme(
    plot.title = element_text(size = 45, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 45, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 30, vjust = 0.5),
    axis.text.y = element_text(size = 30),
    axis.title.x = element_text(size = 40, margin = margin(t = 15)),
    axis.title.y = element_text(size = 40, margin = margin(r = 15)),
    plot.background = element_rect(color = "black", linewidth = 3),
    plot.margin = margin(17, 17, 17, 17),
    panel.spacing = unit(.3, "cm"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  ) +
  labs(
    x = "Time Dependent Rule",
    y = "Parakeet ID"
  )

TDR_individ_heatmap_G2
#ggsave("ch1_indiv_combined_2022G1.pdf", plot = TDR_individ_heatmap_G2, width = 18, height = 15, units = "in")



#G3
indiv_combined_G3 <- subset(indiv_combined, groupid == "Group 3")


#rescale within a group; if all values are the same, map to 0.5 so tiles aren't white
rescale_within <- function(x) {
  if (all(is.na(x))) return(x)
  rng <- range(x, na.rm = TRUE)
  if (rng[1] == rng[2]) rep(0.5, length(x)) else scales::rescale(x, to = c(0, 1), from = rng)
}

#Create a per rule scaled value
indiv_combined_G3 <- indiv_combined_G3 %>%
  group_by(rule) %>%
  mutate(scaled_count = rescale_within(count)) %>%
  ungroup()

# Plot G3

library(scales)

TDR_individ_heatmap_G3 <- ggplot(indiv_combined_G3, aes(x = rule, y = actor_2, fill = scaled_count)) + 
  geom_tile(colour = "grey20", size = 0.2, linewidth = 0.3) +
  geom_text(aes(label = count), color = "black", size = 12) +
  scale_fill_distiller(
    palette = "Spectral",
    direction = -1,
    limits = c(0, 1),
    name = "Within-motif intensity",
    labels = label_percent(accuracy = 1)
    # na.value = "white"  # uncomment if you want explicit NA tiles to be white
  ) +
  theme_minimal() +
  ggtitle("Group 2") +
  theme(
    plot.title = element_text(size = 45, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 45, face = "bold"),
    axis.text.x = element_text(angle = 90, size = 30, vjust = 0.5),
    axis.text.y = element_text(size = 30),
    axis.title.x = element_text(size = 40, margin = margin(t = 15)),
    axis.title.y = element_text(size = 40, margin = margin(r = 15)),
    plot.background = element_rect(color = "black", linewidth = 3),
    plot.margin = margin(17, 17, 17, 17),
    panel.spacing = unit(.3, "cm"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  ) +
  labs(
    x = "Time Dependent Rule",
    y = "Parakeet ID"
  )

TDR_individ_heatmap_G3
#ggsave("ch1_indiv_combined_2022G2.pdf", plot = TDR_individ_heatmap_G3, width = 18, height = 15, units = "in")




library(gridExtra)
library(grid)
# Empty space for spacing
spacer <- nullGrob()  
 grid.arrange(TDR_individ_heatmap_G1, spacer, TDR_individ_heatmap_G2, spacer, TDR_individ_heatmap_G3, ncol = 1, heights = c(1, 0.1, 1, 0.1, 1))
#18 38 portrait
 
 
 library(dplyr)
library(ggplot2)
library(patchwork)
library(scales)

#Helper: rescale within rule
rescale_within <- function(x) {
  if (all(is.na(x))) return(x)
  rng <- range(x, na.rm = TRUE)
  if (rng[1] == rng[2]) rep(0.5, length(x)) else scales::rescale(x, to = c(0, 1), from = rng)
}

#Universal plotting helper
make_heatmap <- function(df_group, group_title) {
  dfp <- df_group %>%
    group_by(rule) %>%
    mutate(scaled_count = rescale_within(count)) %>%
    ungroup()

  ggplot(dfp, aes(x = rule, y = actor_2, fill = scaled_count)) +
    geom_tile(colour = "grey20", linewidth = 0.25) +
    geom_text(aes(label = count), color = "black", size = 4.5) +
    scale_fill_distiller(
      palette = "Spectral", direction = -1,
      limits = c(0, 1),
      name = "Within-motif intensity",
      labels = label_percent(accuracy = 1)
    ) +
    coord_fixed(ratio = 0.5) +  #wider cells (shorter vertically)
    ggtitle(group_title) +
    theme_minimal(base_size = 11) +
    theme(
      plot.title   = element_text(size = 18, face = "bold", hjust = 0.5),
      axis.text.x  = element_text(angle = 90, size = 9, vjust = 0.5, hjust = 1),
      axis.text.y  = element_text(size = 9),
      axis.title.x = element_text(size = 10, margin = margin(t = 6)),
      axis.title.y = element_text(size = 10, margin = margin(r = 6)),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.6),
      plot.margin  = margin(5, 5, 5, 5),
      legend.position = "bottom",
      legend.title = element_text(size = 8),
      legend.text  = element_text(size = 7)
    ) +
    labs(x = "Time-Dependent Rule", y = "Parakeet ID")
}

#Make plots for each group 
indiv_combined_G1 <- subset(indiv_combined, groupid == "Group 1")
indiv_combined_G2 <- subset(indiv_combined, groupid == "Group 2")
indiv_combined_G3 <- subset(indiv_combined, groupid == "Group 3")

p1 <- make_heatmap(indiv_combined_G1, "Group 1")
p2 <- make_heatmap(indiv_combined_G2, "Group 2")
p3 <- make_heatmap(indiv_combined_G3, "Group 3")

# Group 1 (tall) on left, Groups 2 + 3 stacked on right
final_plot <-
  (p1 | (p2 / p3)) +
  plot_layout(widths = c(1.25, 1), guides = "collect") +
  plot_annotation(theme = theme(legend.position = "bottom"))

final_plot
#10 by 12

 
```
#bar graph version individuals
```{r}
library(dplyr)
library(ggplot2)
library(forcats)

# Make sure the rule order is fixed
rule_order <- c("Retaliation", "Pile-on", "Pass-along", "Opportunism", "Punishment")
indiv_combined$rule <- factor(indiv_combined$rule, levels = rule_order)

# Summarize counts by group, rule, and parakeet
df_bar <- indiv_combined %>%
  group_by(groupid, rule, actor_2) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop") %>%
  mutate(rule = factor(rule, levels = rule_order))

# Helper function to plot one group
make_group_barplot <- function(df, group_name) {
  ggplot(df, aes(x = actor_2, y = count)) +
    geom_col(width = 0.7) +
    facet_wrap(~ rule, nrow = 1, scales = "free_x") +
    labs(
      title = paste0(group_name),
      x = "Parakeet ID",
      y = "Number of rule follows"
      
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
      strip.text = element_text(face = "bold", size = 11),
      panel.grid.major.x = element_blank(),
      legend.position = "none"
    )
}

# Subset data for each group
df_G1 <- subset(df_bar, groupid == "Group 1")
df_G2 <- subset(df_bar, groupid == "Group 2")
df_G3 <- subset(df_bar, groupid == "Group 3")

# Generate plots
p_G1 <- make_group_barplot(df_G1, "Group 1")
p_G2 <- make_group_barplot(df_G2, "Group 2")
p_G3 <- make_group_barplot(df_G3, "Group 3")

# Show them
p_G1
p_G2
p_G3

```

#just 1 rule
```{r}
library(dplyr)
library(ggplot2)

# Filter to only Retaliation
df_retaliation <- indiv_combined %>%
  filter(rule == "Retaliation") %>%
  group_by(groupid, actor_2) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# Plot: one facet per group
retaliation_plot <- ggplot(df_retaliation, aes(x = actor_2, y = count)) +
  geom_col(width = 0.7) +
  facet_wrap(~ groupid, nrow = 1, scales = "free_x") +
  labs(
    x = "Parakeet ID",
    y = "Number of rule follows",
    fill = "Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
    strip.text = element_text(face = "bold", size = 11),
    panel.grid.major.x = element_blank(),
    legend.position = "none"  # each facet already labeled by group
  )

retaliation_plot

```
#try
```{r}
library(dplyr)
library(ggplot2)
library(patchwork)

# Data: Retaliation only, summed by parakeet within group
df_retaliation <- indiv_combined %>%
  filter(rule == "Punishment") %>%
  group_by(groupid, actor_2) %>%
  summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

# Optional: share the same y-axis across all three plots for comparability
y_max <- df_retaliation$count |> max(na.rm = TRUE)

make_plot <- function(dfg, title) {
  ggplot(dfg, aes(x = actor_2, y = count)) +
    geom_col(width = 0.7) +
    scale_y_continuous(limits = c(0, y_max), expand = expansion(mult = c(0, 0.05))) +
    labs( x = "Parakeet ID", y = "Number of rule follows") +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
      panel.grid.major.x = element_blank(),
      plot.margin = margin(5, 8, 5, 8)
    )
}

p1 <- make_plot(filter(df_retaliation, groupid == "Group 1"), "Group 1")
p2 <- make_plot(filter(df_retaliation, groupid == "Group 2"), "Group 2")
p3 <- make_plot(filter(df_retaliation, groupid == "Group 3"), "Group 3")

# Layout: Group 1 on top; Groups 2 & 3 side by side on bottom
final_retaliation <-
  p1 / (p2 | p3) +
  plot_layout(heights = c(1.2, 1)) +   # make top row taller
  plot_annotation(
    title = "Retaliation Rule Follows by Parakeet",
    theme = theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5))
  )

final_retaliation

```

#richness, eveness, diversity calculations

```{r}
library(vegan)

indiv_combined <- read_csv(here("ch1 data","indiv_combined.csv"))


head(indiv_combined)

#store each rule as a factor in a particular order so when binded back on they stay in order of simple to complex
indiv_combined$rule <- factor(indiv_combined$rule, levels = c("Retaliation", "Pile-on", "Pass-along", "Opportunism", "Punishment"))

actor_col <- if ("actor" %in% names(indiv_combined)) "actor" else "actor_2"

#counts of rule follows per bird
by_actor <- indiv_combined %>%
  group_by(groupid, rule, actor = .data[[actor_col]]) %>%
  summarise(n = sum(count, na.rm = TRUE), .groups = "drop")

#metrics per group × rule 
metrics_all_groups <- by_actor %>%
  group_by(groupid, rule) %>%
  summarise(
    Richness  = sum(n > 0, na.rm = TRUE),                    
    Diversity = if (sum(n, na.rm = TRUE) == 0) 0 else vegan::diversity(n, index = "shannon"),
    Evenness  = dplyr::if_else(Richness > 1, Diversity / log(Richness), NA_real_),
    .groups = "drop"
  )


#group size denominator = unique individuals in each group across both roles so proportional richness is comparable across groups)
group_sizes <- indiv_combined %>%
  group_by(groupid) %>%
  summarise(N = n_distinct(actor_2, na.rm = TRUE), .groups = "drop")


#calculate proportional richness 
metrics_all_groups <- metrics_all_groups %>%
  left_join(group_sizes, by = "groupid") %>%
  mutate(
    Proportional_Richness = if_else(
      rule %in% c("Retaliation","Pile-on"),
      Richness / N,
      Richness / pmax(N - 1, 1)  # avoid divide-by-zero
    )
  ) %>%
  arrange(groupid, rule)

metrics_all_groups




diversity_summary_all <- metrics_all_groups


diversity_summary_2021 <- subset(diversity_summary_all, groupid == "Group 1")

# 
# 

#group 1
#Convert to long format
div_long_2021 <- diversity_summary_2021 %>%
  pivot_longer(cols = c(Proportional_Richness, Diversity, Evenness),
               names_to = "Metric",
               values_to = "Value")


div_long_2021$Metric[div_long_2021$Metric == "Proportional_Richness"] <- "Proportional Richness"
div_long_2021$Metric[div_long_2021$Metric == "Diversity"] <- "Shannon's Diversity"
div_long_2021$Metric[div_long_2021$Metric == "Evenness"] <- "Pilou's Evenness"




div_long_2021$Metric <- factor(div_long_2021$Metric, levels = c("Proportional Richness", "Shannon's Diversity", "Pilou's Evenness"))
# Plot
div_long_2021_plot <- ggplot(div_long_2021, aes(x = rule, y = Value, fill = Metric)) +
  geom_col(show.legend = FALSE, width = 0.7) +
  facet_wrap(~ Metric, scales = "free_y") +
  scale_fill_viridis_d(option = "D") +
  
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(, angle = 90, size = 12, hjust = 1),
    strip.text = element_text(face = "bold", size = 13),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5)
  ) +
  labs(
    title = "Diversity Metrics by Rule: Group 1",
    x = NULL,
    y = NULL
  )

div_long_2021_plot

#ggsave("figure.pdf", plot = div_long_2021_plot, width = 10, height = 6, units = "in")



#group 2
#Convert to long format
diversity_summary_2022G1 <- subset(diversity_summary_all, groupid == "Group 2")

div_long_2022G1 <- diversity_summary_2022G1 %>%
  pivot_longer(cols = c(Proportional_Richness, Diversity, Evenness),
               names_to = "Metric",
               values_to = "Value")


div_long_2022G1$Metric[div_long_2022G1$Metric == "Proportional_Richness"] <- "Proportional Richness"
div_long_2022G1$Metric[div_long_2022G1$Metric == "Diversity"] <- "Shannon's Diversity"
div_long_2022G1$Metric[div_long_2022G1$Metric == "Evenness"] <- "Pilou's Evenness"

div_long_2022G1$Metric <- factor(div_long_2022G1$Metric, levels = c("Proportional Richness", "Shannon's Diversity", "Pilou's Evenness"))


# Plot
div_long_2022G1_plot <- ggplot(div_long_2022G1, aes(x = rule, y = Value, fill = Metric)) +
  geom_col(show.legend = FALSE, width = 0.7) +
  facet_wrap(~ Metric, scales = "free_y") +
  scale_fill_viridis_d(option = "D") +
  
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(, angle = 90, size = 12, hjust = 1),
    strip.text = element_text(face = "bold", size = 13),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5)
  ) +
  labs(
    title = "Diversity Metrics by Rule: Group 2",
    x = NULL,
    y = NULL
  )


div_long_2022G1_plot
#ggsave("figure.pdf", plot = div_long_2022G1_plot, width = 10, height = 6, units = "in")











#group 3
#Convert to long format
diversity_summary_2022G2 <- subset(diversity_summary_all, groupid == "Group 3")

div_long_2022G2 <- diversity_summary_2022G2 %>%
  pivot_longer(cols = c(Proportional_Richness, Diversity, Evenness),
               names_to = "Metric",
               values_to = "Value")


div_long_2022G2$Metric[div_long_2022G2$Metric == "Proportional_Richness"] <- "Proportional Richness"
div_long_2022G2$Metric[div_long_2022G2$Metric == "Diversity"] <- "Shannon's Diversity"
div_long_2022G2$Metric[div_long_2022G2$Metric == "Evenness"] <- "Pilou's Evenness"

div_long_2022G2$Metric <- factor(div_long_2022G2$Metric, levels = c("Proportional Richness", "Shannon's Diversity", "Pilou's Evenness"))


# Plot
div_long_2022G2_plot <- ggplot(div_long_2022G2, aes(x = rule, y = Value, fill = Metric)) +
  geom_col(show.legend = FALSE, width = 0.7) +
  facet_wrap(~ Metric, scales = "free_y") +
  scale_fill_viridis_d(option = "D") +
  
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(, angle = 90, size = 12, hjust = 1),
    strip.text = element_text(face = "bold", size = 13),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5)
  ) +
  labs(
    title = "Diversity Metrics by Rule: Group 3",
    x = NULL,
    y = NULL
  )
#ggsave("figure.pdf", plot = div_long_2022G2_plot, width = 10, height = 6, units = "in")



div_long_all <- rbind(div_long_2021, div_long_2022G1, div_long_2022G2)



div_long_all$Metric <- factor(div_long_all$Metric, levels = c("Proportional Richness", "Shannon's Diversity", "Pilou's Evenness"))
div_long_all$groupid <- factor(div_long_all$groupid, levels = c("Group 1", "Group 2", "Group 3"))


library(scales)

# Plot
div_long_all_plot <- ggplot(div_long_all, aes(x = rule, y = Value, fill = Metric)) +
  geom_col(show.legend = FALSE, width = 0.7) +
    facet_grid(Metric ~ groupid, scales = "free_y", labeller = labeller(Metric = c(
             "Shannon's Diversity" = "Shannon's\nDiversity",
             "Pilou's Evenness" = "Pielou's\nEvenness",
             "Proportional Richness" = "Proportional\nRichness"))) +
  scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = label_number(accuracy = 0.1)) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "grey85"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(, angle = 90, size = 12, hjust = 1),
    strip.text = element_text(face = "bold", size = 12),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5)
  ) +
  labs(
    title = "Diversity Metrics of Response Motifs by Rule",
    x = NULL,
    y = NULL
  )
#ggsave("figure.pdf", plot = div_long_all_plot, width = 8, height = 6, units = "in")



```







#boxplot that compares rule speed across rules for each group
```{r}
head (retal_length_ready_2021)
head(pass_length_ready)
head(pile_length_ready_2021)
head(pun_length_ready_2021)
head(opp_length_ready_2021)


#add rule type column and standardize column name for all datasets
#could use loop but wanted to do one by one to keep a close eye

retal_comb <- mutate(retal_length_ready_2021, rule_type = "Retaliation", rule_time = retal_length_ready_2021$time_to_retal, Group = "1")

retal_comb_sub <- subset(retal_comb, select = c(rule_type, rule_time, Group))



pass_comb <- mutate(pass_length_ready, rule_type = "Pass-along", rule_time = pass_length_ready$time_to_pass, Group = "1")

pass_comb_sub <- subset(pass_comb, select = c(rule_type, rule_time, Group))



pile_comb <- mutate(pile_length_ready_2021, rule_type = "Pile-on", rule_time = pile_length_ready_2021$time_to_pile, Group = "1")

pile_comb_sub <- subset(pile_comb, select = c(rule_type, rule_time, Group))


pun_comb <- mutate(pun_length_ready_2021, rule_type = "Punishment", rule_time = pun_length_ready_2021$time_to_pun, Group = "1")

pun_comb_sub <- subset(pun_comb, select = c(rule_type, rule_time, Group))



opp_comb <- mutate(opp_length_ready_2021, rule_type = "Opportunism", rule_time = opp_length_ready_2021$time_to_opp, Group = "1")

opp_comb_sub <- subset(opp_comb, select = c(rule_type, rule_time, Group))




retal_length_ready2022G1



retal_comb_2022G1 <- mutate(retal_length_ready2022G1, rule_type = "Retaliation", rule_time = retal_length_ready2022G1$time_to_retal, Group = "2")

retal_comb_2022G1_sub <- subset(retal_comb_2022G1, select = c(rule_type, rule_time, Group))



pass_comb_2022G1 <- mutate(pass_length_ready_2022G1, rule_type = "Pass-along", rule_time = pass_length_ready_2022G1$time_to_pass, Group = "2")

pass_comb_2022G1_sub <- subset(pass_comb_2022G1, select = c(rule_type, rule_time, Group))



pile_comb_2022G1 <- mutate(pile_length_2022G1, rule_type = "Pile-on", rule_time = pile_length_2022G1$time_to_pile, Group = "2")

pile_comb_2022G1_sub <- subset(pile_comb_2022G1, select = c(rule_type, rule_time, Group))


pun_comb_2022G1 <- mutate(pun_length_ready_2022G1, rule_type = "Punishment", rule_time = pun_length_ready_2022G1$time_to_pun, Group = "2")

pun_comb_2022G1_sub <- subset(pun_comb_2022G1, select = c(rule_type, rule_time, Group))



opp_comb_2022G1 <- mutate(opp_length_ready_2022G1, rule_type = "Opportunism", rule_time = opp_length_ready_2022G1$time_to_opp, Group = "2")

opp_comb_2022G1_sub <- subset(opp_comb_2022G1, select = c(rule_type, rule_time, Group))










retal_comb_2022G2 <- mutate(retal_length_ready2022G2, rule_type = "Retaliation", rule_time = retal_length_ready2022G2$time_to_retal, Group = "3")

retal_comb_2022G2_sub <- subset(retal_comb_2022G2, select = c(rule_type, rule_time, Group))



pass_comb_2022G2 <- mutate(pass_length_ready_2022G2, rule_type = "Pass-along", rule_time = pass_length_ready_2022G2$time_to_pass, Group = "3")

pass_comb_2022G2_sub <- subset(pass_comb_2022G2, select = c(rule_type, rule_time, Group))



pile_comb_2022G2 <- mutate(pile_length_2022G2, rule_type = "Pile-on", rule_time = pile_length_2022G2$time_to_pile, Group = "3")

pile_comb_2022G2_sub <- subset(pile_comb_2022G2, select = c(rule_type, rule_time, Group))


pun_comb_2022G2 <- mutate(pun_length_ready_2022G2, rule_type = "Punishment", rule_time = pun_length_ready_2022G2$time_to_pun, Group = "3")

pun_comb_2022G2_sub <- subset(pun_comb_2022G2, select = c(rule_type, rule_time, Group))



opp_comb_2022G2 <- mutate(opp_length_ready_2022G2, rule_type = "Opportunism", rule_time = opp_length_ready_2022G2$time_to_opp, Group = "3")

opp_comb_2022G2_sub <- subset(opp_comb_2022G2, select = c(rule_type, rule_time, Group))





#bind all dfs together
rules.length.data.frame <- rbind(retal_comb_sub, pile_comb_sub, pass_comb_sub, pun_comb_sub, opp_comb_sub, retal_comb_2022G1_sub, pile_comb_2022G1_sub, pass_comb_2022G1_sub, pun_comb_2022G1_sub, opp_comb_2022G1_sub, retal_comb_2022G2_sub, pile_comb_2022G2_sub, pass_comb_2022G2_sub, pun_comb_2022G2_sub, opp_comb_2022G2_sub) 
head(rules.length.data.frame)

rules.length.data.frame$rule_time <- as.numeric(rules.length.data.frame$rule_time)

#write.csv(rules.length.data.frame,  here("ch1 data","rules.length.data.frame.csv"), row.names = FALSE)




rules.length.data.frame <- read_csv(here("ch1 data","rules.length.data.frame.csv"))




#group 2 has 0 retlaiations so adding a 0 manually so it shows up on plot
rules.length.data.frame <- subset(rules.length.data.frame, select = c("rule_type", "rule_time", "Group"))

rules.length.data.frame <- rbind(rules.length.data.frame, data.frame(rule_type = "Retaliation", rule_time = 0, Group = 2))


#ensure that group number is a factor
rules.length.data.frame$Group<- as.character( rules.length.data.frame$Group)





#compute quantiles for each rule
qs_rule <- rules.length.data.frame %>%
  group_by(rule_type) %>%
  summarize(q50 = quantile(rule_time, 0.50, na.rm = TRUE),
            q75 = quantile(rule_time, 0.75, na.rm = TRUE),
            q95 = quantile(rule_time, 0.95, na.rm = TRUE),
            .groups = "drop") %>%
  pivot_longer(starts_with("q"), names_to = "q", values_to = "y")

#Expand  rule-level quantiles to each Group so they can be dodged
all_groups <- unique(rules.length.data.frame$Group)
qs_rule_group <- tidyr::crossing(qs_rule, Group = all_groups)

#keep same Group order as data to help dodge alignment)
qs_rule_group$Group <- factor(qs_rule_group$Group,
                              levels = levels(factor(rules.length.data.frame$Group)))

pd <- position_dodge2(width = 0.75, preserve = "single")

test <- ggplot(rules.length.data.frame, aes(x = rule_type, y = rule_time, fill = Group)) +
  geom_boxplot(position = pd) +
  geom_crossbar(
    data = qs_rule_group,
    aes(x = rule_type, y = y, ymin = y, ymax = y, linetype = q, group = Group),
    inherit.aes = FALSE,
    position = pd,
    width = 1,      # line length inside each group’s box slot
    size = 0.25,
    color = "blue"
  ) +
  scale_linetype_manual(
    values = c(q50 = "dashed", q75 = "dashed", q95 = "dashed"),
    labels = c(q50 = "50%", q75 = "75%", q95 = "95%"),
    name = "Quantiles"
  ) +
  scale_x_discrete(limits = c("Retaliation", "Pile-on", "Pass-along", "Opportunism", "Punishment")) +
  labs(x = "Time Dependent Rule", y = "Speed of rule follows (s)") +
  theme_minimal(base_size = 15) +
  theme(legend.position = c(0.92, 0.92)) +
  scale_fill_brewer(palette = "Set2") +
  coord_cartesian(ylim = c(0, 2750)) +
  scale_y_continuous(breaks = seq(0, 2750, 300))

test
#weird but workable. fix position of quantiles in inkscape











#plot boxplot of comparisons of rule speeds for all 3 groups
it <- ggplot(rules.length.data.frame,aes(x =rule_type, y = rule_time, fill = Group)) + 
      stat_boxplot(geom = "errorbar") +

  geom_boxplot(show.legend = FALSE) + 
  scale_x_discrete(limits=c("Retaliation", "Pile-on", "Pass-along", "Opportunism" , "Punishment")) + 
  labs(x="Time Dependent Rule Sequence", y="Sequence speed (s)") +
  theme(legend.position = "none")  +
  scale_fill_brewer(palette = "Set2") +
  scale_color_brewer(palette = "Set2")  + 
    coord_cartesian(ylim = c(0, 2750)) + scale_y_continuous(breaks = seq(0, 2750, 300)) + theme_minimal(base_size = 15)


it













```



#GLMM and EMD to compare rules speeds between rules across the three groups
Group 1
```{r}
#GLMM and EMD
 library(multcomp)
library(emmeans)
library(MuMIn)


#Group 1

#whole procedure could be done with loop but did it one by one to keep a close eye step by step
#prepare dfs for mixed model
retal_comb_ready_for_mm_2021 <- subset(retal_comb, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 
pile_comb_ready_for_mm_2021 <- subset(pile_comb, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 
pass_comb_ready_for_mm_2021 <- subset(pass_comb, select = c(actor, subject, actor_2, subject_2,  rule_type, rule_time, date_period, Group)) 
opp_comb_ready_for_mm_2021 <- subset(opp_comb, select = c(actor, subject, actor_2, subject_2,  rule_type, rule_time, date_period, Group)) 
pun_comb_ready_for_mm_2021 <- subset(pun_comb, select = c(actor, subject, actor_2, subject_2,  rule_type, rule_time, date_period, Group)) 




#combine into one summary_heatmap_file for mixed model and make sure rule time is numeric
rules.length.data.frame_MM_2021 <- rbind(retal_comb_ready_for_mm_2021, pile_comb_ready_for_mm_2021, pass_comb_ready_for_mm_2021, opp_comb_ready_for_mm_2021, pun_comb_ready_for_mm_2021)
rules.length.data.frame_MM_2021$rule_time <- as.numeric(rules.length.data.frame_MM_2021$rule_time )



#write.csv(rules.length.data.frame_MM_2021,  here("ch1 data","rules.length.data.frame_MM_2021.csv"), row.names = FALSE)




rules.length.data.frame_MM_2021 <- read_csv(here("ch1 data","rules.length.data.frame_MM_2021.csv"))








#mixed model
obs_time_mod_2021 <- glmer(rule_time ~ rule_type + (1 | date_period)   + (1|subject)  , data = rules.length.data.frame_MM_2021 , control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)), family = Gamma(link = "log"))

summary_obs_time_mod_2021 <- summary(obs_time_mod_2021)

summary_obs_time_mod_2021

#CIs for base model
ci_model <- confint(obs_time_mod_2021, method = "Wald")
ci_model

library(MuMIn)
r.squaredGLMM(obs_time_mod_2021)

#post hoc multiple comparisons
posthoc_2021 <- glht(obs_time_mod_2021, linfct = mcp(rule_type = "Tukey"))
summary(posthoc_2021)

#CIs for post hoc comparisons
ci_tukey <- confint(posthoc_2021)
ci_tukey








#EMD ret pile pass opp pun
emd_ret_vs_pile_2021 <- emdist::emd(summary.pile_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
emd_ret_vs_pile_2021
#192.5961



emd_ret_vs_pass_2021 <- emdist::emd(summary.pass_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
emd_ret_vs_pass_2021
#367.9533

emd_ret_vs_opp_2021 <- emdist::emd(summary.opp_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
emd_ret_vs_opp_2021
#415.3581

emd_ret_vs_pun_2021 <- emdist::emd(summary.pun_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
emd_ret_vs_pun_2021
#226.5101


emd_pile_vs_pass_2021 <- emdist::emd(summary.pile_length_ready_2021_density_no_count_reordered_mat, summary.pass_length_ready_2021_density_no_count_reordered_mat)
emd_pile_vs_pass_2021
#542.8511

emd_pile_vs_opp_2021 <- emdist::emd(summary.pile_length_ready_2021_density_no_count_reordered_mat, summary.opp_length_ready_2021_density_no_count_reordered_mat)
emd_pile_vs_opp_2021
#606.71

emd_pile_vs_pun_2021 <- emdist::emd(summary.pile_length_ready_2021_density_no_count_reordered_mat, summary.pun_length_ready_2021_density_no_count_reordered_mat)
emd_pile_vs_pun_2021
#368.1497

emd_pass_vs_opp_2021 <- emdist::emd(summary.pass_length_ready_2021_density_no_count_reordered_mat, summary.opp_length_ready_2021_density_no_count_reordered_mat)
emd_pass_vs_opp_2021
#81.64326

emd_pass_vs_pun_2021 <- emdist::emd(summary.pass_length_ready_2021_density_no_count_reordered_mat, summary.pun_length_ready_2021_density_no_count_reordered_mat)
emd_pass_vs_pun_2021
#203.1274

emd_opp_vs_pun_2021 <- emdist::emd(summary.opp_length_ready_2021_density_no_count_reordered_mat, summary.pun_length_ready_2021_density_no_count_reordered_mat)
emd_opp_vs_pun_2021
#242.0948

emd_2021_all <- c(emd_pass_vs_opp_2021, emd_pile_vs_opp_2021, emd_opp_vs_pun_2021,  emd_ret_vs_opp_2021, emd_pile_vs_pass_2021, emd_pass_vs_pun_2021, emd_ret_vs_pass_2021, emd_pile_vs_pun_2021, emd_ret_vs_pile_2021, emd_ret_vs_pun_2021)



mean_rule_time_2021_summary <- rules.length.data.frame_MM_2021 %>% group_by(rule_type) %>%
  dplyr::summarize (mean(rule_time))
mean_rule_time_2021_summary
```
Group 2
```{r}
#GLMM and EMD
 library(multcomp)

#Group 2
#prepare dfs for mixed model
#retal_comb_ready_for_mm_2022G1 <- subset(retal_comb_2022G1, select = c(actor, subject, actor_2, rule_type, rule_time, date_period, Group)) 
pile_comb_ready_for_mm_2022G1 <- subset(pile_comb_2022G1, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 
pass_comb_ready_for_mm_2022G1 <- subset(pass_comb_2022G1, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 
opp_comb_ready_for_mm_2022G1 <- subset(opp_comb_2022G1, select = c(actor, subject, actor_2,  subject_2, rule_type, rule_time, date_period, Group)) 
pun_comb_ready_for_mm_2022G1 <- subset(pun_comb_2022G1, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 




#combine into one summary_heatmap_file for mixed model and make sure rule time is numeric
rules.length.data.frame_MM_2022G1 <- rbind( pile_comb_ready_for_mm_2022G1, pass_comb_ready_for_mm_2022G1, opp_comb_ready_for_mm_2022G1, pun_comb_ready_for_mm_2022G1)
rules.length.data.frame_MM_2022G1$rule_time <- as.numeric(rules.length.data.frame_MM_2022G1$rule_time )


#write.csv(rules.length.data.frame_MM_2022G1,  here("ch1 data","rules.length.data.frame_MM_2022G1.csv"), row.names = FALSE)




rules.length.data.frame_MM_2022G1 <- read_csv(here("ch1 data","rules.length.data.frame_MM_2022G1.csv"))








#mixed model
obs_time_mod_2022G1 <- glmer(rule_time ~ rule_type + (1 | date_period)     , data = rules.length.data.frame_MM_2022G1 , control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)), family = Gamma(link = "log"))

summary_obs_time_mod_2022G1 <- summary(obs_time_mod_2022G1)

summary_obs_time_mod_2022G1

#CIs for base model
ci_model <- confint(obs_time_mod_2022G1, method = "Wald")
ci_model


r.squaredGLMM(obs_time_mod_2022G1)

#post hoc multiple comparisons
posthoc_2022G1 <- glht(obs_time_mod_2022G1, linfct = mcp(rule_type = "Tukey"))
summary(posthoc_2022G1)

#CIs for post hoc comparisons
ci_tukey <- confint(posthoc_2022G1)
ci_tukey




#EMD ret pile pass opp pun
#emd_ret_vs_pile_2022G1 <- emdist::emd(summary.pile_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
#emd_ret_vs_pile_2022G1
#



#emd_ret_vs_pass_2022G1 <- emdist::emd(summary.pass_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
#emd_ret_vs_pass_2022G1
#

#emd_ret_vs_opp_2022G1 <- emdist::emd(summary.opp_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
#emd_ret_vs_opp_2022G1
#

#emd_ret_vs_pun_2022G1 <- emdist::emd(summary.pun_length_ready_2021_density_no_count_reordered_mat, summary.retal_length_ready_2021_density_no_count_reordered_mat)  
#emd_ret_vs_pun_2022G1
#


emd_pile_vs_pass_2022G1 <- emdist::emd(summary.pile_length_ready_2022G1_density_no_count_reordered_mat, summary.pass_length_ready_2022G1_density_no_count_reordered_mat)
emd_pile_vs_pass_2022G1
#229.910

emd_pile_vs_opp_2022G1 <- emdist::emd(summary.pile_length_ready_2022G1_density_no_count_reordered_mat, summary.opp_length_ready_2022G1_density_no_count_reordered_mat)
emd_pile_vs_opp_2022G1
#522.444


emd_pile_vs_pun_2022G1 <- emdist::emd(summary.pile_length_ready_2022G1_density_no_count_reordered_mat, summary.pun_length_ready_2022G1_density_no_count_reordered_mat)
emd_pile_vs_pun_2022G1
#167.085

emd_pass_vs_opp_2022G1 <- emdist::emd(summary.pass_length_ready_2022G1_density_no_count_reordered_mat, summary.opp_length_ready_2022G1_density_no_count_reordered_mat)
emd_pass_vs_opp_2022G1
#296.2308

emd_pass_vs_pun_2022G1 <- emdist::emd(summary.pass_length_ready_2022G1_density_no_count_reordered_mat, summary.pun_length_ready_2022G1_density_no_count_reordered_mat)
emd_pass_vs_pun_2022G1
#105.9524

emd_opp_vs_pun_2022G1 <- emdist::emd(summary.opp_length_ready_2022G1_density_no_count_reordered_mat, summary.pun_length_ready_2022G1_density_no_count_reordered_mat)
emd_opp_vs_pun_2022G1
#356.327






emd_2022G1_all <- c(emd_pass_vs_opp_2022G1, emd_pile_vs_opp_2022G1, emd_opp_vs_pun_2022G1, emd_pile_vs_pass_2022G1, emd_pass_vs_pun_2022G1,  emd_pile_vs_pun_2022G1)




mean_rule_time_2022G1_summary <- rules.length.data.frame_MM_2022G1 %>% group_by(rule_type) %>%
  dplyr::summarize (mean(rule_time))
mean_rule_time_2022G1_summary

```

Group 3
```{r}
#GLMM and EMD
 library(multcomp)

#Group 3
#prepare dfs for mixed model
retal_comb_ready_for_mm_2022G2 <- subset(retal_comb_2022G2, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 
pile_comb_ready_for_mm_2022G2 <- subset(pile_comb_2022G2, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 
pass_comb_ready_for_mm_2022G2 <- subset(pass_comb_2022G2, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 
opp_comb_ready_for_mm_2022G2 <- subset(opp_comb_2022G2, select = c(actor, subject, actor_2,  subject_2, rule_type, rule_time, date_period, Group)) 
pun_comb_ready_for_mm_2022G2 <- subset(pun_comb_2022G2, select = c(actor, subject, actor_2, subject_2, rule_type, rule_time, date_period, Group)) 




#combine into one summary_heatmap_file for mixed model and make sure rule time is numeric
rules.length.data.frame_MM_2022G2 <- rbind(retal_comb_ready_for_mm_2022G2, pile_comb_ready_for_mm_2022G2, pass_comb_ready_for_mm_2022G2, opp_comb_ready_for_mm_2022G2, pun_comb_ready_for_mm_2022G2)
rules.length.data.frame_MM_2022G2$rule_time <- as.numeric(rules.length.data.frame_MM_2022G2$rule_time )



#write.csv(rules.length.data.frame_MM_2022G2,  here("ch1 data","rules.length.data.frame_MM_2022G2.csv"), row.names = FALSE)




rules.length.data.frame_MM_2022G2 <- read_csv(here("ch1 data","rules.length.data.frame_MM_2022G2.csv"))








#mixed model
obs_time_mod_2022G2 <- glmer(rule_time ~ rule_type + (1 | date_period)     , data = rules.length.data.frame_MM_2022G2 , control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)), family = Gamma(link = "log"))

summary_obs_time_mod_2022G2 <- summary(obs_time_mod_2022G2)

summary_obs_time_mod_2022G2

#CIs for base model
ci_model <- confint(obs_time_mod_2022G2, method = "Wald")
ci_model


r.squaredGLMM(obs_time_mod_2022G2)

#post hoc multiple comparisons
posthoc_2022G2 <- glht(obs_time_mod_2022G2, linfct = mcp(rule_type = "Tukey"))
summary(posthoc_2022G2)

#CIs for post hoc comparisons
ci_tukey <- confint(posthoc_2022G2)
ci_tukey




#EMD ret pile pass opp pun
emd_ret_vs_pile_2022G2 <- emdist::emd(summary.pile_length_ready_2022G2_density_no_count_reordered_mat, summary.retal_length_ready2022G2_density_no_count_reordered_mat)  
emd_ret_vs_pile_2022G2
#332.314



emd_ret_vs_pass_2022G2 <- emdist::emd(summary.pass_length_ready_2022G2_density_no_count_reordered_mat, summary.retal_length_ready2022G2_density_no_count_reordered_mat)  
emd_ret_vs_pass_2022G2
#163.516

emd_ret_vs_opp_2022G2 <- emdist::emd(summary.opp_length_ready_2022G2_density_no_count_reordered_mat, summary.retal_length_ready2022G2_density_no_count_reordered_mat)  
emd_ret_vs_opp_2022G2
#186.9

emd_ret_vs_pun_2022G2 <- emdist::emd(summary.pun_length_ready_2022G2_density_no_count_reordered_mat, summary.retal_length_ready2022G2_density_no_count_reordered_mat)  
emd_ret_vs_pun_2022G2
#207.436


emd_pile_vs_pass_2022G2 <- emdist::emd(summary.pile_length_ready_2022G2_density_no_count_reordered_mat, summary.pass_length_ready_2022G2_density_no_count_reordered_mat)
emd_pile_vs_pass_2022G2
#226.501

emd_pile_vs_opp_2022G2 <- emdist::emd(summary.pile_length_ready_2022G2_density_no_count_reordered_mat, summary.opp_length_ready_2022G2_density_no_count_reordered_mat)
emd_pile_vs_opp_2022G2
#307.814

emd_pile_vs_pun_2022G2 <- emdist::emd(summary.pile_length_ready_2022G2_density_no_count_reordered_mat, summary.pun_length_ready_2022G2_density_no_count_reordered_mat)
emd_pile_vs_pun_2022G2
#146.332

emd_pass_vs_opp_2022G2 <- emdist::emd(summary.pass_length_ready_2022G2_density_no_count_reordered_mat, summary.opp_length_ready_2022G2_density_no_count_reordered_mat)
emd_pass_vs_opp_2022G2
#82.245

emd_pass_vs_pun_2022G2 <- emdist::emd(summary.pass_length_ready_2022G2_density_no_count_reordered_mat, summary.pun_length_ready_2022G2_density_no_count_reordered_mat)
emd_pass_vs_pun_2022G2
#95.987

emd_opp_vs_pun_2022G2 <- emdist::emd(summary.opp_length_ready_2022G2_density_no_count_reordered_mat, summary.pun_length_ready_2022G2_density_no_count_reordered_mat)
emd_opp_vs_pun_2022G2
#165.888




emd_2022G2_all <- c(emd_pass_vs_opp_2022G2, emd_pile_vs_opp_2022G2, emd_opp_vs_pun_2022G2,  emd_ret_vs_opp_2022G2, emd_pile_vs_pass_2022G2, emd_pass_vs_pun_2022G2, emd_ret_vs_pass_2022G2, emd_pile_vs_pun_2022G2, emd_ret_vs_pile_2022G2, emd_ret_vs_pun_2022G2)





mean_rule_time_2022G2_summary <- rules.length.data.frame_MM_2022G2 %>% group_by(rule_type) %>%
  dplyr::summarize (mean(rule_time))
mean_rule_time_2022G2_summary
```




#time faceted plot
```{r}
####WHEN YOU DO THIS MAKE SURE P VALS ARE UPDATED HERE AND IN MANU

#could use loop 


#subset dfs to prepare for combining and plotting
retal_time_comb_2021 <- subset(retal_length_ready_2021_comb, select = c(time_to_retal, data_type))
retal_time_comb_2022G1 <- subset(retal_length_ready_comb_2022G1, select = c(time_to_retal, data_type))
retal_time_comb_2022G2 <- subset(retal_length_ready_2022G2_comb, select = c(time_to_retal, data_type))



pile_time_comb_2021 <- subset(pile_length_ready_2021_comb, select = c(time_to_pile, data_type))
pile_time_comb_2022G1 <- subset(pile_length_ready_2022G1_comb, select = c(time_to_pile, data_type))
pile_time_comb_2022G2 <- subset(pile_length_ready_2022G2_comb, select = c(time_to_pile, data_type))




pass_time_comb_2021 <- subset(pass_length_ready_comb, select = c(time_to_pass, data_type))
pass_time_comb_2022G1 <- subset(pass_length_ready_2022G1_comb, select = c(time_to_pass, data_type))
pass_time_comb_2022G2 <- subset(pass_length_ready_2022G2_comb, select = c(time_to_pass, data_type))



opp_time_comb_2021 <- subset(opp_length_ready_2021_comb, select = c(time_to_opp, data_type))
opp_time_comb_2022G1 <- subset(opp_length_ready_2022G1_comb, select = c(time_to_opp, data_type))
opp_time_comb_2022G2 <- subset(opp_length_ready_2022G2_comb, select = c(time_to_opp, data_type))


pun_time_comb_2021 <- subset(pun_length_ready_2021_comb, select = c(time_to_pun, data_type))
pun_time_comb_2022G1 <- subset(pun_length_ready_2022G1_comb, select = c(time_to_pun, data_type))
pun_time_comb_2022G2 <- subset(pun_length_ready_2022G2_comb, select = c(time_to_pun, data_type))




#make all column names consistent
colnames(retal_time_comb_2021)[1] <- 'time_to_rule'
colnames(retal_time_comb_2022G1)[1] <- 'time_to_rule'
colnames(retal_time_comb_2022G2)[1] <- 'time_to_rule'
colnames(pile_time_comb_2021)[1] <- 'time_to_rule'
colnames(pile_time_comb_2022G1)[1] <- 'time_to_rule'
colnames(pile_time_comb_2022G2)[1] <- 'time_to_rule'
colnames(pass_time_comb_2021)[1] <- 'time_to_rule'
colnames(pass_time_comb_2022G1)[1] <- 'time_to_rule'
colnames(pass_time_comb_2022G2)[1] <- 'time_to_rule'
colnames(opp_time_comb_2021)[1] <- 'time_to_rule'
colnames(opp_time_comb_2022G1)[1] <- 'time_to_rule'
colnames(opp_time_comb_2022G2)[1] <- 'time_to_rule'
colnames(pun_time_comb_2021)[1] <- 'time_to_rule'
colnames(pun_time_comb_2022G1)[1] <- 'time_to_rule'
colnames(pun_time_comb_2022G2)[1] <- 'time_to_rule'




# here



#bind together into one summary_heatmap_file
tdr_time_comb<- rbind(
  cbind(retal_time_comb_2021, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(retal_time_comb_2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(retal_time_comb_2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(pile_time_comb_2021, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(pile_time_comb_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(pile_time_comb_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(pass_time_comb_2021, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(pass_time_comb_2022G1, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(pass_time_comb_2022G2, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(opp_time_comb_2021, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(opp_time_comb_2022G1, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(opp_time_comb_2022G2, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(pun_time_comb_2021, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(pun_time_comb_2022G1, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(pun_time_comb_2022G2, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

#write.csv(tdr_time_comb,  here("ch1 data","tdr_time_comb.csv"), row.names = FALSE)


#tdr_time_comb <- read_csv(here("ch1 data","tdr_time_comb.csv"))




#tdr_time_comb$Group <- factor(tdr_time_comb$Group, levels = c("Retalialtion Group 1", "Retaliation Group 2", "Retaliation Group 3", "Pile-on Group 1", "Pile-on Group 2", "Pile-on Group 3", "Pass-along Group 1", "Pass-along Group 2", "Pass-along Group 3", "Opportunism Group 1", "Opportunism Group 2", "Opportunism Group 3", "Punishment Group 1", "Punishment Group 2", "Punishment Group 3"))

#store each rule as a factor in the particular order I would like them displayed in my plot
custom_order <- c("Retalialtion Group 1", "Retaliation Group 2", "Retaliation Group 3", "Pile-on Group 1", "Pile-on Group 2", "Pile-on Group 3", "Pass-along Group 1", "Pass-along Group 2", "Pass-along Group 3", "Opportunism Group 1", "Opportunism Group 2", "Opportunism Group 3", "Punishment Group 1", "Punishment Group 2", "Punishment Group 3")

# Convert group column to a factor with the desired order
tdr_time_comb$Group <- factor(tdr_time_comb$Group, levels = custom_order)

#row with NA from retal 2022G1. store and chop off so I can add p values
row_with_na <- tdr_time_comb[apply(tdr_time_comb, 1, function(x) any(is.na(x))), ]

#remove NAs
tdr_time_comb <- na.omit(tdr_time_comb)





#store p values in the particular order Id like to place on my faceted plots
pval_time_df <- data.frame(
  Group = custom_order,    # Match facet group names
  label = c("p = 0.042", "p = NA", "p = 0.847", "p < 0.001",
            "p < 0.001", "p < 0.001", "p < 0.001", "p = 0.856",
            "p = 0.359", "p < 0.001", "p = 0.033", "p = 0.012",
            "p = 0.367", "p = 0.392", "p = 0.282"), # p values
    # 90% of max count in each group to keep text inside
    x_pos = max(tdr_time_comb$time_to_rule) * 0.9,  
    # 90% of max density per group
    y_pos = max(density(tdr_time_comb$time_to_rule)$y) * 0.9  
)





bold_values_time <- c("p < 0.001", "p = 0.042", "p = 0.033", "p = 0.012")

pval_time_df$font_weight <- ifelse(pval_time_df$label %in% bold_values_time, "bold", "plain")




# Ensure pval_time_df respect the custom order
pval_time_df$Group <- factor(pval_time_df$Group, levels = custom_order)

#bind the neccessary na row back onto tdr_time_comb
tdr_time_comb <- rbind(tdr_time_comb, row_with_na)



#faceted set of split violin plots that compares observed rule speed to random rule speed for each rule in each group
tdr_time_plot <- ggplot(tdr_time_comb, aes(x = rule, y = time_to_rule, fill = data_type)) +
  introdataviz::geom_split_violin(alpha= .9, trim = TRUE, show.legend = FALSE) +
  geom_boxplot(width = .2, alpha = .2, fatten = 2, show.legend = FALSE, outlier.shape = NA) +
  geom_text(data = pval_time_df, aes(x = Inf, y = Inf, label = label, fontface = font_weight),
                inherit.aes = FALSE,  
  hjust = 1.5, vjust = 1.5, color = "black", size = 9) +  # Custom text in top-right
  facet_wrap( ~ Group, scales = "free_x", ncol = 3, nrow = 5 ) +
  theme_minimal() + 
  theme(
    strip.text = element_text(size = 30, face = "bold"),
    axis.text = element_text(size = 30),  # Increase axis text size
    axis.title = element_text(size = 30, face = "bold"),
    axis.text.x = element_blank(),
  axis.ticks.x = element_line(),
    panel.spacing = unit(1, "lines"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    )#,
    #axis.text.x = element_blank(),  # Remove x-axis labels (since x is empty)
    #axis.ticks.x = element_blank()   # Remove x-axis ticks
  ) +
scale_x_discrete(name = "Density") +  
     coord_cartesian(ylim = c(0, 1000)) +
  scale_y_continuous(name = "Speed of rule follows (s)",
                     breaks = seq(0, 1000, 200) 
                     ) + scale_fill_manual(values=c("red","gray"), name = "Data Type")
  
#ggsave("figure.pdf", plot = tdr_time_plot, width = 20, height = 22, units = "in")


#18 by 20 either







#store p values in particular order 
pval_df <- data.frame(
  Group = custom_order,    # Match facets with p values
  label = c("p = 0.174", "p = 0.16", "p = 0.036", "p = 0",
            "p = 0.122", "p = 0.042", "p = 0", "p = 0",
            "p = 0", "p = 0", "p = 0", "p = 0.492",
            "p = 0", "p = 0.028", "p = 0.003"),   
  # 90% of max count in each group to keep text inside
    x_pos = max(tdr_count_rand_comb$count_of_rule_follows) * 0.9,  
   # 90% of max density per group
    y_pos = max(density(tdr_count_rand_comb$count_of_rule_follows)$y) * 0.9 
)


bold_values_count <- c("p = 0.036", "p = 0","p = 0.042", "p = 0.028", "p = 0.003")

pval_df$font_weight <- ifelse(pval_df$label %in% bold_values_count, "bold", "plain")


# Ensure pval_df respect the custom order
pval_df$Group <- factor(pval_df$Group, levels = custom_order)






# Plot faceted set of violin plots that show how often each rule was followed compared to random chance. Observed rule follows vs randomized reference mdoels
rule_count_faceted_plot <- ggplot(tdr_count_rand_comb, aes(x = "", y = count_of_rule_follows)) +  # x is empty for density distribution
  geom_violin(fill = "grey", color = "black", alpha = 1) +  # Violin plot
    geom_point(data = tdr_count_real_comb, aes(x = "", y = count_of_rule_follows), color = "red", size = 5, shape = 8, stroke = 2) +  # Add red star
  geom_text(data = pval_df, aes(x = Inf, y = Inf, label = label, fontface = font_weight), 
            hjust = 1.5, vjust = 1.5, color = "black", size = 7) +  # Custom text in top-right
  facet_wrap( ~ Group, scales = "free_x", ncol = 3, nrow = 5 ) +
  theme_minimal() + 
  theme(
    strip.text = element_text(size = 30, face = "bold"),
    axis.text = element_text(size = 30),  # Increase axis text size
    axis.title = element_text(size = 30, face = "bold"),
    axis.text.x = element_text(size = 30),
  axis.ticks.x = element_line(),
    panel.spacing = unit(1, "lines"),# Y-axis tit
    panel.border = element_rect(          # Add boxes around each plot
      color = "black", fill = NA, linewidth = 1
    )#,
    #axis.text.x = element_blank(),  # Remove x-axis labels (since x is empty)
    #axis.ticks.x = element_blank()   # Remove x-axis ticks
  ) +
  labs(
       x = "Density",
       y = "Count")

rule_count_faceted_plot

#ggsave("figure.pdf", plot = rule_count_faceted_plot, width = 20, height = 22, units = "in")




```












#Richness and eveness plots for indiviudal rule use 2021
```{r}

diversity_summary_all 


richness_2021 <- c(retal_richness_2021, pile_richness_2021, pass_along_richness_2021, opp_richness_2021, punishment_richness_2021)

rule_names <-c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment")

richness_2021_df <- data.frame(rule_names, richness_2021)


ggplot(richness_2021_df, aes(x=rule_names, y=richness_2021, fill=rule_names)) +
         geom_bar(stat="identity")+theme_minimal() + scale_x_discrete(limits=c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment"))


evenness_2021 <- c(retal_even_value_2021$PielouEven, pile_even_value_2021$PielouEven, pass_along_even_value_2021$PielouEven, opp_even_value_2021$PielouEven, punishment_even_value_2021$PielouEven)

evenness_2021_df <- data.frame(rule_names, evenness_2021)

ggplot(evenness_2021_df, aes(x=rule_names, y=evenness_2021, fill=rule_names)) +
         geom_bar(stat="identity")+theme_minimal() + scale_x_discrete(limits=c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment"))

```



#Richness and eveness plots for indiviudal rule use 2022G1
```{r}
richness_2022G1 <- c(retal_richness_2022G1, pile_richness_2022G1, pass_along_richness_2022G1, opp_richness_2022G1, punishment_richness_2022G1)

rule_names <-c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment")

richness_2022G1_df <- data.frame(rule_names, richness_2022G1)


ggplot(richness_2022G1_df, aes(x=rule_names, y=richness_2022G1, fill=rule_names)) +
         geom_bar(stat="identity")+theme_minimal() + scale_x_discrete(limits=c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment"))


evenness_2022G1 <- c(retal_even_value_2022G1$PielouEven, pile_even_value_2022G1$PielouEven, pass_along_even_value_2022G1$PielouEven, opp_even_value_2022G1$PielouEven, punishment_even_value_2022G1$PielouEven)

evenness_2022G1_df <- data.frame(rule_names, evenness_2022G1)

ggplot(evenness_2022G1_df, aes(x=rule_names, y=evenness_2022G1, fill=rule_names)) +
         geom_bar(stat="identity")+theme_minimal() + scale_x_discrete(limits=c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment"))

```



#Richness and eveness plots for indiviudal rule use 2022G2
```{r}
richness_2022G2 <- c(retal_richness_2022G2, pile_richness_2022G2, pass_along_richness_2022G2, opp_richness_2022G2, punishment_richness_2022G2)

rule_names <-c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment")

richness_2022G2_df <- data.frame(rule_names, richness_2022G2)


ggplot(richness_2022G2_df, aes(x=rule_names, y=richness_2022G2, fill=rule_names)) +
         geom_bar(stat="identity")+theme_minimal() + scale_x_discrete(limits=c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment"))


evenness_2022G2 <- c(retal_even_value_2022G2$PielouEven, pile_even_value_2022G2$PielouEven, pass_along_even_value_2022G2$PielouEven, opp_even_value_2022G2$PielouEven, punishment_even_value_2022G2$PielouEven)

evenness_2022G2_df <- data.frame(rule_names, evenness_2022G2)

ggplot(evenness_2022G2_df, aes(x=rule_names, y=evenness_2022G2, fill=rule_names)) +
         geom_bar(stat="identity")+theme_minimal() + scale_x_discrete(limits=c("Retaliation", "Pile-on", "Pass-along", "Opportunism","Punishment"))

```