---
title: "punishment 2022G2 final"
author: "Xavier Francis"
date: "2024-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(data.table) 
library(ggplot2)
library(chemodiv)
library(introdataviz)
library(lme4)
library(multcomp)
library(sjPlot)
library(emdist)


```


#count of rule follows
```{r}
#start with both_numbered_sorted2022G2

#ensure that ranks are numeric
both_numbered_sorted2022G2$actor_rank <- as.numeric(both_numbered_sorted2022G2$actor_rank)
both_numbered_sorted2022G2$subject_rank <- as.numeric(both_numbered_sorted2022G2$subject_rank)

head(both_numbered_sorted2022G2)


#procedure to add obs session

#store the first row as a separate df. It will be binded back on later 
row_1_obs <- both_numbered_sorted2022G2[1,]

#put events side by side. This will allow us to subtract consecutive times from one another to determine whether an event took place on a separate observation session. 
obs_period_pt_1 <- mutate(both_numbered_sorted2022G2, actor_2 = lead(actor),
actor_2_rank = lead(actor_rank), 
subject_2 = lead(subject),                                   
subject_2_rank = lead(subject_rank), 
 full_time_2 = lead(full_time), 
date2 = lead(date))

#add empty columns for observation period and date period
obs_period_pt_1$obs_period <- NA
obs_period_pt_1$date_period <- NA

#create ticker for loop
q <- 1

#loop to calculate observation sessions. If consecutive events were on the same day but there was a long time between events, it is determined that the 2nd event took place in a separate observation session. The ticker goes up, and is reset if events took place on different days.
for (i in 1:nrow(obs_period_pt_1)) {ifelse(obs_period_pt_1$date[i] == obs_period_pt_1$date2[i]
                                            & obs_period_pt_1$full_time_2[i] - obs_period_pt_1$full_time[i] >= 1800, q <- q+1, q <- q)
  ifelse(obs_period_pt_1$date[i] != obs_period_pt_1$date2[i], q <- 1, q <- q)
  obs_period_pt_1$obs_period [i] <- q
  }

#remove original columns and keep mutated columns. These will replace original columns
obs_period_pt_2 <- subset(obs_period_pt_1, select = c(date2,  full_time_2, actor_2, actor_2_rank, subject_2, subject_2_rank, obs_period, date_period  ))

#get rid of the bottom, which will be full of NAs, as the last event in the original data does not have anything that comes after
obs_period_pt_3 <- slice(obs_period_pt_2, 1:(n() - 1))

#rename mutated columns with original names 
setnames(obs_period_pt_3, old = c('date2',  'full_time_2', 'actor_2', 'actor_2_rank', 'subject_2', 'subject_2_rank'), 
         new = c('date',  'full_time', 'actor', 'actor_rank', 'subject', 'subject_rank'))

#add observation period "1" onto first row that we subseted off, as the first event in the data has to take place in obs session 1
row_1_obs$obs_period <- 1 
row_1_obs$date_period <- NA

#ensure that row 1 matches the df
row_1_obs_ready <-  subset(row_1_obs, select = c(date,  full_time, actor, actor_rank, subject, subject_rank, obs_period, date_period  ))

#bind first row and df together
obs_period_pt_4 <- rbind(row_1_obs_ready, obs_period_pt_3)
#add date period 
obs_period_pt_4$date_period <- paste(obs_period_pt_4$date,obs_period_pt_4$obs_period)
head(obs_period_pt_4)




head(obs_period_pt_4)


#here

#Prepare/order & give each row an id + numeric time ---
df <- obs_period_pt_4 %>%
  arrange(date_period, date, full_time) %>%
  mutate(.row = row_number(),
         tsec = as.numeric(full_time))  # seconds since midnight

# Previous-by-actor (used for the base "A→B within 60s" check) ---
by_actor <- df %>%
  group_by(date_period, actor) %>%
  mutate(prev_subj_by_actor = lag(subject),
         prev_tsec_by_actor = lag(tsec)) %>%
  ungroup()

#Last involvement time for each participant (actor OR subject) ---
# Build a "long" view with one entry per participant per event
long <- bind_rows(
  by_actor %>% transmute(date_period, .row, participant = actor, tsec),
  by_actor %>% transmute(date_period, .row, participant = subject, tsec)
) %>%
  arrange(date_period, tsec, .row) %>%
  group_by(date_period, participant) %>%
  mutate(prev_involve_tsec = lag(tsec)) %>%
  ungroup()

#For each original row, get the most recent involvement of either A or B before this row
last_involve <- long %>%
  group_by(date_period, .row) %>%
  summarise(last_involve_tsec = max(prev_involve_tsec, na.rm = TRUE), .groups = "drop") %>%
  mutate(last_involve_tsec = ifelse(is.finite(last_involve_tsec), last_involve_tsec, NA_real_))

#Apply  rule of getting rid of consectuive 60 sec possible chases
aggression_no_repeat_custom <- by_actor %>%
  left_join(last_involve, by = c("date_period", ".row")) %>%
  mutate(
    #Base repeat condition: previous-by-same-actor is the same subject and within 60s
    within_actor_repeat = !is.na(prev_tsec_by_actor) &
                          subject == prev_subj_by_actor &
                          (tsec - prev_tsec_by_actor) <= 60,

    #Was there an intervening event involving A or B after that previous A→B?
    interrupted_by_A_or_B = within_actor_repeat &
                            !is.na(last_involve_tsec) &
                            (last_involve_tsec > prev_tsec_by_actor),

    keep = !within_actor_repeat | interrupted_by_A_or_B
  ) %>%
  filter(keep) %>%
dplyr::select(-.row, -tsec,
         -prev_subj_by_actor, -prev_tsec_by_actor,
         -last_involve_tsec, -within_actor_repeat, -interrupted_by_A_or_B, -keep)


obs <- aggression_no_repeat_custom$date_period
#separate data into a list with each df=a date 


#create empty list to store each bird each day each obs
pun_aggression_list_2022G2 <- list()

dater <- unique(obs)
datex <- c(1: length(unique(obs)))
actorer <- actor
actex <- c(1:length(actor))



#to here

#separate data into list of each obs session each day for each bird
for(i in datex) { 
  for (g in actex){
  output <- filter(aggression_no_repeat_custom, aggression_no_repeat_custom$date_period == dater[i] & (subject == actorer[g] | actor == actorer[g]))
   if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
  output <- mutate(output, counter = actorer[g])
      pun_aggression_list_2022G2[[paste0("list", dater[i], actorer[g])]] <- output}}






#ensure df arranged by date and time
pun_aggression_list_2022G2 <- lapply(pun_aggression_list_2022G2, function(x) arrange(x, date, full_time))






#put events side by side
punishment_side_by_side_2022G2 <- lapply(pun_aggression_list_2022G2, function(x) mutate(x, actor_2 = lead(actor), 
                                                                                 actor_2_rank = lead(actor_rank), 
                                                                                 subject_2 = lead(subject), 
                                                                                 subject_2_rank = lead(subject_rank), 
                                                                                 full_time_2 = lead(full_time), 
                                                                                 date2 = lead(date)))

#pick out punishment rule while keeping all dataframes in a list
 punishment_side_by_side_tested_2022G2 <- map(punishment_side_by_side_2022G2, ~ mutate(.x, initiated = ifelse(actor == counter
                                                                                                & subject_2 == counter
                                                                                                & actor_rank > subject_rank
                                                                                                , "yes","no"), 
                                      rule_followed = ifelse( subject_2 == counter
                                                              & actor == counter
                                                              & actor == subject_2
                                                              & full_time_2 - full_time <= 414 
                                                              & full_time_2 - full_time > 0
                                                              & date2 == date
                                                              & actor_2_rank < subject_2_rank
                                                              & actor_rank > subject_rank
                                                              & actor_2 != subject,
                                                              "yes","no")))
 
 #combine dataframes into one huge dataframe
 punishment_huge_2022G2 <- bind_rows(punishment_side_by_side_tested_2022G2, .id = "column_label")
 
 #sort by time and date
 punishment_huge_2022G2_sorted <- punishment_huge_2022G2 %>%
   arrange(punishment_huge_2022G2$date, punishment_huge_2022G2$full_time)
 
#Remove duplicate events, events that have been represented twice, and events that could not confidently be counted as the start of punishment
 punishment_huge_2022G2_sorted_no_dupr <- subset(punishment_huge_2022G2_sorted, initiated != "no") 
 
 
 #subset for desired columns
 punishment_no_dupr <- subset(punishment_huge_2022G2_sorted_no_dupr, select = c(date, full_time, actor, actor_rank, subject, subject_rank, 
                                                                         counter, actor_2, actor_2_rank, subject_2, subject_2_rank, full_time_2, date2, 
                                                                         initiated, rule_followed))
 
 table(punishment_no_dupr$rule_followed)
 
 #determine how often punishment was followed
 punishment_real_2022G2 <- length(which(punishment_no_dupr$rule_followed == "yes"))
 
punishment_real_2022G2_frame <- data.frame(punishment_real_2022G2)
 
 
 
 #create df of only rule follows
 punishment_doner_all_yes_2022G2 <-filter(punishment_no_dupr, rule_followed == "yes")
 #
 
 
 
#22 for 1328 
 
 
 
 
 
 
 
 
#13 for 169
```

#individ rule use
```{r}


#test for individual use of rule. How many times did each bird use rule
unique(punishment_doner_all_yes_2022G2$actor_2)
length(unique(punishment_doner_all_yes_2022G2$actor_2)
)

summary.pun.individ_2022G2 <- punishment_doner_all_yes_2022G2 %>% group_by(actor_2) %>%
  summarize (
    count =n())

summary.pun.individ_2022G2



#show Liz
#unique(`2022_interactions_cleaned`$actor)


#compare to raw dataset to see if any birds didnt use rule at all
unique(t2022_aggression_onlyG2$actor)
unique(t2022_aggression_onlyG2$subject)

#setdiff(bird.list, unique(t2022_aggression_only_readyG2$subject) )

#for each rule keep in mind how many birds can actually use the rule
#bird who didnt do rule was bottom ranked and cant do rule
punishment_nofollow <- setdiff(unique(t2022_aggression_onlyG2$actor), unique(punishment_doner_all_yes_2022G2$actor_2)
)
punishment_nofollow
#DO FIRST BEFORE ADDING ZEROS AND DONT DO AGAIN WITHOUT RUNNING ABOVE CODE
#Calculate Richness score for rule following
punishment_richness <- length(summary.pun.individ_2022G2$actor_2)/10
#did -1 because pgg cant ever punish since its lowest rank

 #add in 0s for individuals who did not follow rule for plotting purposes
punishment_zeros <- rep(c(0),times=c(length(punishment_nofollow)))

punishment_nofollow_zeros<- data.frame(punishment_nofollow,punishment_zeros) 

colnames(punishment_nofollow_zeros) <- c("actor_2","count")

summary.pun.individ_2022G2 <- rbind(summary.pun.individ_2022G2, punishment_nofollow_zeros)




individ_pun_plot_2022G2 <-ggplot(summary.pun.individ_2022G2, aes(x = actor_2, y = count )) + 
  geom_bar(stat = "identity", width = 0.6) + # Consistent bar width
scale_y_continuous( breaks = seq(0, 60, 2)) + 
  theme_minimal()  +
  theme(
    axis.text.x = element_text(angle = 90, vjust=0.5),
    axis.title.x = element_text( face = "bold",  margin = margin(t = 15)),  # X-axis title
    axis.title.y = element_text( face = "bold",  margin = margin(r = 15)),
    ) +
  labs(
    x = "Parakeet ID",  # Add your custom x-axis title here
    y = "Count of rule follows"   # Add your custom y-axis title here
  )
individ_pun_plot_2022G2

#Calculate eveness metric
pun_for_even <- subset(summary.pun.individ_2022G2, select = c(count))

final_pun_for_even <- as.data.frame(t(pun_for_even))

pun_even_value <-calcDiv(final_pun_for_even, type = "PielouEven", q = 1) 

pun_even_value

```

#rule speed
```{r}


#######################################################################################################
#follows similar procedure as calculating the count of rule follows, with some differences in the rule detection loop and the structure

#test for length of punishment in seconds
punishment_side_by_side_tested_time_2022G2 <-  map(punishment_side_by_side_2022G2, ~ mutate(.x, initiated = ifelse(actor == counter
                                                                                                & subject_2 == counter
                                                                                                & actor_rank > subject_rank
                                                                                                , "yes","no"), 
                                      rule_followed = ifelse( subject_2 == counter
                                                              & actor == counter
                                                              & actor == subject_2
                                                              & full_time_2 - full_time > 0
                                                              & date2 == date
                                                              & actor_2_rank < subject_2_rank
                                                              & actor_rank > subject_rank
                                                              & actor_2 != subject,
                                                              "yes","no"), time_to_pun = full_time_2 - full_time))
 
 
#bind list of dfs into one df
 punishment_huge_2022G2_time <- bind_rows(punishment_side_by_side_tested_time_2022G2, .id = "column_label")
 
 #sort by time and date
 punishment_huge_2022G2_sorted_time <- punishment_huge_2022G2_time %>%
   arrange(punishment_huge_2022G2_time$date, punishment_huge_2022G2_time$full_time)
 
#remove any duplicate or double counted events 
 punishment_huge_2022G2_sorted_no_dupr_time <- subset(punishment_huge_2022G2_sorted_time, initiated != "no") 
 
 #subset for columns of interest
 punishment_no_dupr_time <- subset(punishment_huge_2022G2_sorted_no_dupr_time, select = c(date, full_time, actor, actor_rank, subject, subject_rank, 
                                                                         counter, actor_2, actor_2_rank, subject_2, subject_2_rank, full_time_2, date2, 
                                                                         initiated, rule_followed, time_to_pun,date_period))
 
 
 #get df where rule was followed
 pun_length <- subset(punishment_no_dupr_time, rule_followed != "no")
 
 #subset for columns of interest
 pun_length_ready_2022G2 <- subset( pun_length, select = c(time_to_pun, actor, subject, actor_2, subject_2, date_period))
 
 #obtain value for number of times rule was followed
 pun <- length(pun_length_ready_2022G2$time_to_pun)
 
 
 #summarize the speed of rule follows. Here we are checking whether the speed of the rule in question tends to be clustered
 summary.pun <- pun_length_ready_2022G2 %>% group_by(time_to_pun) %>%
   summarize (
     count =n())
 
 
 summary.pun
 
 
 
 #quick glimpses of the distribution of rule speeds
 ggplot(pun_length_ready_2022G2, aes(x=time_to_pun)) + 
   geom_histogram(aes(y=..density..),      
                  binwidth=50,
                  colour="black", fill="white") + 
   geom_density()
 
 
 ggplot(pun_length_ready_2022G2, aes(x=time_to_pun)) + 
   geom_histogram(aes(y=..density..),      
                  binwidth=10,
                  colour="black", fill="white") + 
   geom_density() + scale_x_continuous(name = "length in s of punishment events", limits = c(0,1000)) 
 

```

#Permutation to randomize order of events
```{r}

#######################################################################################################
#loop process

 head(aggression_no_repeat_custom)


obs <- aggression_no_repeat_custom$date_period
#separate data into a list with each df=a date period so we can randomize within days and obs periods

datelist <- list()
dater <- unique(obs)
datex <- c(1: length(unique(obs)))

#loop to create a separate dataframe for of each obs session each day 
#Each dataframe is stored within a list to make the dataframes easy to work with
#We wish to conserve our permutation randomization within days and, more specifically, within observation sessions
for(i in datex) { 
  output <- filter(aggression_no_repeat_custom, aggression_no_repeat_custom$date_period == dater[i] )
   if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
      datelist[[paste0("list", dater[i])]] <- output}

#to here



#run ids should be 1 through however many times you want to run through the permutation. Here 1000 times
runIDs <- seq(1,1000)

#create an empty masterframe to be filled with each run of the permutation
masterframe_punishment_2022G2.df <- data.frame(runIDs = character(),
                                        actor = character(),
                                        subject = character(),
                                        date = character(),
                                        full_time = integer(),
                                        counter = character(),
                                        actor_2 = character(),
                                        subject_2 = character(),
                                        full_time_2 = integer(),
                                        date2 = character(),
                                        initiated = character(),
                                        rule_followed = character(),
                                        stringsAsFactors=FALSE)


nruns <- 1000
#add progress bar
pb <- txtProgressBar(min=0, max = nruns, style = 3)
#set seed
set.seed(71)
#seed for 1404 = 20
#seed for unrestricted = 90
#seed for 402 sec = 200
#seed for 169 sec = 789
#seed fpr 150 = 706
#seed for 431 = 25
#seed for 1363 = 106
#seed = 98 for 169
#845
#71 = 414
#for loop starts here'
for (r in 1:length(runIDs)) { #r = 1 is a test you can uncomment to see if it works before #running it 100 times
  #r = 1
  run <- r
setTxtProgressBar(pb, r) #update progress bar

  loop.run <- paste("run", run) 
  loop.run
  #for example, if r = 1, looprun will say run 1 which will allow you to keep each run's #data straight
  
  
  #loop.seed <- set.seed[r] didnt work so ignore. not needed
  
  
  
  #renamed data
  loop.datalist <- datelist
  head(loop.datalist[[3]])
  xorder <- c()
  listlength <- as.numeric(length(loop.datalist))
  datex2 <- 1:listlength
  full_time_list <-c()
  loop.data.randomized.time.added <- c()
  #get rid of anything thats not actor and receiver (we will bind back on the date and time #later anyways to preserve
  #their order)
  


   for(i in datex2) {
     #store time
     full_time_list[[i]] <- subset(loop.datalist[[i]], select = c(full_time))
     #remove time from each df in the list. randomizing the order of time wouldnt make sense
     loop.datalist[[i]] <- subset(loop.datalist[[i]], select = -c(full_time))

      #create a randomized sequence of numbers to bind to the subset data (ex. 2,6,1,3,5,4)
      #sort the data frame to put the randomized numbers in order from 1 to whatever and voila, the order of events has been randomized (ex.  from above 1,2,3,4,5,6)
    xorder[[i]] <- data.frame(xorder = sample(1:length(loop.datalist[[i]]$actor),
                                              length(loop.datalist[[i]]$actor)))
    #bind random order to df
    loop.datalist[[i]] <- mutate(loop.datalist[[i]], 
                                 xorderx = xorder[[i]]$xorder)
    #put randomized order of numbers in order to randomize rows in the dataset
    loop.datalist[[i]] <-
      loop.datalist[[i]][order(loop.datalist[[i]]$xorderx),]
    #bind time back on
    loop.data.randomized.time.added[[i]] <- bind_cols(loop.datalist[[i]], full_time_list[[i]])
    loop.data.randomized.time.added[[i]] <- subset(loop.data.randomized.time.added[[i]], select = -c(xorderx))

   }

  
#separate data into list of each obs session each day for each bird
actorer <- actor
actex <- c(1:length(actor))
dater <- unique(obs)
list_split_by_bird <- c()

for(i in datex2) {
for (g in actex) {output <- subset(loop.data.randomized.time.added[[i]], actor == actorer[g] |  subject == actorer[g])
   #if one of the dfs in the list is empty, add NA
if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
  output <- mutate(output, counter = actorer[g])
  list_split_by_bird[[paste0("list", dater[i], actorer[g])]] <- output
}}

 
  #arrange by time and date
list_split_by_bird <- lapply(list_split_by_bird, function(x) arrange(x, date, full_time))

  
  
  
  
  #rename df
 punishment_aggression_list_2022_random <- list_split_by_bird


  
  #put events side by side
  punishment_side_by_side <- lapply(punishment_aggression_list_2022_random, function(x) mutate(x, actor_2 = lead(actor), 
                                                                                               actor_2_rank = lead(actor_rank), 
                                                                                               subject_2 = lead(subject), 
                                                                                               subject_2_rank = lead(subject_rank), 
                                                                                               full_time_2 = lead(full_time), 
                                                                                               date2 = lead(date)))
  

#loop to set conditionals to pick out punishment

#counter corresponds with which birds dataframe I am looking at and is used to control for false positives 
#for example, if punishment is detected for a bird A whos specific dataframe I am not looking at (im viewing bird Bs df) , this may be a false positive, as that bird A's interactions are not all listed in the dataframe of another bird B. In other words, bird A may have done something else to another bird (C, D, F etc.) that is not listed, as it would be only listed in that specific birds dataframe, that being Bird A's own dataframe   
  punishment_side_by_side_tested <-map(punishment_side_by_side, ~ mutate(.x, initiated = ifelse(actor == counter
                                                                                                & subject_2 == counter
                                                                                                & actor_rank > subject_rank
                                                                                                , "yes","no"), 
                                      rule_followed = ifelse( subject_2 == counter
                                                              & actor == counter
                                                              & actor == subject_2
                                                              & full_time_2 - full_time <= 414 
                                                              & full_time_2 - full_time > 0
                                                              & date2 == date
                                                              & actor_2_rank < subject_2_rank
                                                              & actor_rank > subject_rank
                                                              & actor_2 != subject,
                                                              "yes","no")))
  #combine dataframes into one huge dataframe
  punishment_huge_2022 <- bind_rows(punishment_side_by_side_tested, .id = "column_label")
  
  #sort by time and date
  punishment_huge_2022_sorted <- punishment_huge_2022 %>%
    arrange(punishment_huge_2022$date, punishment_huge_2022$full_time)
  

  #Remove duplicate events, events that have been represented twice, and events that could not confidently be counted as the start of punishment
    punishment_huge_2022_sorted_no_dupr <- subset(punishment_huge_2022_sorted, initiated != "no") 
  
  punishment_no_dupr <- subset(punishment_huge_2022_sorted_no_dupr, select = c(date, full_time, actor, actor_rank, subject, subject_rank, 
                                                                               counter, actor_2, actor_2_rank, subject_2, subject_2_rank, full_time_2, date2, 
                                                                               initiated, rule_followed))
  
  
  
  
  #bind vector of run IDs for each run to the dataframe so that you can have a numeric run #id next to each data point
  #that coresponds with that run which will allow you to sort/put them in ascending order later
  runIDs <- rep(loop.run, length(punishment_no_dupr$actor))
  
  punishment_complete <- cbind(runIDs, punishment_no_dupr)
  
  masterframe_punishment_2022G2.df  <- rbind(masterframe_punishment_2022G2.df, punishment_complete)
}

#######################################################LOOP#END#######################################################


#write.csv(masterframe_punishment_2022G2.df,  here("ch1 data","punishment2022G2countmaster_global_75perc.csv"), row.names = FALSE)

```

#observed rule follows vs random
```{r}



#masterframe_punishment_2022G2.df <- read_csv(here("ch1 data","punishment2022G2countmaster_global_75perc.csv"))

#put run #s in an ascending vector
#create vector for number of time the rules were followed to be stored

cum_run_IDS <- unique(masterframe_punishment_2022G2.df$runIDs)
random_times_punishment_followed <- c()


#put the number of times the rule was followed into one vector for analysis 
for (j in cum_run_IDS) {
  len_of_punishment <- length(which(masterframe_punishment_2022G2.df$rule_followed == "yes" & 
                                      masterframe_punishment_2022G2.df$runIDs == j))
  
  random_times_punishment_followed <- c(random_times_punishment_followed, len_of_punishment)
}


rand_pun_.frame_2022G2 <- data.frame(random_times_punishment_followed)

#visualize observed rule follow vs reference distribution
 pun_2022G2_count_plot <- ggplot(rand_pun_.frame_2022G2, aes(x= random_times_punishment_followed)) + 
   geom_histogram(aes(y=..density..),      
                  binwidth=2,
                  colour="black", fill="grey") +
   geom_density() + 
   geom_vline(xintercept = punishment_real_2022G2, colour = "red", size = 2) + geom_density(lwd = 1, linetype = 1) + theme_minimal() + theme(axis.text = element_text(size=25),  axis.title = element_blank()) +
  scale_x_continuous(name = "Count of Punishment events", breaks=seq(0, 50, 10)) +  
     coord_cartesian(xlim = c(0, 50), ylim = c(0,.125)) +
scale_y_continuous(breaks = seq(0, .125, .025)) 
 
 pun_2022G2_count_plot

#calculate p value (proportion of randomized reference values that are more "extreme" than the observed count of rule follows)
pun_tally2022G2 <-length(which(rand_pun_.frame_2022G2$random_times_punishment_followed > punishment_real_2022G2))

finalpun_tally2022G2 <-  1 - pun_tally2022G2/length(rand_pun_.frame_2022G2$random_times_punishment_followed)
finalpun_tally2022G2
#p = 0.022 left





#rule spec 
#1328 95%
#402 75% 0.147
# #169 50% 
# 
#


#GLOABL 
#1350 95% 0.029
#414 75% 0.156
#146 50%








# p = 0 for 1328 (95% rule spec) 
# p = 0 for 430  (75% rule spec) 
# p = 0 for 169  (50% rule spec) 



#p = 0  for 150 (50%global)
#p = 0 for 431 (75% global)
#p =  0 for 1386 (95% global)







#p = 0.001 for no restriction 
#p =  .01 for 8 sec
#p =  .052 for 85 sec
#p =  .003 for 336 sec
```



#RSavg Observed punishments (thresholds = rule specific average)
```{r}

#create empty list to store each bird each day each obs
pun_aggression_list <- list()

dater <- unique(obs)
datex <- c(1: length(unique(obs)))
actorer <- actor
actex <- c(1:length(actor))


#separate data into list of each obs session each day for each bird
for(i in datex) { 
  for (g in actex){
  output <- filter(obs_period_pt_4, obs_period_pt_4$date_period == dater[i] & (subject == actorer[g] | actor == actorer[g]))
   if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
  output <- mutate(output, counter = actorer[g])
      pun_aggression_list[[paste0("list", dater[i], actorer[g])]] <- output}}




#arrange df by date and time
pun_list_ordered <- lapply(pun_aggression_list, function(x) arrange(x, date, full_time))



pun_list_ordered <- lapply(pun_list_ordered, function(x) mutate(x, lag_actor = lag(actor),
    lag_subject = lag(subject),
     lag_time = lag(full_time),
    lag_obs_period = lag(obs_period),
    lag_time = lag(full_time),
    lag_date = lag(date),
    same_pair = actor == lag_actor & subject == lag_subject,
    time_diff = full_time - lag_time,
    keep = !same_pair | is.na(same_pair) | time_diff > 60
  )  %>%
  filter(keep) %>%
  subset(select = -c(lag_actor,lag_subject, lag_time, same_pair,time_diff, keep )))



#arrange df by date and time
pun_list_ordered <- lapply(pun_list_ordered, function(x) arrange(x, date, full_time))



#modified each data frame to put consecutive events side by side

pun_side_by_side_rsavg <- lapply(pun_list_ordered, function(x) mutate(x, actor_2 = lead(actor), actor_2_rank = lead(actor_rank), subject_2 = lead(subject), subject_2_rank = lead(subject_rank), full_time_2 = lead(full_time), date2 = lead(date))) 

#counter corresponds with which birds dataframe I am looking at and is used to control for false positives 
#for example, if punishment is detected for a bird A whos specific dataframe I am not looking at (im viewing bird Bs df) , this may be a false positive, as that bird A's interactions are not all listed in the dataframe of another bird B. In other words, bird A may have done something else to another bird (C, D, F etc.) that is not listed, as it would be only listed in that specific birds dataframe, that being Bird A's own dataframe




pun_side_by_side_rsavg_tested <- map(pun_side_by_side_rsavg, ~ mutate(.x, initiated = ifelse(actor == counter
                                                                                                & subject_2 == counter
                                                                                                & actor_rank > subject_rank
                                                                                                , "yes","no"), 
                                      rule_followed = ifelse( subject_2 == counter
                                                              & actor == counter
                                                              & actor == subject_2
                                                              & full_time_2 - full_time <= 365 
                                                              & full_time_2 - full_time > 0
                                                              & date2 == date
                                                              & actor_2_rank < subject_2_rank
                                                              & actor_rank > subject_rank
                                                              & actor_2 != subject,
                                                              "yes","no")))





#combine dataframes into one huge dataframe
pun_side_by_side_huge_rsavg <- bind_rows(pun_side_by_side_rsavg_tested, .id = "column_label")

#sort by time and date
punhugesorted_rsavg <- pun_side_by_side_huge_rsavg %>%
  arrange(pun_side_by_side_huge_rsavg$date, pun_side_by_side_huge_rsavg$full_time)




#Remove duplicate events, events that have been represented twice, and events that could not confidently be counted as the start of punishment
punhugesorted_no_dupr_rsavg <- subset(punhugesorted_rsavg, initiated != "no") 

#subset for desired columns
punishment_no_dupr_rsavg <- subset(punhugesorted_no_dupr_rsavg, select = c(date, full_time, actor, subject, 
                                                               counter, actor_2, subject_2, full_time_2, date2, 
                                                               initiated, rule_followed))
head(punishment_no_dupr_rsavg)


#determine how many time rule was follows
pun_real_rsavg <- length(which(punishment_no_dupr_rsavg$rule_followed == "yes"))

pun_real_rsavg.frame <- data.frame(pun_real_rsavg)


pun_real_rsavg.frame

#create df of only rule follows
pun_doner_all_yes_rsavg <- filter(punishment_no_dupr_rsavg, rule_followed == "yes")

#19
```






#Rsavg Individual rule following (Interval = rsavg )
```{r}
#test for individual use of rule. How many times did each bird use rule
unique(pun_doner_all_yes_rsavg$actor_2)
length(unique(pun_doner_all_yes_rsavg$actor_2)
)

rsavg_summary.pun.individ_2022G2 <- pun_doner_all_yes_rsavg %>% group_by(actor_2) %>%
  dplyr::summarize (
    count =n())

rsavg_summary.pun.individ_2022G2




#compare to raw dataset to see if any birds didnt use rule at all
unique(t2022_aggression_onlyG2$actor)
unique(t2022_aggression_onlyG2$subject)


pun_nofollow_rsavg <- setdiff(unique(t2022_aggression_onlyG2$actor), unique(pun_doner_all_yes_rsavg$actor_2)
 )

#Calculate Richness score for rule following
#DO FIRST BEFORE ADDING ZEROS AND DONT DO AGAIN WITHOUT RUNNING ABOVE CODE
pun_richness_rsavg <- length(rsavg_summary.pun.individ_2022G2$actor_2)/length((unique(t2022_aggression_onlyG2$actor)
))

pun_richness_rsavg

#add in 0s for individuals who did not follow rule for plotting purposes
pun_zeros <- rep(c(0),times=c(length(pun_nofollow_rsavg)))

pun_nofollow_rsavg_zeros<- data.frame(pun_nofollow_rsavg,pun_zeros) 

colnames(pun_nofollow_rsavg_zeros) <- c("actor_2","count")

rsavg_summary.pun.individ_2022G2 <- rbind(rsavg_summary.pun.individ_2022G2, pun_nofollow_rsavg_zeros)




# Plot
individ_pun_plot_2022G2 <- ggplot(rsavg_summary.pun.individ_2022G2, aes(x = actor_2, y = count )) + 
  geom_bar(stat = "identity", width = 0.6) + # Consistent bar width
scale_y_continuous( breaks = seq(0, 60, 2)) + 
  theme_minimal()  +
  theme(
    axis.text.x = element_text(angle = 90, vjust=0.5),
    axis.title.x = element_text( face = "bold",  margin = margin(t = 15)),  # X-axis title
    axis.title.y = element_text( face = "bold",  margin = margin(r = 15)),
    ) +
  labs(
    x = "Parakeet ID",  # Add your custom x-axis title here
    y = "Count of rule follows"   # Add your custom y-axis title here
  )
individ_pun_plot_2022G2

#Calculate eveness metric
pun_for_even_rsavg <- subset(rsavg_summary.pun.individ_2022G2, select = c(count))

final_pun_for_even_rsavg <- as.data.frame(t(pun_for_even_rsavg))

pun_even_value_rsavg <-calcDiv(final_pun_for_even_rsavg, type = "PielouEven", q = 1) 

pun_even_value_rsavg





```






#Rsavge randomize order of events
```{r}

#start with "both_numbered_sorted2022G2" as my dataset


head(both_numbered_sorted2022G2)


#procedure to randomize within obs sesh
#store first  row
row_1_obs <- both_numbered_sorted2022G2[1,]

#put events side by side. This will allow us to subtract consecutive times from one another to determine whether an event took place on a separate observation session. 
obs_period_pt_1 <- mutate(both_numbered_sorted2022G2, actor_2 = lead(actor),
actor_2_rank = lead(actor_rank), 
subject_2 = lead(subject),                                   
subject_2_rank = lead(subject_rank), 
 full_time_2 = lead(full_time), 
date2 = lead(date))

#add empty columns for observation period and date period
obs_period_pt_1$obs_period <- NA
obs_period_pt_1$date_period <- NA

#create ticker for loop
q <- 1

#loop to calculate observation sessions. If consecutive events were on the same day but there was a long time between events, it is determined that the 2nd event took place in a separate observation session. The ticker goes up, and is reset if events took place on different days.
for (i in 1:nrow(obs_period_pt_1)) {ifelse(obs_period_pt_1$date[i] == obs_period_pt_1$date2[i]
                                            & obs_period_pt_1$full_time_2[i] - obs_period_pt_1$full_time[i] >= 1800, q <- q+1, q <- q)
  ifelse(obs_period_pt_1$date[i] != obs_period_pt_1$date2[i], q <- 1, q <- q)
  obs_period_pt_1$obs_period [i] <- q
  }

#remove original columns and keep mutated columns. These will replace original columns
obs_period_pt_2 <- subset(obs_period_pt_1, select = c(date2,  full_time_2, actor_2, actor_2_rank, subject_2, subject_2_rank, obs_period, date_period  ))

#get rid of the bottom, which will be full of NAs, as the last event in the original data does not have anything that comes after
obs_period_pt_3 <- slice(obs_period_pt_2, 1:(n() - 1))

#rename mutated columns with original names 
setnames(obs_period_pt_3, old = c('date2',  'full_time_2', 'actor_2', 'actor_2_rank', 'subject_2', 'subject_2_rank'), 
         new = c('date',  'full_time', 'actor', 'actor_rank', 'subject', 'subject_rank'))

#add observation period "1" onto first row that we subseted off, as the first event in the data has to take place in obs session 1
row_1_obs$obs_period <- 1 
row_1_obs$date_period <- NA

#ensure that row 1 matches the df
row_1_obs_ready <-  subset(row_1_obs, select = c(date,  full_time, actor, actor_rank, subject, subject_rank, obs_period, date_period  ))

#bind first row and df together
obs_period_pt_4 <- rbind(row_1_obs_ready, obs_period_pt_3)
#add date period 
obs_period_pt_4$date_period <- paste(obs_period_pt_4$date,obs_period_pt_4$obs_period)



obs <- obs_period_pt_4$date_period
#separate data into a list with each df=a date period so we can randomize within days and obs periods

datelist <- list()
dater <- unique(obs)
datex <- c(1: length(unique(obs)))

#loop to create a separate dataframe for of each obs session each day 
#Each dataframe is stored within a list to make the dataframes easy to work with
#We wish to conserve our permutation randomization within days and, more specifically, within observation sessions
for(i in datex) { 
  output <- filter(obs_period_pt_4, obs_period_pt_4$date_period == dater[i] )
   if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
      datelist[[paste0("list", dater[i])]] <- output}




#run ids should be 1 through however many times you want to run through the permutation. Here 1000 times
runIDs <- seq(1,1000)

#create an empty masterframe to be filled with each run of the permutation
masterframe_punishment_rsavg.df_2022G2 <- data.frame(runIDs = character(),
                             actor = character(),
                             subject = character(),
                             date = character(),
                             full_time = integer(),
                             counter = character(),
                             actor_2 = character(),
                             subject_2 = character(),
                             full_time_2 = integer(),
                             date2 = character(),
                             initiated = character(),
                             rule_followed = character(),
                             stringsAsFactors=FALSE)
nruns <- 1000
#progress bar so we can keep track of progress
pb = txtProgressBar(min=0, max = nruns, style = 3)

#seed 87

#set seed
set.seed(7)
#forloop starts here'
for (r in 1:length(runIDs)) { #r = 1 is a test you can uncomment to see if it works before running it 1000 times
#r = 1
  run <- r
    setTxtProgressBar(pb, r) #update progress bar
  loop.run <- paste("run", run) 
  loop.run
  #for example, if r = 1, looprun will say run 1 which will allow you to keep each run's data straight
  
    

  
  #renamed data
  loop.datalist <- datelist
  head(loop.datalist[[3]])
  
  
  xorder <- c()
  listlength <- as.numeric(length(loop.datalist))
  datex2 <- 1:listlength
  full_time_list <-c()
  loop.data.randomized.time.added <- c()
  #get rid of anything thats not actor and receiver (we will bind back on the date and time #later anyways to preserve
  #their order)
  


   for(i in datex2) {
     #store time
     full_time_list[[i]] <- subset(loop.datalist[[i]], select = c(full_time))
     #remove time from each df in the list. randomizing the order of time wouldnt make sense
     loop.datalist[[i]] <- subset(loop.datalist[[i]], select = -c(full_time))

      #create a randomized sequence of numbers to bind to the subset data (ex. 2,6,1,3,5,4)
      #sort the data frame to put the randomized numbers in order from 1 to whatever and voila, the order of events has been randomized (ex.  from above 1,2,3,4,5,6)
    xorder[[i]] <- data.frame(xorder = sample(1:length(loop.datalist[[i]]$actor),
                                              length(loop.datalist[[i]]$actor)))
    #bind random order to df
    loop.datalist[[i]] <- mutate(loop.datalist[[i]], 
                                 xorderx = xorder[[i]]$xorder)
    #put randomized order of numbers in order to randomize rows in the dataset
    loop.datalist[[i]] <-
      loop.datalist[[i]][order(loop.datalist[[i]]$xorderx),]
    #bind time back on
    loop.data.randomized.time.added[[i]] <- bind_cols(loop.datalist[[i]], full_time_list[[i]])
    loop.data.randomized.time.added[[i]] <- subset(loop.data.randomized.time.added[[i]], select = -c(xorderx))

   }
#rename df
  pun_rsavglist <- loop.data.randomized.time.added
  
  
#separate data into list of each obs session each day for each bird
actorer <- actor
actex <- c(1:length(actor))
dater <- unique(obs)
pun_rsavglist_split_by_bird <- c()

for(i in datex2) {
for (g in actex) {output <- subset(pun_rsavglist[[i]], actor == actorer[g] |  subject == actorer[g])
#if one of the dfs in the list is empty, add NA
   if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
  output <- mutate(output, counter = actorer[g])
  pun_rsavglist_split_by_bird[[paste0("list", dater[i], actorer[g])]] <- output
}}


#sort by date and time
pun_rsavglist_ordered <- lapply(pun_rsavglist_split_by_bird, function(x) arrange(x, date, full_time))




#modified each data frame to put consecutive events side by side
pun_rsavglist_ordered_mut <- lapply(pun_rsavglist_ordered, function(x) mutate(x, actor_2 = lead(actor), actor_2_rank = lead(actor_rank), subject_2 = lead(subject), subject_2_rank = lead(subject_rank), full_time_2 = lead(full_time), date2 = lead(date))) 




#loop to set conditionals to pick out punishment

#counter corresponds with which birds dataframe I am looking at and is used to control for false positives 
#for example, if punishment is detected for a bird A whos specific dataframe I am not looking at (im viewing bird Bs df) , this may be a false positive, as that bird A's interactions are not all listed in the dataframe of another bird B. In other words, bird A may have done something else to another bird (C, D, F etc.) that is not listed, as it would be only listed in that specific birds dataframe, that being Bird A's own dataframe



 pun_rsavglist_ordered_mut_tested <- map(pun_rsavglist_ordered_mut, ~ mutate(.x, initiated = ifelse(actor == counter
                                                                                                & subject_2 == counter
                                                                                                & actor_rank > subject_rank
                                                                                                , "yes","no"), 
                                      rule_followed = ifelse( subject_2 == counter
                                                              & actor == counter
                                                              & actor == subject_2
                                                              & full_time_2 - full_time <= 365 
                                                              & full_time_2 - full_time > 0
                                                              & date2 == date
                                                              & actor_2_rank < subject_2_rank
                                                              & actor_rank > subject_rank
                                                              & actor_2 != subject,
                                                              "yes","no")))











#combine dataframes into one huge dataframe
pun_rsavghuge <- bind_rows(pun_rsavglist_ordered_mut_tested, .id = "column_label")

#sort by time and date
pun_rsavghugesorted <- pun_rsavghuge %>%
  arrange(pun_rsavghuge$date, pun_rsavghuge$full_time)



#Remove duplicate events, events that have been represented twice, and events that could not confidently be counted as the start of punishment
pun_rsavghugesorted_no_dup <- subset(pun_rsavghugesorted, initiated != "no") 

#subset for columns of interest
punishment_no_dup <- subset(pun_rsavghugesorted_no_dup, select = c(date, full_time, actor, subject, 
                                                               counter, actor_2, subject_2, full_time_2, date2, 
                                                               initiated, rule_followed))

head(punishment_no_dup)




  
  #end of punishment process
  #end of rule code
  
  
  
  #bind vector of run IDs for each run to the dataframe so that you can have a numeric run #id next to each data point
  #that coresponds with that run which will allow you to sort/put them in ascending order later
 runIDs <- rep(loop.run, length(punishment_no_dup$actor))
 
  punishment_complete <- cbind(runIDs, punishment_no_dup)
  
  
#put each runs data into teh masterframe
  masterframe_punishment_rsavg.df_2022G2 <- rbind(masterframe_punishment_rsavg.df_2022G2, punishment_complete)
  
}

#write.csv(masterframe_punishment_rsavg.df_2022G2, "C:/Users/Xmanf/Desktop/Ch1 TDR files/pun2022G2countmaster_rsavg.csv")
```



#Rsavg Compare observed rule follows to reference model
```{r}
#masterframe_punishment_rsavg.df_2022G2 <- read.csv("C:/Users/Xmanf/Desktop/Ch1 TDR files/pun2022G2countmaster_rsavg.csv", header=TRUE, stringsAsFactors=FALSE)





#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_punishment_rsavg.df_2022G2$runIDs)

#create empty vector for number of time the rules were followed to be stored
random_times_punishment_followed <- c()


#put the number of times the rule was followed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_punishments <- length(which(masterframe_punishment_rsavg.df_2022G2$rule_followed == "yes" & 
                                        masterframe_punishment_rsavg.df_2022G2$runIDs == j))

random_times_punishment_followed <- c(random_times_punishment_followed, len_of_punishments)
}

#turn vector into a dataframe for analysis
rand_pun_rsavg.frame <- data.frame(random_times_punishment_followed)

#plot (the x intercept is the # of time a rule was used in the real data, found with 
#"pun_real <- length(which(punishment_no_dupr$rule_followed == "yes"))

#visualize observed rule follow vs reference distribution
pun_2022G2_count_plot <- ggplot(rand_pun_rsavg.frame, aes(x=random_times_punishment_followed)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=1,
                 colour="black", fill="grey") +
  geom_density(lwd = 1, linetype = 1) + theme_minimal() +
  geom_vline(xintercept = pun_real_rsavg, colour = "red", size = 2)  + 
  theme(axis.text = element_text(size=25),
        axis.title = element_text(size = 25, face = "bold")) +
  scale_x_continuous(name = "Count of punishment events", breaks=seq(0, 100, 50)) +  
     coord_cartesian(xlim = c(0, 100), ylim = c(0,.15)) +
scale_y_continuous( breaks = seq(0, .15, .05)) +
  labs(
       y = "Density")

pun_2022G2_count_plot

#calculate p value (proportion of randomized reference values that are more "extreme" than the observed count of rule follows)
pun_tally_rsavg_2022G2 <-length(which(pun_real_rsavg <rand_pun_rsavg.frame$random_times_punishment_followed ))

ppun_tally_rsavg_2022G2 <-   1 - pun_tally_rsavg_2022G2/length(rand_pun_rsavg.frame$random_times_punishment_followed)
ppun_tally_rsavg_2022G2
#p = 0.006

```









#randomize rule speed
```{r}
#Here we are randomizing time between events for the permutation
#randomize time between initiation and response between events
#start with 
obs <- aggression_no_repeat_custom$date_period



#separate data into a list with each df=a date 


#create empty list to store each bird each day each obs
practdrlist_ordered <- list()

dater <- unique(obs)
datex <- c(1: length(unique(obs)))
actorer <- actor
actex <- c(1:length(actor))


#separate data into list of each obs session each day for each bird
for(i in datex) { 
  for (g in actex){
  output <- filter(aggression_no_repeat_custom, aggression_no_repeat_custom$date_period == dater[i] & (subject == actorer[g] | actor == actorer[g]))
   if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
  output <- mutate(output, counter = actorer[g])
      practdrlist_ordered[[paste0("list", dater[i], actorer[g])]] <- output}}







#arrange df by date and time
practdrlist_ordered <- lapply(practdrlist_ordered, function(x) arrange(x, date, full_time))



#randomize time between initiation and response between events

punishment_side_by_side_2022G2 <- lapply(practdrlist_ordered, function(x) mutate(x, actor_2 = lead(actor), 
                                                                                      actor_2_rank = lead(actor_rank), 
                                                                                      subject_2 = lead(subject), 
                                                                                      subject_2_rank = lead(subject_rank), 
                                                                                      full_time_2 = lead(full_time), 
                                                                                      date2 = lead(date))) 
#Loop to detect instances and speed of rule use for the rule in question 
 punishment_side_by_side_tested_time_2022G2 <-  map(punishment_side_by_side_2022G2, ~ mutate(.x, initiated = ifelse(actor == counter
                                                                                                & subject_2 == counter
                                                                                                & actor_rank > subject_rank
                                                                                                , "yes","no"), 
                                      rule_followed = ifelse( subject_2 == counter
                                                              & actor == counter
                                                              & actor == subject_2
                                                              & full_time_2 - full_time > 0
                                                              & date2 == date
                                                              & actor_2_rank < subject_2_rank
                                                              & actor_rank > subject_rank
                                                              & actor_2 != subject,
                                                              "yes","no"), time_to_pun = full_time_2 - full_time))
 
 
 punishment_huge_2022G2_time <- bind_rows(punishment_side_by_side_tested_time_2022G2, .id = "column_label")
 

 
 
 
 

#Remove duplicate and double counted sequences
punishment_huge_2022G2_time <- distinct(punishment_huge_2022G2_time, date, full_time, actor, actor_rank, subject, subject_rank, obs_period, date_period, actor_2, subject_2, full_time_2, date2, time_to_pun, .keep_all= TRUE) 



#remove instances of rule use whos speed were calculated to be 0. In this case we cant be certain of which event actually came first
punishment_huge_2022G2_time <- filter(punishment_huge_2022G2_time, time_to_pun != 0)


date_x <- punishment_huge_2022G2_time$date
obs <- punishment_huge_2022G2_time$date_period
#separate data into a list with each df=a date so we can randomize within days
#sort_by_date <- function(x, i, y) {y <- filter(x, date == i )
#mutate(y, date_counter = i)}

rule_time_ready_for_rand_list <- list()
dater <- unique(obs)
datex <- c(1: length(unique(obs)))
#boop <- filter(both_numbered_sorted2022G2, both_numbered_sorted2022G2$date == datex [2])

#loop to create a separate dataframe for each bird that only contains interactions #pertaining to that bird.
#Each dataframe is stored within a list to make the dataframes easy to work with
#separate data into list of each obs session each day 
for(i in datex) { 
  output <- filter(punishment_huge_2022G2_time, punishment_huge_2022G2_time$date_period == dater[i] )
   if (nrow(output) < 1) {output [ nrow(output) + 1 , ] <- NA
    } 
      rule_time_ready_for_rand_list[[paste0("list", dater[i])]] <- output}






 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 

pb = txtProgressBar(min=0, max = nruns, style = 3)

#run ids should be 1 through however many times you want to run through the permutation
runIDs <- seq(1,1000)

#create an empty masterframe to be filled with each run of the permutation
masterframe_punishment.df_time_2022G2 <- data.frame(runIDs = character(),
                             actor = character(),
                             subject = character(),
                             date = character(),
                             full_time = integer(),
                             counter = character(),
                             actor_2 = character(),
                             subject_2 = character(),
                             full_time_2 = integer(),
                             date2 = character(),
                             initiated = character(),
                             rule_followed = character(),
                             time_to_pun = integer(),
                             date_period = character(),
                             stringsAsFactors=FALSE)


set.seed(81)
#forloop starts here'
for (r in 1:length(runIDs)) { #r = 1 is a test you can uncomment to see if it works before #running it 100 times
#r = 1
  run <- r
      setTxtProgressBar(pb, r) #update progress bar
  loop.run <- paste("run", run) 
  loop.run
  #for example, if r = 1, looprun will say run 1 which will allow you to keep each run's #data straight
  
    

  
  #renamed data
  loop.datalist <- rule_time_ready_for_rand_list
  head(loop.datalist[[3]])
  xorder <- c()
  listlength <- as.numeric(length(loop.datalist))
datex2 <- 1:listlength
  full_time_to_rule <-c()
  loop.data.randomized.time_to_rule.added <- c()
    
    

  for(i in datex2) {
         #store time
     full_time_to_rule[[i]] <- subset(loop.datalist[[i]], select = c(time_to_pun))     
     #remove time from each df in the list. randomizing the order of time wouldnt make sense
     loop.datalist[[i]] <- subset(loop.datalist[[i]], select = -c(time_to_pun))
      #create a randomized sequence of numbers to bind to the subset data (ex. 2,6,1,3,5,4)
      #sort the data frame to put the randomized numbers in order from 1 to whatever and voila, the time between events (speed) has been  randomized (ex. from above 1,2,3,4,5,6)
    xorder[[i]] <- data.frame(xorder = sample(1:length(full_time_to_rule[[i]]$time_to_pun),
                                              length(full_time_to_rule[[i]]$time_to_pun)))
    #bind random order to df
    full_time_to_rule[[i]] <- mutate(full_time_to_rule[[i]], 
                                 xorderx = xorder[[i]]$xorder)
    #put randomized order of numbers in order to randomize speed in the dataset
    full_time_to_rule[[i]] <-
      full_time_to_rule[[i]][order(full_time_to_rule[[i]]$xorderx),]
    #bind time back on
    loop.data.randomized.time_to_rule.added[[i]] <- bind_cols(loop.datalist[[i]], full_time_to_rule[[i]])
    loop.data.randomized.time_to_rule.added[[i]] <- subset(loop.data.randomized.time_to_rule.added[[i]], select = -c(xorderx))

   }
  

  
  
  
  
  
  
  

#combine dataframes into one huge dataframe
loop.datalist_time_randomized_binded_tested_pun_2022G2 <- bind_rows(loop.data.randomized.time_to_rule.added, .id = "column_label")


 

#sort by time and date
loop.datalist_time_randomized_binded_tested_pun_2022G2 <- loop.datalist_time_randomized_binded_tested_pun_2022G2 %>%
  arrange(loop.datalist_time_randomized_binded_tested_pun_2022G2$date, loop.datalist_time_randomized_binded_tested_pun_2022G2$full_time)






#Remove duplicate events, events that have been represented twice, and events that could not confidently be counted as the start of punishment
testhugesorted_no_dup_rand_time <- subset(loop.datalist_time_randomized_binded_tested_pun_2022G2, initiated != "no") 

punishment_no_dup_rand_time <- subset(testhugesorted_no_dup_rand_time, select = c(date, full_time, actor, subject, 
                                                               counter, actor_2, subject_2, full_time_2, date2, 
                                                               initiated,  rule_followed,time_to_pun,date_period))





  
  #end of punishment process
  #end of rule code
  
  
  
  #bind vector of run IDs for each run to the dataframe so that you can have a numeric run #id next to each data point
  #that coresponds with that run which will allow you to sort/put them in ascending order later
 runIDs <- rep(loop.run, length(punishment_no_dup_rand_time$actor))
 
  punishment_complete_rand_time_2022G2 <- cbind(runIDs, punishment_no_dup_rand_time)
  
  
#put each runs data into teh masterframe
  masterframe_punishment.df_time_2022G2 <- rbind(masterframe_punishment.df_time_2022G2, punishment_complete_rand_time_2022G2)
  
}


#write.csv(masterframe_punishment.df_time_2022G2,  here("ch1 data","pun2022G2timemaster.csv"), row.names = FALSE)

```


#rule speed vs random
```{r}
#masterframe_punishment.df_time_2022G2 <- read_csv(here("ch1 data","pun2022G2timemaster.csv"))

head(masterframe_punishment.df_time_2022G2)

#subset masterframe so that only rule follows are included
masterframe_punishment.df_time_2022G2_only_follows <- subset(masterframe_punishment.df_time_2022G2, rule_followed == "yes" )

#isolate time it takes to carry out rule...
masterframe_pun_time_sub.df_2022G2 <- subset( masterframe_punishment.df_time_2022G2_only_follows, select = c(actor, subject, actor_2,subject_2, time_to_pun,date_period))
head(masterframe_pun_time_sub.df_2022G2)



#add data label for the randomized rule speed
masterframe_pun_time_ready.df <- mutate(masterframe_pun_time_sub.df_2022G2, data_type = "randomized")

#ensure that time is numeric
masterframe_pun_time_ready.df$time_to_pun <- as.numeric(masterframe_pun_time_ready.df$time_to_pun)


head(pun_length_ready_2022G2)

#add data label for the observed rule speed
pun_length_ready_2022G2_label <- mutate(pun_length_ready_2022G2, data_type = "observed")


#combine df for observed rule speed and randomized rule speed into one df
pun_length_ready_2022G2_comb <- rbind(masterframe_pun_time_ready.df, pun_length_ready_2022G2_label )
unique(pun_length_ready_2022G2_comb$data_type)




#provide label for rule in question
pun_length_ready_2022G2_comb <- mutate(pun_length_ready_2022G2_comb, rule = "punishment")


#make sure that rule speed is numeric
pun_length_ready_2022G2_comb$time_to_pun <- as.numeric(pun_length_ready_2022G2_comb$time_to_pun)


#observed vs randomized raw plot. quick look at how the distriubtuions compare
ggplot(pun_length_ready_2022G2_comb, aes(x=time_to_pun, fill = data_type)) + 
  geom_histogram(aes(y=..density..),      
                 binwidth=1) +#,
                 #colour="black", fill="white") + 
  geom_density() +   scale_x_continuous(name = "length in s of pun events random vs observed", limits = c(0,300)) 


#obtain mode of observed and randomized rule speed for plotting purposes
#getmode <- function(v) {
   #uniqv <- unique(v)
   #uniqv[which.max(tabulate(match(v, uniqv)))]
#}

#pun_2022G2_mode_obs <- getmode(pun_length_ready_2022G2_label$time_to_pun)

#pun_2022G2_mode_rand <-getmode(masterframe_pun_time_ready.df$time_to_pun)

#bind mode to dfs
#pun_2022G2_mode <-  rbind(pun_2022G2_mode_obs,pun_2022G2_mode_rand)

#test for homogeneity of variance
#leveneTest(time_to_pun ~ data_type, data = pun_length_ready_2022G2_comb)


#split violin plot of observed vs randomized rule speed
pun_violin_plot <- ggplot(pun_length_ready_2022G2_comb, aes(rule, time_to_pun, fill = data_type))  +
  introdataviz::geom_split_violin(alpha= .9, trim = TRUE, show.legend = FALSE) +
  geom_boxplot(width = .2, alpha = .2, fatten = 2, show.legend = FALSE, outlier.shape = NA) +
  #stat_summary(fun = "mean", geom = "point",  colour = "darkblue", show.legend = F, 
               #position = position_dodge(.200))  +
  scale_x_discrete(name = "Time Dependent Rule") +  
     coord_cartesian(ylim = c(0, 1000)) +
  scale_y_continuous(name = "Speed of punishment events (s)",
                     breaks = seq(0, 1000, 200) 
                     ) + scale_fill_manual(values=c("red","gray")) +
  theme_minimal(base_size = 20)  + theme(axis.text.y = element_text(size=25), axis.title= element_blank(), axis.text.x = element_blank())


pun_2022G2_time_plot <- pun_violin_plot

pun_2022G2_time_plot

#calculate the means for observed and randomized rule speed
mean_obs_rand_punishment_2022G2 <- pun_length_ready_2022G2_comb %>% group_by(data_type) %>%
  dplyr::summarize (mean(time_to_pun))

mean_obs_rand_punishment_2022G2

#obs 187.8182
#rand 295.8466





median_obs_rand_pun_2022G2 <- pun_length_ready_2022G2_comb %>% group_by(data_type) %>%
  dplyr::summarize (median(time_to_pun))

median_obs_rand_pun_2022G2

#obs 102
#rand 127







 
 
 #linear mixed model to compare means of observed vs randomized rule speed

m1_pun_2022G2 <- glmer(time_to_pun ~ data_type  + (1 | date_period) , data = pun_length_ready_2022G2_comb ,   control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5)), family = Gamma(link = "log"))
summary(m1_pun_2022G2)
  
#p = 0.282            



#find confidence intervals
confint_pun_2022G2 <- round(confint(m1_pun_2022G2, method = "Wald", level = 0.95), 3)
confint_pun_2022G2
#(Intercept)          4.688  5.867
#data_typerandomized -0.169  0.900
#tukey
summary(glht(m1_pun_2022G2, linfct = mcp(data_type="Tukey")), test = adjusted("none"))
##r2 values
library(MuMIn)
r.squaredGLMM(m1_pun_2022G2 )


```
#Earth movers distance
```{r}


#ensure that observed rule speed is numeric
pun_length_ready_2022G2$time_to_pun <- as.numeric(pun_length_ready_2022G2$time_to_pun)


head(pun_length_ready_2022G2)

#convert to a density matrix
#summarize the count of various rule speeds

summary.pun_length_ready_2022G2 <- pun_length_ready_2022G2 %>% group_by(time_to_pun) %>%
  dplyr::summarize (
    count =n())

summary.pun_length_ready_2022G2
#Add density column
summary.pun_length_ready_2022G2_density <- mutate(summary.pun_length_ready_2022G2, density = count/sum(count))
#remove count column
summary.pun_length_ready_2022G2_density_no_count <- subset(summary.pun_length_ready_2022G2_density, select = -c(count))
#reorder the columns
summary.pun_length_ready_2022G2_density_no_count_reordered <- summary.pun_length_ready_2022G2_density_no_count[, c(2, 1)]
head(summary.pun_length_ready_2022G2_density_no_count_reordered)
#convert to matrix
summary.pun_length_ready_2022G2_density_no_count_reordered_mat <- as.matrix(summary.pun_length_ready_2022G2_density_no_count_reordered)





masterframe_length_ready <- subset(masterframe_pun_time_sub.df_2022G2, select = c(time_to_pun))
head(masterframe_length_ready)



#ensure that observed rule speed is numeric
masterframe_length_ready$time_to_pun <- as.numeric(masterframe_length_ready$time_to_pun)






#convert to a density matrix
#summarize the count of various rule speeds

summary.masterframe_length_ready <- masterframe_length_ready %>% group_by(time_to_pun) %>%
  dplyr::summarize (
    count =n())

summary.masterframe_length_ready

#Adddensity column
summary.masterframe_length_ready_density <- mutate(summary.masterframe_length_ready, density = count/sum(count))
#remove count column
summary.masterframe_length_ready_density_no_count <- subset(summary.masterframe_length_ready_density, select = -c(count))
#reorder the columns
summary.masterframe_length_ready_density_no_count_reordered <- summary.masterframe_length_ready_density_no_count[, c(2, 1)]
head(summary.masterframe_length_ready_density_no_count_reordered)
#convert to matrix
summary.masterframe_length_ready_density_no_count_reordered_mat <- as.matrix(summary.masterframe_length_ready_density_no_count_reordered)


#Calculate EMD between observed and randomized rule speed
pun_obs_vs_rand_emd <- emdist::emd(summary.pun_length_ready_2022G2_density_no_count_reordered_mat, summary.masterframe_length_ready_density_no_count_reordered_mat)
pun_obs_vs_rand_emd
# 119.1294
```
